### 2021-09-07 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_CalculateTPM.Rmd
Fix consolidated similar stages and gam marker plot ordering

---
title: "P.berghei pir calculate TPM"
author: "Timothy Little"
date: "31/08/2021"
output: 
  html_document:
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---

```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, results = TRUE, message = FALSE, warning = FALSE)
options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
```

```{r load_packages, include = FALSE, message = FALSE}
library('readxl')
library('readr')
library('dplyr')
library('purrr')
library('ggplot2')
library('DESeq2')
library("RColorBrewer")
library('colorspace')
library('data.table')
library('gridExtra')
library('tibble')
library('scales')
library('stringr')
library('kableExtra')
library('viridis')
library('corrplot')
library('reshape2')
library('circlize') # for circular plots
library('cluster')    # clustering algorithms
library('factoextra') # clustering algorithms & visualization
library('ggfortify') # so ggplot understands prcomp objects
library('ape') # for phylogeny plotting
library('dendextend') # for dendrogram plotting
library('plotly')
library('ggpattern')
library('ComplexHeatmap')
library('ggdendro') #Drawing ggplot dendrograms

#Colour blind friendly palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
show_col(cbPalette)

#Want the defaults for arrange, select, mutate and summarise to be from dplyr
arrange <- dplyr::arrange
mutate <- dplyr::mutate
summarise <- dplyr::summarise
select <- dplyr::select

```

```{r sourcing_functions}
source('../rfunctions/uniquify_fun.R')
source('../rfunctions/ordering_fun.R')
source('../rfunctions/tpm_fun.R')
source('../rfunctions/usefulFunctions_mround_my-max_firstup_fun.R')
```

```{r ordering_vectors}
order_vector <- c('Sporo', 'Liv-pre4h', 'Liv-pre47h', 'Liv-post48h',
                 'Liv-2h','Liv-4h','Liv-12h', 'Liv-18h',
                 'Liv-24h' ,'Liv-36h','48h','54h','60h',
                 'Mero','Asex','Schiz','Gam','-Pre3h','BldMl-24h','Ook')

r_order_vector <- str_replace_all(order_vector, 
                                  pattern = '-', 
                                  replacement = '\\.')
```

```{r pir_id, message=FALSE, results='hide'}

# Getting the pir gene names and transcript lengths needed for calculating the transcript-per-million

# First we need to get the _pir gene names. These have been downloaded using a search of PlasmoDB P. berghei ANKA genome v3 using  "\*pir\*" search and manual removal of genes which clearly were not pir.

allgene.info.file <- '../standard_files/PlasmoDb-47_beta_Pberghei_GenesByTaxon_Summary.csv'
allgene.info <- as_tibble(
  read_csv(allgene.info.file)
) %>%
  mutate(Geneid = `Gene ID`, 
         Description = `Product Description`,
         is_pseudo = `Is Pseudo`,
         Transcript.Length = `Transcript Length`,
         paralog_count = `Paralog count`) %>% 
  select(Geneid, 
         Description, 
         is_pseudo, 
         Transcript.Length, 
         paralog_count)

pir_info <- allgene.info %>% 
  filter(str_detect(Description, 
                    pattern = 'PIR(?!.+pseud)'), 
         is_pseudo == 'No') %>% 
  filter(!Geneid %in% 'PBANKA_0524600')
#Removing the pir-like gene.

pir_id <- pir_info$Geneid

#pir

pir_genes_list <- lapply(
  unique(pir_info$Description), 
  function(pir_group){
    select(
      filter(
        pir_info, 
        Description %in% pir_group
      ),
      Geneid
    )
  }
)
names(pir_genes_list) <- unique(pir_info$Description)

#How many pirs are present on each chromosome?

pir_tbl = as_tibble(data.frame(Geneid = pir_id))
pir_chrom_tbl = pir_tbl %>% 
  mutate(chrom = unlist( #Getting the chromosomes by a complicated set of lapply and sapplys...
    sapply(
      lapply(
        sapply(
          as.character(pir_tbl$Geneid), 
          strsplit,
          split = ''
        ), 
        '[', 
        8:9
      ), 
      paste, 
      collapse = ''
    )
  )
  ) 
pir_chrom_tbl %>% 
  group_by(chrom) %>% 
  dplyr::summarise(number_of_pirs_in_each_chromosome = n(),
                   percentage_of_all_pirs = round(n()/nrow(pir_tbl), 3)*100) %>% 
  dplyr::arrange(desc(percentage_of_all_pirs))

#Getting the loci information
pir_loci <- data.frame(
  Geneid = pir_id,
  chrom = str_extract(pir_id, 
                      pattern = '(?<=_)[[:digit:]]{2}'),
  number = as.numeric(str_extract(pir_id, 
                                  pattern = '(?<=_[[:digit:]]{2})[[:digit:]]+'))
) %>% 
  filter(!chrom %in% '00') %>% #Remove chrom00 (contigs)
  group_by(chrom) %>% 
  mutate(jump = ifelse(lag(number, 
                           default = 1001) < 1000 & number - lag(number, 
                                                                 default = 0) > 1000,
                       1,
                       0)) %>% 
  mutate(
    jump = ifelse(cumsum(jump) == 1 | jump == 1, 1, 0)
  ) %>% 
  mutate(
    locus = paste0(
      chrom,
      ifelse(jump == 0 & number < 1000, 
             'Left',
             'Right')
    )
  )

# Detect 'jumps' of the gene number over 1000 within each chromosome, setting the lag default as 1001 in the
#   first part of the ifelse statement so that chrom2 doesn't have an erroneous jump as it starts at >1000
# Later cumsum is do that every gene after the 'jump' is included allowing us to define telomere ends.

pir_clade <- read_delim('../standard_files/Pberghei_pir_clades.txt', 
                        delim = '\t')
pir_clade <- mutate(pir_clade, 
                    locus = pir_loci$locus[match(Geneid, pir_loci$Geneid)])
pir_id <- pir_clade$Geneid

pir_clade_list <- lapply(unique(pir_clade$clade), function(clade_name){
  pir_clade$Geneid[pir_clade$clade %in% clade_name]
})
names(pir_clade_list) <- as.character(unique(pir_clade$clade))
```

```{r transcript_redundancy, message = FALSE, results='hide'}

#Note that there is some redundancy
two.transcript <- allgene.info %>% 
  dplyr::count(Geneid) %>% 
  filter(n > 1)

#Nine genes have two transcripts - need to give them one average of the transcripts since featurecounts has only
# calculated one count for the locus
two.transcript.info <- allgene.info[allgene.info$Geneid %in% two.transcript$Geneid,]
two.transcript.mean = aggregate(Transcript.Length~Geneid, data=two.transcript.info, 
                                FUN=function(x) c(mean=mean(x)))

#Remove the first match
allgene.info <- allgene.info[-(match(two.transcript.mean$Geneid, 
                                     allgene.info$Geneid)),]

#For the second match replace the transcript value with the mean calculated with aggregate
allgene.info$Transcript.Length <- as.double(allgene.info$Transcript.Length)
allgene.info[match(two.transcript.mean$Geneid, 
                   allgene.info$Geneid),]$Transcript.Length <- two.transcript.mean$Transcript.Length
lengths.transcript <- subset(allgene.info, select = c(Geneid, Transcript.Length), drop = F)
```

## Read in counts and calculate TPM

```{r read_files_calc_TPM}
#Reading in the counts from file
berghei_counts <- read.csv('tl20-05_berghei_counts.csv',
                               row.names = 1)

berghei_tpm <- tpm(count_data = berghei_counts,
                   lengths = lengths.transcript)

apply(berghei_tpm[-1], 2, sum)
```

```{r check_tpm}

#Used `dput` function to create a df of randomly chosen part of what I assume to be the true values.
check_tpm <- structure(list(E03_Gam_1 = c(578.506468410824, 43.8844462052653, 
                                          93.7895540428778, 55.9251999927489, 28.3887997796848, 49.7689021855188
), 
E03_Gam_2 = c(491.460414371885, 38.6239066701, 119.30578922894, 
              148.452713617397, 38.8846967267782, 25.7641457887425), 
E03_Gam_3 = c(824.81294645121, 
              48.1865616211443, 92.5270154766448, 74.1594147167308, 34.8902736305123, 
              74.9586138010564), 
E03_Ook_1 = c(340.238340611313, 206.852529849517, 
              43.9667720316821, 26.3982623497325, 14.9136412690561, 6.22485349746656
), 
E03_Ook_2 = c(713.689689765822, 178.051382530752, 58.5009548960498, 
              52.6871822865333, 18.8623926433347, 12.0840592930611), 
E03_Ook_3 = c(409.621633449974, 
              104.471439607932, 38.7705779604298, 31.3685913631761, 20.1312670544009, 
              14.7857267550803)), 
row.names = 3345:3350, 
class = "data.frame")

all.equal(berghei_tpm[3345:3350,colnames(check_tpm)],check_tpm)

```

```{r save_tpm_oldNames_allSamples}
#Save the 'raw' TPM before any samples have been removed or sample names combined.

write.csv(berghei_tpm,
          'berghei_tpm_oldNamesAllSamples.csv')

```

```{r remove_E06-6h}
#Remove the Caldelari (E06) 6h samples, as they did in their paper.
berghei_tpm[,str_subset(colnames(berghei_tpm), 'E06_Liv.6h')] <- NULL
```

# Hierarchical clustering plots.

```{r hclusts, message = FALSE}
experiment_id <- unique(na.omit(str_extract(colnames(berghei_tpm), pattern = 'E[[:digit:]]{2}')))

lapply(experiment_id, 
       function(experiment_name){
         tpm_data <- select(berghei_tpm, matches(experiment_name))
         if(ncol(tpm_data) > 4){ 
           #No point doing hclusts for any of the expts with under four samples
           dists = dist(t(tpm_data))
           par(cex=0.7)
           p <- ggdendrogram(hclust(dists)) +
             ggtitle(experiment_name)
           ggsave(filename = paste0('thesis_chapter3_figure_hclust_', 
                                    experiment_name, 
                                    '.png'),
                  plot = p,
                  path = 'plots')
           return(p)
         }
       }
)
```

Going to combine certain samples based on their similar profiles, e.g. <4h, >48h liver stages.

```{r combine_similar_samples_byNames}
col_stage <- colnames(berghei_tpm)

#Combine all Liver samples 4h or before into LiverPre4h
col_stage <- str_replace_all(col_stage,
                             pattern = 'Liv\\.(2|4)h_[[:digit:]]+',
                             replacement = 'Liv.pre4h')
#Combine all Liver samples 48h or onwards into LiverPost48h
col_stage <- str_replace_all(col_stage,
                             pattern = 'Liv\\.(4[8-9]|[5-9][0-9])h_[[:digit:]]+',
                             replacement = 'Liv.post48h')
#Combine all blood meal samples 3h or before into BldMlPre3h
col_stage <- str_replace_all(col_stage,
                             pattern = '(1h|3h)_[[:digit:]]+', 
                             replacement = 'Pre3h')
#Uniquify the name names
col_stage[grep(col_stage, 
               perl = TRUE,
               invert = TRUE,
               pattern = '_[[:digit:]]{1,}$')] <- uniquify(col_stage[grep(col_stage, 
                                                                          perl = TRUE,
                                                                          invert = TRUE,
                                                                          pattern = '_[[:digit:]]{1,}$')])

colnames(berghei_tpm) <- col_stage
```

```{r remove_badly_clustering_samples}
berghei_tpm$`E08_Liv.pre4h_1` <- NULL
berghei_tpm$`E03_Asex.Schiz_1` <- NULL
berghei_tpm$`E03_Asex.Schiz_3` <- NULL
```

## Saving the TPM

```{r save_tpm}
write.csv(berghei_tpm,
          'berghei_tpm.csv')
```