### 2022-07-21 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_calculateTPM.Rmd
Change more names to chab_tpm base and save pseudogene image

---
title: "Manuscript_Github_Pchabaudi_calculateTPM"
author: "Timothy Little"
date: "22/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      results = TRUE, 
                      message = FALSE, 
                      warning = FALSE)

knitr::opts_knit$set(root.dir = 
                       ".")
#Note that the above doesn't give an error if the dir doesn't exist

options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
```

```{r loading_packages, include = FALSE, message = FALSE}
# Load the required packages

library('readr')
library('dplyr')
library('ggplot2')
# library("pheatmap")
library("RColorBrewer")
library('data.table')
library('gridExtra')
library('tibble')
library('stringr')
library('kableExtra')
library('viridis')
# library('corrplot')
# library('reshape2')
# library('circlize') # for circular plots and colours
# library('cluster')    # clustering algorithms
# library('factoextra') # clustering algorithms & visualization
# library('ggfortify') # so ggplot understands prcomp objects
# library('ape') # for phylogeny plotting
# library('dendextend') # for dendrogram plotting
# library('readxl') # for reading in Excel spreadsheets
# library('plotly')
# library('scales')
# library('processx') # for downloading the plotly plots
# library('cowplot')
# library('ComplexHeatmap')
# library('ggpattern')

#Colour blind friendly palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
#show_col(cbPalette)

#Want the defaults for arrange, mutate and summarise to be from dplyr
arrange <- dplyr::arrange
mutate <- dplyr::mutate
summarise <- dplyr::summarise
select <- dplyr::select
```

```{r sourcing_functions}
source('../rfunctions/uniquify_fun.R')
source('../rfunctions/ordering_fun.R')
source('../rfunctions/tpm_fun.R')
source('../rfunctions/usefulFunctions_mround_my-max_firstup_fun.R')
```

```{r order_vec}

order_vector <- c("Early.Rings",
                  "Early.Mid.Rings",
                  "Late.Mid.Rings",
                  "Late.Rings",
                  "Rings.Troph.conversion",
                  "Early.Trophs",
                  "Mid.Trophs",
                  "Late.Trophs")
```

The _cir_ gene names were downloaded using a search of PlasmoDB _P. chabaudi_ genome. _cir_ genes are the names of _pir_ genes in _P chabaudi_.

```{r cir_id, message=FALSE, results='hide'}

# Getting the cir gene names and transcript lengths needed for calculating the transcript-per-million

chabaudi_gene_info <- read.csv('../standard_files/PlasmoDB-47_beta_Pchabaudichabaudi_GenesByTaxon_Summary.csv',
                              header = TRUE,
                              stringsAsFactors = FALSE)

chabaudi_transcript_lengths <- select(chabaudi_gene_info, c(Gene.ID, Transcript.Length)) %>% 
  arrange(Gene.ID) %>% 
  mutate(Geneid = Gene.ID) %>% 
  select(Geneid, Transcript.Length)

chabaudi_gene_info <- dplyr::rename(chabaudi_gene_info, 
                             Geneid = Gene.ID,
                             Description = Product.Description)

```

```{r cir_info}

cir_info <- read_delim('../standard_files/Pchabaudi_pir_info.txt',
           delim = '\t')
cir_id <- cir_info$Geneid

cir_clade_info <- read_delim('../standard_files/Pchabaudi_pir_clades.txt',
            delim = '\t')
```

```{r times_stages_labels}
#Use the rMT stage differential count in order to name the samples, but keep time to retain order
times_stages_table <- data.frame(time = c('02h', '05h', '08h', '11h', '14h', '17h','20h','23h'),
                                 stage_diff_count = c('Late-Rings',
                                                      'Rings-Troph-conversion',
                                                      'Early-Trophs',
                                                      'Mid-Trophs',
                                                      'Late-Trophs',
                                                      'Early-Rings',
                                                      'Early-Mid-Rings',
                                                      'Late-Mid-Rings'))
# R will convert the dashes to '.' at some point so let's have this notation here also
times_stages_table$r_stage_diff_count <- str_replace_all(times_stages_table$stage_diff_count, 
                                                         pattern = '-', 
                                                         replacement = '\\.')
```

```{r calc_tpm}
#Read in the count data and remove one sample which is unknown and had no description.
chab_count_table <- read.csv('data/tl19-07_twentyfour_counts_samples.csv') %>% 
  select(-rMT_Unknown)
#Calculate the TPM
chab_tpm.df <- tpm(count_data = chab_count_table, 
                          lengths = chabaudi_transcript_lengths)

#Do they all sum to 1e6? Note that all.equal allows for a tiny bit of difference which is okay.
all.equal(apply(chab_tpm.df[-1], 2, sum), 
          rep(1e6, times = dim(chab_tpm.df[-1])[2]), 
          check.names = FALSE)
```

```{r super-pseudogene}

chab_tpm_melt.df <- reshape2::melt(chab_tpm.df,
                                   variable.name = 'sample',
                                   value.name = 'tpm') %>% 
  mutate(time = str_extract(sample, 
                            pattern = '([[:digit:]]{2}h_.+|sch.culture)(?=_[[:digit:]])'),
         transmission = str_extract(sample,
                                    pattern = 'rMT|SBP'))

(plot <- ggplot(data = filter(chab_tpm_melt.df, 
                     Geneid %in% 'PCHAS_0420800'), 
       aes(x = time, 
           y = log1p(tpm),
           fill = transmission)) +
  stat_summary(fun = median, fun.min = median, fun.max = median,
               geom = "bar", width = 0.75, size = 0.25, position = 'dodge') +
  geom_point(stat = 'identity', position = position_dodge(width = 0.75)) +
  scale_fill_manual(values = cbPalette) +
  theme_classic() +
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5)))

ggsave(filename = 'plots/pseudogene_tpm_barchart.png',
       plot)

```

```{r tpm_rescale}

#Setting chab_tpm_ct core dataset - remove pseudogene and re-scaled

chab_tpm_nopseudo <- chab_tpm.df %>% 
  filter(!Geneid %in% 'PCHAS_0420800')

chab_tpm_rescale_nopseudo <- lapply(colnames(chab_tpm_nopseudo), 
                                          function(column_name){
                                            column <- getElement(chab_tpm_nopseudo, 
                                                                 paste(column_name))
                                            if(is.numeric(column)){
                                              #is.numeric to avoid trying to scale the Geneid column
                                              tpm_scale_factor <- 1e6 / sum(column)
                                              tpm_scaled <- as.numeric(column*tpm_scale_factor)
                                            }else{
                                              tpm_scaled <- column
                                            }
                                            return(tpm_scaled)
                                          }) %>% 
  bind_cols

colnames(chab_tpm_rescale_nopseudo) <- colnames(chab_tpm_nopseudo)
# apply(chab_tpm_rescale_nopseudo[,-1], 2, sum)

#Setting the re-scaled version as the default from now on
#Setting chab_tpm_ct core dataset - removed pseudogene and re-scaled
# chab_tpm.df <- chab_tpm_rescale_nopseudo
```

```{r clean_up_and_average}
chab_tpm_avg.df <- reshape2::melt(chab_tpm.df, 
                           variable.name = 'sample',
                           value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, pattern = '.+(?=_[[:digit:]])')) %>% 
  group_by(Geneid, stage) %>% 
  summarise(mean_tpm = mean(tpm)) %>% 
  reshape2::dcast(Geneid ~ stage)
```

```{r save}
write_csv(twentyfour_rMT,
          file = 'Pchabaudi_TPM_allBioReps.csv')

write_csv(twentyfour_rMT_avg,
          file = 'Pchabaudi_TPM_average.csv')
```