### 2022-07-21 chapter4_results_chabaudi24h/old_Pchabaudi24h/tl19-07_chabaudi24h_PCA_3D_rgl.Rmd

---
title: "tl19-07_chabaudi24h_PCA_3D_rgl"
author: "Timothy Little"
date: "12/01/2021"
output: 
  html_document:
    theme: cosmo
    code_download: true
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      results = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
knitr::opts_knit$set(root.dir = 
                       "/Users/littlet/Dropbox (The Francis Crick)/Data/tl19-07_audreybrugat24hchabaudi")
#Note that the above doesn't give an error if the dir doesn't exist
options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
```

```{r}
library('rgl')
library('magick')
library('dplyr')
library('readxl')
library('tibble')
library('factoextra')
library('stringr')
library('viridis')
```

```{r}
rmt <- read_xlsx('tl19-07_TPM_rMT_nopseudoRescaled_labelsStage_allSamples.xlsx') %>% 
  select(!matches('sch')) %>% 
  rename_with(~ paste0('rMT_',
                       .x),
              matches('[[:digit:]]{2}h'))
sbp <- read_xlsx('tl19-07_TPM_SBP_nopseudoRescaled_labelsStage_allSamples.xlsx') %>% 
  select(!matches('sch')) %>% 
  rename_with(~ paste0('SBP_',
                       .x),
              matches('[[:digit:]]{2}h'))
tpm_data <- merge(rmt, 
                  sbp) %>% 
  column_to_rownames(var = 'Geneid')
tpm_data <- tpm_data[,sort(colnames(tpm_data))]

# apply(tpm_data, 2, function(x) 1e10*(sum(x) - 1e6))
```

# 3D PCA to save

```{r 3d_PCA_save_rgl}

twentyfour_TPM_pca <- prcomp(tpm_data,
                             center = TRUE, 
                             scale. = TRUE)

twentyfour_TPM_pca_var <- get_pca_var(twentyfour_TPM_pca)
twentyfour_TPM_pca_coords <- data.frame(twentyfour_TPM_pca_var$coord)

pca_importance <- as.data.frame(summary(twentyfour_TPM_pca)$importance)
pc1_importance <- round(pca_importance$PC1[2]*100, digits = 1)
pc2_importance <- round(pca_importance$PC2[2]*100, digits = 1)
pc3_importance <- round(pca_importance$PC3[2]*100, digits = 1)

twentyfour_TPM_pca_coords <- data.frame(
  sample = rownames(twentyfour_TPM_pca_var$coord),
  PC1 = twentyfour_TPM_pca_coords$Dim.1, 
  PC3 = twentyfour_TPM_pca_coords$Dim.3, 
  PC2 = twentyfour_TPM_pca_coords$Dim.2
) %>% 
  dplyr::mutate(stage = gsub(sample,
                             pattern = "_[[:digit:]]$", 
                             replacement = "")) %>% 
  dplyr::mutate(time = str_extract(sample, 
                                   pattern = '[[:digit:]]{2}h')) %>% 
  dplyr::mutate(transmission = str_split_fixed(sample, 
                                               pattern = "_", 
                                               n = 3)[, 1])

colornames <- viridis(length(unique(twentyfour_TPM_pca_coords$time)))
colornames <- setNames(colornames, sort(unique(twentyfour_TPM_pca_coords$time)))

twentyfour_TPM_pca_coords$cols <- colornames[
  match(twentyfour_TPM_pca_coords$time,
        names(colornames))
]

twentyfour_TPM_pca_coords_rmt <- filter(twentyfour_TPM_pca_coords, 
                                        transmission == 'rMT')
twentyfour_TPM_pca_coords_sbp <-  filter(twentyfour_TPM_pca_coords, 
                                        transmission == 'SBP')

```

```{r movie3d_with_loop}
movie3d_loop <- function(f, duration, dev = cur3d(), ..., fps = 10, 
                         movie = "movie", frames = movie, dir = tempdir(), 
                         convert = NULL, clean = TRUE, verbose = TRUE,
                         top = TRUE, type = "gif", startTime = 0){
  olddir <- setwd(dir)
  on.exit(setwd(olddir))
  for (i in round(startTime * fps):(duration * fps)) {
    time <- i/fps
    if (cur3d() != dev) 
      set3d(dev)
    stopifnot(cur3d() != 0)
    args <- f(time, ...)
    subs <- args$subscene
    if (is.null(subs)) 
      subs <- currentSubscene3d(dev)
    else args$subscene <- NULL
    for (s in subs) par3d(args, subscene = s)
    filename <- sprintf("%s%03d.png", frames, i)
    if (verbose) {
      cat(gettextf("Writing '%s'\r", filename))
      flush.console()
    }
    rgl.snapshot(filename = filename, fmt = "png", top = top)
  }
  cat("\n")
  if (.Platform$OS.type == "windows") 
    system <- shell
  if (is.null(convert) && requireNamespace("magick")) {
    m <- NULL
    for (i in round(startTime * fps):(duration * fps)) {
      filename <- sprintf("%s%03d.png", frames, i)
      frame <- magick::image_read(filename)
      if (is.null(m)) 
        m <- frame
      else m <- c(m, frame)
      if (clean) 
        unlink(filename)
    }
    m <- magick::image_animate(m, fps = fps, loop = 0, dispose = "previous") 
    #Only change is here; 'loop = 1' -> 'loop = 0'
    magick::image_write(m, paste0(movie, ".", type))
    return(invisible(m))
  }
  else if (is.null(convert)) {
    warning("R package 'magick' is not installed; trying external package.")
    convert <- TRUE
  }
  if (is.logical(convert) && convert) {
    progname <- "magick"
    version <- try(system2(progname, "--version", stdout = TRUE, 
                           stderr = TRUE), silent = TRUE)
    if (inherits(version, "try-error") || !length(grep("ImageMagick", 
                                                       version))) {
      progname <- "convert"
      version <- try(system2(progname, "--version", stdout = TRUE, 
                             stderr = TRUE), silent = TRUE)
    }
    if (inherits(version, "try-error") || !length(grep("ImageMagick", 
                                                       version))) 
      stop("'ImageMagick' not found")
    filename <- paste0(movie, ".", type)
    if (verbose) 
      cat(gettextf("Will create: %s\n", file.path(dir, 
                                                  filename)))
    convert <- paste(progname, "-delay 1x%d %s*.png %s.%s")
  }
  if (is.character(convert)) {
    convert <- sprintf(convert, fps, frames, movie, type, 
                       duration, dir)
    if (verbose) {
      cat(gettextf("Executing: '%s'\n", convert))
      flush.console()
    }
    system(convert)
    if (clean) {
      if (verbose) 
        cat(gettext("Deleting frames\n"))
      for (i in 0:(duration * fps)) {
        filename <- sprintf("%s%03d.png", frames, i)
        unlink(filename)
      }
    }
  }
  invisible(convert)
}
```

```{r}
r3dDefaults$windowRect <- c(100, 100, 750, 750)
rgl.init()
plot3d(
  twentyfour_TPM_pca_coords_rmt$PC1,
  twentyfour_TPM_pca_coords_rmt$PC2,
  twentyfour_TPM_pca_coords_rmt$PC3,
  col = twentyfour_TPM_pca_coords_rmt$cols,
  type = "s", 
  radius = .025,
  alpha = 0.65,
  xlab = '',
  ylab = '',
  zlab = ''
)
cubes <- shapelist3d(
  cube3d(),
  twentyfour_TPM_pca_coords_sbp$PC1,
  twentyfour_TPM_pca_coords_sbp$PC2,
  twentyfour_TPM_pca_coords_sbp$PC3,
  size = 0.017,
  alpha = 0.65,
  # ambient = 'black',
  emission = twentyfour_TPM_pca_coords_sbp$cols,
  col = twentyfour_TPM_pca_coords_sbp$cols,
  # shininess = 25,
  # specular = 'black'
  lit = TRUE
)
#Add outlines (copy pasted from 
# https://stackoverflow.com/questions/39730889/rgl-drawing-a-cube-with-colored-faces-vertex-points-and-lines )
for (i in 1:length(cubes)){
  c3d <- cubes[[i]]
  for (i in 1:6){
    lines3d(t(c3d$vb)[c3d$ib[,i],])
  }
}
# add titles
title3d('', 
        '', 
        paste0('PC1 ', pc1_importance, '%'), 
        paste0('PC2 ', pc2_importance, '%'), 
        paste0('PC3 ', pc3_importance, '%'))

empty_length <- floor(length(unique(twentyfour_TPM_pca_coords$time))/2)
legend3d("topright",
         legend = c(
           unique(twentyfour_TPM_pca_coords$time),
           rep_len('', 
                   length.out = empty_length),
           'RMT',
           'SBP',
           rep_len('', 
                   length.out = empty_length)
         ), 
         pch = c(
           rep_len(16, 
                   length.out = empty_length*2),
           rep_len(NA, 
                   length.out = empty_length) ,
           16,
           15,
           rep_len(NA, 
                   length.out = empty_length)
           ), 
         col = c(
           unique(twentyfour_TPM_pca_coords$cols),
           rep_len(NA, 
                   length.out = empty_length),
           'blue',
           'blue',
           rep_len(NA, 
                   length.out = empty_length)
         ), 
         cex=1, 
         inset=c(0.02),
         ncol = 2)
aspect3d("iso")

# spin3d rpm defaults at 5 rpm - so 60/5 = 12s carries one full rotation.
movie3d_loop(spin3d(axis = c(0,0,1)),
        duration = 12,
        movie = 'tl19-07_pca3d_test',
        type = 'gif',
        dir = getwd())

# pca3dplot <- scene3d()
```

This works alright at the moment, although the additional shape I add for the 'SBP' samples seems stretched, probably because of the different lengths of the xyz axes. "iso" setting in aspect3d fixes this, so it is indeed to do with that.