### 2022-07-21 chapter4_results_chabaudi24h/old_Pchabaudi24h/tl19-07_chabaudi24h_pir_GeneSetEnrichment.Rmd

---
title: "tl19-07_chabaudi24h_pir_GeneSetEnrichment"
author: "Timothy Little"
date: "21/01/2021"
output: 
  html_document:
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = TRUE, message = FALSE, warning = FALSE)

# Set the working directory for the document where the data is located.
knitr::opts_knit$set(root.dir = 
                       "/Users/littlet/Dropbox (The Francis Crick)/Data/tl19-07_audreybrugat_24h_chabaudi")
# Change options so that numbers are displayed with all the zeros (e,g, as 1000000 not 1e6)
options(scipen = 999) 
```

```{r load_packages}
library('dplyr')
library('readr')
library('readxl')
library('data.table')
library('tibble')
library('qusage')
library('ggplot2')
library('stringr')
library('RColorBrewer')
```

```{r read_count_data}
# chab_counts <- read_xlsx('Manuscript_Frontiers_chabaudi_24h_counts.xlsx') %>% 
#   filter(!Geneid %in% 'PCHAS_0420800') #Removing the meddlesome pseudogene

chab_tpm <- read_xlsx('Manuscript_Frontiers_SuppInfo_chabaudi_24h.xlsx',
          sheet = 'chab_TPM_bioreps')

# chab_tpm %>% filter(Geneid %in% 'PCHAS_0420800')

#Filter low expressed genes (greater than 10 tpm in at least one sample), convert to a matrix, add 1 for the log transformation, then transform to log2
chab_tpm <- column_to_rownames(chab_tpm, var = 'Geneid')

chab_tpm <- chab_tpm[apply(chab_tpm, 1, function(x) {any(x >= 10)}),]

chab_tpm_log2plus1.mat <- chab_tpm %>% 
  as.matrix %>% 
  `+`(.+1) %>% 
  log(base = 2)

chab_tpm_log2plus1.mat[1:5,1:5]
```

```{r generate_labels}
#Use make.names to remove the '-' from the labels
colnames(chab_tpm_log2plus1.mat) <- make.names(colnames(chab_tpm_log2plus1.mat))

labels <- colnames(chab_tpm_log2plus1.mat)

#We want the life cycle stages to be the different groups for now.

lifeCycle_labels <- str_extract(labels, 
                                pattern = '[[:digit:]]{2}h')
unique(lifeCycle_labels) %>% sort

transmission_labels <- str_extract(labels, 
                                pattern = 'rMT|SBP')
unique(transmission_labels) %>% sort

```

```{r generate_genesets}

pir_info <- read_xlsx('Manuscript_Frontiers_SuppInfo_chabaudi_24h.xlsx',
                      sheet = 1)

pir_geneset <- list(pirs = pir_info$Geneid)

chrom_list <- lapply(unique(pir_info$chrom), 
                     function(chrom_nam) filter(pir_info, 
                                                chrom == chrom_nam)$Geneid)

names(chrom_list) <- unique(pir_info$chrom)

#Remove groups with only three members or fewer.
chrom_list <- chrom_list[unlist(lapply(chrom_list, function(x) length(x) > 3))]

subfam_list <- lapply(unique(pir_info$subfam), 
                     function(clade_nam) filter(pir_info, 
                                                subfam == clade_nam)$Geneid)
names(subfam_list) <- unique(pir_info$subfam)
subfam_list <- subfam_list[unlist(lapply(subfam_list, function(x) length(x) > 3))]

ShortLong_list <- 
  lapply(unique(pir_info$SorL), 
                     function(SorL_nam) filter(pir_info, 
                                                SorL == SorL_nam)$Geneid)
names(ShortLong_list) <- unique(pir_info$SorL)
ShortLong_list <- ShortLong_list[unlist(lapply(ShortLong_list, function(x) length(x) > 3))]

```

```{r running_qusage_pirs}
#For each life cycle stage:
qs.results.list <- lapply(
  unique(transmission_labels), 
  function(transmission){

    index_trans <- str_detect(colnames(chab_tpm_log2plus1.mat), 
                              pattern = transmission)

    #Perform qusage
    qs.results <- qusage(chab_tpm_log2plus1.mat[,index_trans],
                         lifeCycle_labels[index_trans],
                         '14h-02h',
                         pir_geneset,
                         n.points = 2^18)
  }
)

names(qs.results.list) <- unique(transmission_labels)

qsTable(qs.results.list$SBP)
qsTable(qs.results.list$rMT)
```

```{r}
 qs.results.trans.chrom <- qusage(chab_tpm_log2plus1.mat,
                              transmission_labels,
                              'rMT-SBP',
                              chrom_list,
                              n.points = 2^18)
```

```{r}
qsTable(qs.results.trans.chrom)
```

```{r}
 qs.results.trans.pirs <- qusage(chab_tpm_log2plus1.mat,
                              transmission_labels,
                              'rMT-SBP',
                              pir_geneset,
                              n.points = 2^18)
```

```{r}
qsTable(qs.results.trans.pirs)
```

WTF

```{r}

index_14h <- str_detect(colnames(chab_tpm_log2plus1.mat), pattern = '14h')

qs.results.trans.pirs.14h <- qusage(chab_tpm_log2plus1.mat[,index_14h],
                                transmission_labels[index_14h],
                                'rMT-SBP',
                                pir_geneset,
                                n.points = 2^18)

qsTable(qs.results.trans.pirs.14h)

plotGeneSetDistributions(qs.results.trans.pirs.14h)
```

```{r qusage_pirs-expressed-higher-RMT}

rmt_14h_hi <- read_xlsx('Manuscript_Frontiers_SuppInfo_chabaudi_24h.xlsx',
                      sheet = 'chab_TPM_bioreps') %>% 
  filter(Geneid %in% cir_id) %>%
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(
    stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
    transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1],
    diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3],
    time = str_split_fixed(sample, pattern = '_', n = 3)[,2],
    subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)]
  ) %>% 
  filter(time == '14h') %>% 
  group_by(transmission, time, Geneid) %>% 
  summarise(tpm_mean = mean(tpm)) %>% 
  dcast(Geneid ~ transmission, 
        margins = c('tpm_mean')) %>% 
  filter((rMT - SBP) > 150) %>% 
  .$Geneid

qs.results.trans.pirs.14h.rmthi <- qusage(chab_tpm_log2plus1.mat,
                                          transmission_labels,
                                          'rMT-SBP',
                                          list(rmt_hi = rmt_14h_hi),
                                          n.points = 2^18)

qsTable(qs.results.trans.pirs.14h.rmthi)

```