### 2022-07-21 chapter4_results_chabaudi24h/old_Pchabaudi24h/tl19-07_manuscript_figures.Rmd

---
title: "tl19-07_manuscript_figures"
author: "Timothy Little"
date: "06/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      results = TRUE, 
                      message = FALSE, 
                      warning = FALSE)

#for server
# knitr::opts_knit$set(root.dir = '~/data/tl19-07_chabaudi_24h_audreyMTbrugatSBP')
knitr::opts_knit$set(root.dir = 
                       "/Users/littlet/Dropbox (The Francis Crick)/Data/tl19-07_audreybrugat24hchabaudi")
#Note that the above doesn't give an error if the dir doesn't exist

options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
```

```{r loading_packages, include = FALSE, message = FALSE}
# Load the required packages

# library('tidyverse')
library('readr')
library('dplyr')
library('tidyr')
library('ggplot2')
library('DESeq2')
library("pheatmap")
library("RColorBrewer")
library('data.table')
library('gridExtra')
# library('apeglm')
# library('ashr')
library('tibble')
library('stringr')
# library('cowplot')
library('kableExtra')
library('viridis')
# library('VennDiagram')
library('corrplot')
library('reshape2')
library('circlize') # for circular plots
library('cluster')    # clustering algorithms
library('factoextra') # clustering algorithms & visualization
library('ggfortify') # so ggplot understands prcomp objects
library('ape') # for phylogeny plotting
library('dendextend') # for dendrogram plotting
library('readxl') # for reading in Excel spreadsheets
library('plotly')
library('scales')
library('processx') # for downloading the plotly plots
library('cowplot')
library('writexl')

library('devtools')
#These packages were installed using github
library('ComplexHeatmap')
library('ggbiplot')
library('ggpattern')

#Colour blind friendly palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
show_col(cbPalette)

#Want the defaults for arrange, mutate and summarise to be from dplyr
arrange <- dplyr::arrange
mutate <- dplyr::mutate
summarise <- dplyr::summarise
select <- dplyr::select
```

```{r function_mround}
mround <- function(x,base){ 
  base*ceiling(x/base) 
} 
```

```{r function_my_max}
my_max <- function(x) ifelse( !all(is.na(x)), max(x, na.rm=T), NA) #max robust to NAs
```

```{r function_ordering}
ordering <- function(to_order, order_vector){
  order = unlist(sapply(order_vector, 
                        function(x){unique(str_subset(to_order, 
                                                      pattern = paste(x)))}))
  order_missing = unique(to_order[!to_order %in% order])
  unique(c(order, order_missing), fromLast = TRUE)
}

order_vector <- c("Early.Rings",
                  "Early.Mid.Rings",
                  "Late.Mid.Rings",
                  "Late.Rings",
                  "Rings.Troph.conversion",
                  "Early.Trophs",
                  "Mid.Trophs",
                  "Late.Trophs")
```

The _cir_ gene names were downloaded using a search of PlasmoDB _P. chabaudi_ genome.

```{r cir_id, message=FALSE, results='hide'}

# Getting the cir gene names and transcript lengths needed for calculating the transcript-per-million

chabaudi_gene_info = read.csv('chabaudi_PccAS_PlasmoDBv44_geneinformation_GenesByTaxon_Summary.csv',
                              header = TRUE,
                              stringsAsFactors = FALSE)
#.txt file is weirdly truncated, don't use it

chabaudi_transcript_lengths = select(chabaudi_gene_info, c(Gene.ID, Transcript.Length)) %>% 
  arrange(Gene.ID)

cir_info = chabaudi_gene_info %>% 
  select(Gene.ID, Product.Description, Transcript.Length) %>% 
  filter(Product.Description == 'PIR protein')
nrow(cir_info)

#Brugat 2017 cirs
list.files()
brugat_cir_names = read_xlsx('chabaudi_pir_cir_genes_brugat_2017_nmicrobiol2016276-s4.xlsx') %>% 
  .$id
length(brugat_cir_names)
#Two extra genes compared to from PlasmoDB

#Adam Reid Personal Comm 2019 cirs
reid_cir_clade = read.table('AdamReid_chabaudi_PccAS_v3_aa.pir.clades.simple.txt',
                            header = FALSE,
                            stringsAsFactors = FALSE,
                            sep = '\t',
                            col.names = c('Geneid','clade'))
reid_cir_names = reid_cir_clade$Geneid
length(reid_cir_names)
#An extra gene compared to Brugat

brugat_cir_names[!brugat_cir_names %in% cir_info$Gene.ID] #Not annotated as cirs by PlasmoDB v Brugat
cir_info$Gene.ID[!cir_info$Gene.ID %in% brugat_cir_names] #Not annotated as cirs by Brugat 2017? v PlasmoDB
cir_info$Gene.ID[!cir_info$Gene.ID %in% reid_cir_names] #Not annotated as cirs by Reid v PlasmoDB
reid_cir_names[!reid_cir_names %in% brugat_cir_names] #Not annotated as cirs by Brugat v Reid
reid_cir_names[!reid_cir_names %in% cir_info$Gene.ID] #Not annotated as cirs by PlasmoDB v Reid

#Adam Reid's list contains all the pirs we should assume, however I don't understand why it is missing the 'conserved rodent pir' - need to discuss this with him.
#After extensive discussion it has been decided that due to dissimilarity and lack of an InterPro pir domain the
# pan-rodent is not a pir gene

reid_missing_cir = cir_info$Gene.ID[!cir_info$Gene.ID %in% reid_cir_names]

#Add the missing cir to the list of cirs and their clades
cir_clade_info <- reid_cir_clade %>% 
  mutate(SorL = str_extract(clade, 
                            pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'))
write_delim(cir_clade_info,
            '../standard_files/Pchabaudi_pir_clades.txt',
            delim = '\t')

cir_info <- chabaudi_gene_info %>% 
  select(Gene.ID, 
         Product.Description, 
         Transcript.Length) %>% 
  filter(Gene.ID %in% c(reid_cir_names)) %>% 
  mutate(
    Geneid = Gene.ID,
    chrom = str_extract(Gene.ID, pattern = '(?<=PCHAS_)[[:digit:]]{2}'),
    subfam = reid_cir_clade$clade[match(Geneid, reid_cir_clade$Geneid)]
  ) %>% 
  mutate(SorL = str_extract(subfam, pattern = 'L|S|ancestral|unk'))
cir_transcript_lengths = select(cir_info, -Product.Description, -Gene.ID)

cir_id <- cir_info$Gene.ID
length(cir_id)

cir_id_no_anc_or_rodcon <- cir_id[!cir_id %in% 'PCHAS_0101200']

#How many cirs are present on each chromosome?

cir_tbl = as_tibble(data.frame(Geneid = cir_id))
cir_chrom_tbl <- cir_tbl %>% 
  mutate(chrom = unlist(sapply(lapply(sapply(as.character(cir_tbl$Geneid), strsplit,
                                             split = ''), '[', 7:8), 
                               paste, collapse = ''))) 
cir_chrom_tbl %>% 
  group_by(chrom) %>% 
  dplyr::summarise(number_of_cirs_in_each_chromosome = n(),
                   percentage_of_all_cirs = round(n()/nrow(cir_tbl), 3)*100) %>% 
  dplyr::arrange(desc(percentage_of_all_cirs))

pir_clade <- reid_cir_clade %>% 
  mutate(SorL = str_extract(clade, pattern = 'S|L|anc|unk'))

#How many in each clade?
cir_clade_tbl = reid_cir_clade %>% 
  as_tibble() %>% 
  group_by(clade) %>% 
  dplyr::summarise(n())
cir_clade_tbl

chabaudi_gene_info <- dplyr::rename(chabaudi_gene_info, 
                             Geneid = Gene.ID,
                             Description = Product.Description)
```

```{r chapl_info}
cir_loci <- read_xlsx('Brugat2017_pirexpressiondata_41564_2017_BFnmicrobiol2016276_MOESM25_ESM.xlsx',
                            sheet = 1) %>% 
  filter(id %in% cir_id)

cir_chapl_aapl_info <- cir_loci %>%
  mutate(
    Geneid = id,
    locus_type = `Locus type`,
    chapl_aapl = ifelse(is.na(`Locus type`),
                        NA,
                        paste0(Subtelomere, 
                               '_',
                               `Locus type`)
    )
  ) %>% 
  select(Geneid, locus_type, chapl_aapl)

cir_info <- mutate(cir_info, 
                   subtelomere = cir_loci$Subtelomere[match(Geneid, cir_loci$id)],
                   chapl_aapl = cir_loci$`Locus type`[match(Geneid, cir_loci$id)],
                   chapl_aapl_name = cir_chapl_aapl_info$chapl_aapl[match(Geneid, cir_chapl_aapl_info$Geneid)])

```

```{r writing_cir_info_file}

write_delim(cir_info,
            path = 'tl19-07_dataset_pirs_chabaudi_info.txt',
            delim = '\t')

```

```{r times_stages_labels}
#Use the rMT stage differential count in order to name the samples, but keep time to retain order
times_stages_table <- data.frame(time = c('02h', '05h', '08h', '11h', '14h', '17h','20h','23h'),
                                 stage_diff_count = c('Late-Rings',
                                                      'Rings-Troph-conversion',
                                                      'Early-Trophs',
                                                      'Mid-Trophs',
                                                      'Late-Trophs',
                                                      'Early-Rings',
                                                      'Early-Mid-Rings',
                                                      'Late-Mid-Rings'))
# R will convert the dashes to '.' at some point so let's have this notation here also
times_stages_table$r_stage_diff_count <- str_replace_all(times_stages_table$stage_diff_count, 
                                                         pattern = '-', 
                                                         replacement = '\\.')
```

```{r read_tpm_rescaled_data}
#Setting twentyfour_tpm_ct core dataset - removed pseudogene and re-scaled

twentyfour_ct_comb <- list( tpm = read_xlsx('tl19-07_TPM_both_nopseudoRescaled_labelsStage_allSamples.xlsx') )

twentyfour_ct_comb_noschz <- list( tpm = read_xlsx('tl19-07_TPM_both_nopseudoRescaled_labelsStage_allSamples.xlsx') %>% 
                                    dplyr::select(!matches('sch'))  )

twentyfour_avgtpm <- list(
  'rMT' = list(
    tpm = read_xlsx(
                    'tl19-07_TPM_rMT_nopseudoRescaled_labelsStage_meanTPM.xlsx'),
    experiment = 'rMT'
  ),
  'SBP' = list(
    tpm = read_xlsx(
                     'tl19-07_TPM_SBP_nopseudoRescaled_labelsStage_meanTPM.xlsx'),
    experiment = 'SBP'
  )
)

twentyfour_avgtpm_noschz <- lapply(twentyfour_avgtpm, 
                                   function(experiment_data){
                                     tpm_df <- experiment_data$tpm
                                     tpm_df <- tpm_df[!str_detect(colnames(tpm_df), pattern = 'sch')]
                                     # Remove the column with 'sch' pattern in it = schizonts
                                     list(tpm = tpm_df,
                                          experiment = experiment_data$experiment)
                                   })

```

```{r colours_clades_phase}
SorL_col <- setNames(
  c(    '#F0E442',
        '#56B4E9', #L
        '#E69F00', #S
        '#999999'),
  sort(
    unique(cir_clade_info$SorL)
  )
)

SorL_col_num <- SorL_col
names(SorL_col_num) <- sapply(names(SorL_col), 
                              function(clade_name) {
                                paste0(
                                  clade_name, 
                                  ' (n = ',
                                  sum(pir_clade$SorL == clade_name),
                                  ')'
                                )
                              })

clade_col <- setNames(
  c('#F0E442',
    '#56B4E9', #L1
    '#009E73',
    '#0072B2',
    '#E69F00', #S1
    '#D55E00',
    '#CC79A7',
    '#999999'),
  sort(
    unique(cir_clade_info$clade)
  )
)

clade_col_num <- clade_col
names(clade_col_num) <- sapply(names(clade_col), 
                               function(clade_name) {
                                 paste0(
                                   clade_name, 
                                   ' (n = ',
                                   sum(pir_clade$clade == clade_name),
                                   ')'
                                 )
                               })

#Be careful to ensure that the above is consistent with other images in the same paper.

```

# Figure  - Heatmap of top S and L gene expression.

```{r figure_heatmap_table}

#Get the TPM data for the pirs family being looked at
highest_SorL_24h <- lapply(twentyfour_avgtpm_noschz, 
                          function(tpm_input){

                            tpm_data_SorL_stackgenes_top_table <- tpm_input$tpm %>% 
                              filter(Geneid %in% cir_id) %>% 
                              melt(id.vars = 'Geneid', 
                                   variable.name = 'stage', 
                                   value.name = 'total_tpm') %>% 
                              mutate(total_tpm = as.numeric(total_tpm),
                                     SorL = cir_clade_info$SorL[match(Geneid, cir_clade_info$Geneid)],
                                     stage = times_stages_table$time[match(stage, times_stages_table$r_stage_diff_count)]) %>% 
                              filter(str_detect(SorL, 
                                                pattern = 'ancestral|unk', 
                                                negate = TRUE)) %>% 
                              group_by(stage) %>% 
                              mutate(sum_tpm = sum(total_tpm)) %>% 
                              group_by(stage, 
                                       SorL) %>% 
                              slice_max(order_by = total_tpm,
                                        n = 1) %>% 
                              mutate(total_tpm = round(total_tpm, digits = 2),
                                     prop_tpm = round(total_tpm/sum_tpm, digits = 2)) %>% 
                              select(stage, Geneid, SorL, total_tpm, prop_tpm)

                            # write.csv(tpm_data_SorL_stackgenes_top_table,
                            #           file = paste0('tl19-07_combined_SorL_topperstage_table_', 
                            #                         tpm_input$experiment, 
                            #                         '.csv'))

                            # write_xlsx(tpm_data_SorL_stackgenes_top_table,
                            #            paste0('tl19-07_combined_SorL_topperstage_table_', 
                            #                   tpm_input$experiment, 
                            #                   '.xlsx'))

                            SorL_kable_index <- rep.int(2, 
                                                        times = length(unique(tpm_data_SorL_stackgenes_top_table$stage)))
                            names(SorL_kable_index) <- unique(tpm_data_SorL_stackgenes_top_table$stage)

                            return(
                              list(dataframe = tpm_data_SorL_stackgenes_top_table, 
                                   table = 
                                     tpm_data_SorL_stackgenes_top_table %>% 
                                     kable(caption = paste(tpm_input$experiment) )%>% 
                                     kable_styling() %>% 
                                     pack_rows(index = SorL_kable_index) %>% 
                                     scroll_box(width = "100%", height = "500px"),
                                   experiment = tpm_input$experiment
                              )
                            )
                          }
)

lapply(highest_SorL_24h, function(expt) expt$table)

```

```{r figure_heatmap}

heatmap_row_name_fonts <- 20
heatmap_col_name_fonts <- 15
heatmap_col_title_fonts <- 15 
threshold <- 0

highest_SorL_pirs_bystage <- sort(
  unique(
    unlist(
      lapply(
        highest_SorL_24h, function(expt) expt$dataframe$Geneid)
    )))

highest_SorL_pirs_bystage_subfams <- data.frame(
  Geneid = str_replace(as.character(highest_SorL_pirs_bystage), 
                                    pattern = 'PCHAS_', 
                                    replacement = ''),
                       subfamily = cir_clade_info$SorL[match(highest_SorL_pirs_bystage
                                                             , cir_clade_info$Geneid)]
  )

ha = Heatmap(
  as.matrix(highest_SorL_pirs_bystage_subfams$subfamily),
  name = 'Subfamily',
  na_col = 'white',
  col = SorL_col,
  cluster_rows = FALSE, 
  cluster_columns = FALSE,
  column_names_gp = gpar(col = 'darkslategray', 
                         fontsize = 12, 
                         fontface = 'italic'),
  show_column_dend = FALSE,
  rect_gp = gpar(col = 'white'), 
  width = unit(1, "cm"),
  heatmap_legend_param = list(
    labels_gp = gpar(fontsize = 10),
    title_gp = gpar(fontsize = 10)
  ),
  show_heatmap_legend = FALSE
  # Going to hide the legend so I can make it manually later with the grid.points that I use in the heatmap body.
)

#Get the TPM data for the pirs family being looked at
expt_tpms_top_pirs <- lapply(twentyfour_avgtpm_noschz, 
                             function(tpm_input){
                               # print(paste0('Now doing ', tpm_input$experiment))
                               tpms <- tpm_input$tpm %>% 
                                 filter(Geneid %in% highest_SorL_pirs_bystage) %>% 
                                 as.data.frame

                               colnames(tpms) <- ifelse(
                                 !colnames(tpms) %in% 'Geneid',
                                                        times_stages_table$time[match(colnames(tpms),
                                                                                      times_stages_table$r_stage_diff_count)],
                                                        'Geneid'
                               )

                               tpms <- tpms[,
                                            order(colnames(tpms)),
                                            drop = FALSE]
                               #Ordering using the ordering function and order_vector defined in the functions section.

                               #Data for the grid.dots for heatmap to show max for each stage
                               max_SorL <- mutate(tpms, 
                                                  SorL = cir_clade_info$SorL[match(Geneid, 
                                                                                   cir_clade_info$Geneid)]) %>% 
                                 group_by(SorL) %>% 
                                 mutate(across(-matches('Geneid|SorL'), 
                                               ~ ifelse(.x == max(.x), 
                                                        SorL, 
                                                        NA))) %>% 
                                 ungroup() %>% 
                                 select(-Geneid, -SorL)

                               #Remove PCHAS_ from the gene names for simplicity
                               rownames(tpms) <- str_replace(as.character(tpms$Geneid), 
                                                             pattern = 'PCHAS_', 
                                                             replacement = '')
                               tpms <- tpms[str_sort(rownames(tpms), 
                                                     numeric = TRUE),]
                               tpms <- select(tpms, -Geneid)

                               tpms <- apply(tpms, 2, function(x){ifelse(x >= threshold, 
                                                                         x, 
                                                                         NA)}) #Get NA for values under threshold
                               tpms <- log2(1+tpms)
                               if(tpm_input$experiment %>%  is.null){
                                 tpm_input$experiment = ''
                               }
                               list = list(tpm = tpms, 
                                           experiment = tpm_input$experiment,
                                           max_SorL = max_SorL)
                               return(list)
                             })

    maxcol <- max(
      unlist(
        lapply(
          expt_tpms_top_pirs, 
          function(x) max(
            x$tpm,
            na.rm = TRUE
          )
        )
      )
    )
    #Get the max TPM value

heatmaps = lapply(expt_tpms_top_pirs, 
                  function(tpms){

                    if(tpms$experiment != dplyr::last(expt_tpms_top_pirs)$experiment){
                      #If it is the final element don't remove the row names
                      rownames(tpms$tpm) = c()
                    }
                    #Setting the name if one is not provided
                    title_name = tpms$experiment
                    #Drawing each heatmap
                    ht = Heatmap(tpms$tpm, 
                                 name = 'log(TPM + 1)', 
                                 col = colorRamp2(c(threshold, 
                                                    maxcol/4, 
                                                    maxcol/2, 
                                                    3*maxcol/4, 
                                                    maxcol), 
                                                  viridis(5)),
                                 cluster_rows = FALSE, 
                                 cluster_columns = FALSE,
                                 column_title = paste(title_name),
                                 column_title_gp = gpar(fontsize = heatmap_col_title_fonts),
                                 row_names_gp = gpar(fontsize = heatmap_row_name_fonts),
                                 column_title_rot = 0,
                                 column_names_rot = 0,
                                 column_names_centered = TRUE,
                                 column_names_gp = gpar(fontsize = heatmap_col_name_fonts),
                                 na_col = 'white', 
                                 border = TRUE,
                                 heatmap_legend_param = list(
                                   title = 'log(TPM + 1)', 
                                   at = mround(
                                     c(log2(1+threshold), 
                                       seq(0, 
                                           maxcol, 
                                           length.out = 4)[-1]
                                     ), 
                                     base = 0.05),
                                   labels_gp = gpar(fontsize = 10),
                                   title_gp = gpar(fontsize = 10, fontface = "bold")
                                 ),
                                 cell_fun = function(j, i, x, y, width, height, fill) {
                                   if(!is.na(as.matrix(tpms$max_SorL)[i, j])){
                                     ifelse(
                                       as.matrix(tpms$max_SorL)[i, j] == 'S',
                                       grid.points(x, y, pch = 16, size = unit(3, "mm")),
                                       grid.points(x, y, pch = 18, size = unit(4, "mm"))
                                     )
                                   }
                                 }
                    )
                  })

ldg_subfam <- Legend(labels = c('S', 'L'),
                     type = 'points',
                     pch = c(16,18),
                     size = unit.c(unit(2, "mm"), 
                                   unit(3, "mm")),
                     background = SorL_col[c('S','L')],
                     title = 'Subfamily',
                     title_gp = gpar(fontsize = 10, fontface = "bold"))

ht_list = Reduce('+', heatmaps) #To add the list elements together
ht_list <- ha + ht_list
# pdf('tl19-07_pir_SorLhighest_figure2c_heatmap.pdf'
#     , width = 10
#     , height = 4
# )
draw(ht_list,
     row_title_side = 'left',
     column_title = 'Malaria stage',
     column_title_side = 'bottom',
     column_title_gp = gpar(fontsize = 16),
     heatmap_legend_side = 'left',
     ht_gap = unit(0.5, "cm"),
     heatmap_legend_list = list(ldg_subfam))
# dev.off()
```

## Figure 4 - Expression of major expressed cirs rMT/SBP side-by-side.

_pir_ genes selected by being expressed above 1 TPM in all samples.

Adding in housekeeping genes for comparison:
HSP70 putative  PCHAS_0721000
MSP1 PCHAS_0831300
AMA1 PCHAS_0931000

```{r pirs_side-by-side_figure4}

housekeeping_genes <- data.frame('HSP70' = 'PCHAS_0721000',
                             'MSP1' = 'PCHAS_0831300',
                             'AMA1' = 'PCHAS_0931000')

cir_housekeeping_id <- c(cir_id, 
                         'PCHAS_0721000',
                         'PCHAS_0831300',
                         'PCHAS_0931000')

cir_plot <- twentyfour_ct_comb$tpm %>% 
  filter(Geneid %in% cir_housekeeping_id) %>%
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(
    stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
    transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1],
    diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3],
    time = str_split_fixed(sample, pattern = '_', n = 3)[,2],
    subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)]
  ) %>% 
  mutate(subfam = ifelse(is.na(subfam), 'ref', subfam)) %>% 
  filter(!(time %in% c('sch.culture', 
                       'Unknown'))) %>% 
  mutate(subfam = str_replace(subfam, 
                              'ancestral',
                              'a'))

cir_plot$Geneid <- ifelse(
  cir_plot$Geneid %in% housekeeping_genes[1,],
  colnames(housekeeping_genes)[ match(cir_plot$Geneid, housekeeping_genes[1,]) ],
  cir_plot$Geneid
)

limits_cir <- c(0, 
                mround(max(cir_plot$tpm), 
                       base = 100))

cir_plot_order <- cir_plot %>% 
  filter(time == '14h',
         transmission == 'rMT') %>% 
  arrange(desc(tpm))

cir_plot <- mutate(cir_plot, 
                   Geneid = factor(Geneid, 
                                   levels = unique(cir_plot_order$Geneid)))

side_by_side_cir_plot <- function(time_vec = '14h',
                                  log_it = TRUE,
                                  save = TRUE,
                                  limits = limits_cir,
                                  point_size = 1,
                                  text_size = 10,
                                  file_name_end = '',
                                  threshold_min = 100,
                                  threshold_max = Inf){

  cir_plot <- group_by(cir_plot, 
                       Geneid) %>% 
    filter(
      time == time_vec,
      !Geneid %in% cir_id | any(tpm >= threshold_min),
      !Geneid %in% cir_id | all(tpm <= threshold_max)
    )

  # cir_plot %>% group_by(Geneid) %>% summarise(n())
  if(log_it){
    cir_plot$tpm <- log2(1+cir_plot$tpm)
      limits = log2(1+limits)
      breaks <- round(seq.int(0, 
                              max(limits), 
                              length.out = round(max(limits)) + 1, 
                              digits = 0), 
                      digits = 0)
    }else{
      breaks <- seq.int(0, max(limits), length.out = 11)
    }

    cir_plot$Geneid <- factor(
      str_remove(cir_plot$Geneid, 
                 pattern = 'PCHAS_'),
      levels = str_remove(levels(cir_plot$Geneid), 
                          pattern = 'PCHAS_')
    )

    cir_plot$subfam <- factor(
      cir_plot$subfam,
      levels = c(
        sort(unique(cir_plot$subfam)[!(unique(cir_plot$subfam) %in% 'ref')]),
        'ref'
      )
    )

    y <- ggplot(
      data = cir_plot,
      mapping = aes(
        x = Geneid,
        fill = transmission,
        y = tpm
      )
    ) +
      stat_summary(fun.y = median, 
                   fun.ymin = median, 
                   fun.ymax = median,
                   geom = "bar", 
                   width = 0.75, 
                   size = 0.25, 
                   position = 'dodge') +
      geom_point(stat = 'identity', 
                 position = position_dodge(width = 0.75),
                 size = point_size) +
      scale_y_continuous(breaks = breaks,
                         limits = limits) +
      scale_fill_manual(values = c('skyblue','red')) +
      facet_grid(.~subfam, 
                 scales = "free_x",
                 space = "free_x") +
      theme_classic() +
      theme(
        axis.text.x = element_text(size = text_size,
                                   angle = 90)
      ) +
      ylab('log(TPM + 1)')+
      ggtitle(paste(unique(cir_plot$time)))

    #This bit is to change the colour, and text colour, of the facet boxes

    g <- ggplot_gtable(ggplot_build(y))
    strip_t <- which(grepl('strip-t', g$layout$name))
    subfam_vec <- sort(unique(cir_plot$subfam))
    fills <- case_when(str_detect(subfam_vec, 
                                  pattern = 'a') ~ clade_col['ancestral'],
                       subfam_vec %in% names(clade_col) ~ clade_col[paste(subfam_vec)],
                       str_detect(subfam_vec, 
                                  pattern = 'ref') ~ 'white')
    k <- 1
    for (i in strip_t) {
      j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
      g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
      # if(getGrob(g$grobs[[i]]$grobs[[1]], 'GRID', grep = TRUE)$label == 'a'){
      #   #If grob is for the 'a' label we want to change the font
      #   m <- which(grepl('text', g$grobs[[i]]$grobs[[1]]$childrenOrder))
      #   n <- which(grepl('GRID.text', g$grobs[[i]]$grobs[[1]]$children[[m]]$childrenOrder))
      #   g$grobs[[i]]$grobs[[1]]$children[[m]]$children[[n]]$gp$col <- 'white'
      # }
      k <- k+1
    }

    if(save){
      ggsave2(
        paste0('tl20-05_side-by-side_barchart_facetbySubfam_',
               time_vec,
               file_name_end,
               '.png'),
        plot = g,
        dpi = 300,
        width = 28,
        height = 8,
        units = 'cm'
      )
    }

    grid.newpage()
    grid.draw(g)

}

 lapply(
    sort(unique(cir_plot$time)),
    side_by_side_cir_plot,
    threshold_min = 50,
    threshold_max = Inf,
    point_size = 1,
    text_size = 10,
    file_name_end = '_Over100TPM',
    save = TRUE
  )
```

```{r}
  lapply(
    sort(unique(cir_plot$time)),
    side_by_side_cir_plot,
    threshold_min = 0,
    threshold_max = 50,
    point_size = 0.1,
    text_size = 3,
    file_name_end = '_Under50TPM',
    save = TRUE
  )
#Note that the file name above may not be accurate - this is just so that I don't have to make the supp figure from scratch.
```

```{r pirs_side-by-side_figure4}

housekeeping_genes <- data.frame('HSP70' = 'PCHAS_0721000',
                             'MSP1' = 'PCHAS_0831300',
                             'AMA1' = 'PCHAS_0931000')

cir_housekeeping_id <- c(cir_id, 
                         'PCHAS_0721000',
                         'PCHAS_0831300',
                         'PCHAS_0931000')

cir_plot <- twentyfour_ct_comb$tpm %>% 
  filter(Geneid %in% cir_housekeeping_id) %>%
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(
    stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
    transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1],
    diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3],
    time = str_split_fixed(sample, pattern = '_', n = 3)[,2],
    subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)]
  ) %>% 
  mutate(subfam = ifelse(is.na(subfam), 'ref', subfam)) %>% 
  filter(!(time %in% c('sch.culture', 
                       'Unknown'))) %>% 
  mutate(subfam = str_replace(subfam, 
                              'ancestral',
                              'a')) %>% 
  mutate(
    subfam = interaction(
      transmission,
      forcats::fct_shift(
        relevel(
          as.factor(subfam),
          ref = 'ref')
      ),
      sep = '_',
      lex.order = TRUE
    )
  )
#Need to add the transmission name to the subfam so that they are ordered correctly.
# Need to move 'ref' factor to the start of the subfam, then shift the factor levels to the 'left' (so now
# ref is at the end) ensuring that the 'ref' genes are at the end of the graph for each transmission type.

cir_plot$Geneid <- ifelse(
  cir_plot$Geneid %in% housekeeping_genes[1,],
  colnames(housekeeping_genes)[ match(cir_plot$Geneid, housekeeping_genes[1,]) ],
  cir_plot$Geneid
)

limits_cir <- c(0, 
                mround(max(cir_plot$tpm), 
                       base = 100))

cir_plot_order <- cir_plot %>% 
  filter(time == '14h',
         transmission == 'rMT') %>% 
  arrange(desc(tpm))

cir_plot <- mutate(cir_plot, 
                   Geneid = factor(Geneid, 
                                   levels = unique(cir_plot_order$Geneid)))

transmissionSplit_cir_plot <- function(time_vec = '14h',
                                  log_it = TRUE,
                                  save = TRUE,
                                  limits = limits_cir,
                                  point_size = 1,
                                  text_size = 10,
                                  file_name_end = '',
                                  threshold_min = 100,
                                  threshold_max = Inf){

  cir_plot <- group_by(cir_plot, 
                       Geneid) %>% 
    filter(
      time == time_vec,
      !Geneid %in% cir_id | any(tpm >= threshold_min),
      !Geneid %in% cir_id | all(tpm <= threshold_max)
    )

  # cir_plot %>% group_by(Geneid) %>% summarise(n())
  if(log_it){
    cir_plot$tpm <- log2(1+cir_plot$tpm)
      limits = log2(1+limits)
      breaks <- round(seq.int(0, 
                              max(limits), 
                              length.out = round(max(limits)) + 1, 
                              digits = 0), 
                      digits = 0)
    }else{
      breaks <- seq.int(0, max(limits), length.out = 11)
    }

    cir_plot$Geneid <- factor(
      str_remove(cir_plot$Geneid, 
                 pattern = 'PCHAS_'),
      levels = str_remove(levels(cir_plot$Geneid), 
                          pattern = 'PCHAS_')
    )

    y <- ggplot(
      data = cir_plot,
      mapping = aes(
        x = Geneid,
        fill = transmission,
        y = tpm
      )
    ) +
      stat_summary(fun.y = median, 
                   fun.ymin = median, 
                   fun.ymax = median,
                   geom = "bar", 
                   width = 0.75, 
                   size = 0.25, 
                   position = 'dodge') +
      geom_point(stat = 'identity', 
                 position = position_dodge(width = 0.75),
                 size = point_size) +
      scale_y_continuous(breaks = breaks,
                         limits = limits) +
      scale_fill_manual(values = c('skyblue','red')) +
      facet_grid(.~subfam, 
                 scales = "free_x",
                 space = "free_x",
                 labeller = labeller(subfam = function(string) {
                   str_remove_all(string, 
                                  pattern = '(rMT|SBP)_')
                 })) +
      theme_classic() +
      theme(
        axis.text.x = element_text(size = text_size,
                                   angle = 90,
                                   # hjust = 2,
                                   vjust = 0.5)
      ) +
      ylab('log(TPM + 1)')+
      ggtitle(paste(unique(cir_plot$time)))

    #This bit is to change the colour, and text colour, of the facet boxes

    g <- ggplot_gtable(ggplot_build(y))

    # g$widths[5] = 2*g$widths[5]

    strip_t <- which(grepl('strip-t', g$layout$name))
    subfam_vec <- sort(unique(cir_plot$subfam)) %>% str_remove_all(pattern = '(rMT|SBP)_')
    fills <- case_when(str_detect(subfam_vec, 
                                  pattern = 'a') ~ clade_col['ancestral'],
                       subfam_vec %in% names(clade_col) ~ clade_col[paste(subfam_vec)],
                       str_detect(subfam_vec, 
                                  pattern = 'ref') ~ 'white')
    k <- 1
    for (i in strip_t) {
      j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
      g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
      k <- k+1
    }

    #subfam_vec also corresponds to which panel the different clades are, so for the small ones we can use this to
    # manually make them a bit wider.

    short_panels <- which(str_detect(subfam_vec, pattern = 'a|ref'))

    for(i in short_panels){
      num <- g$layout$l[grep(paste0('panel-1-', i, '$'), g$layout$name)]
      g$widths[num] = 1.75*g$widths[num]
      #1.75 times the width of the right panels.
      # print(num)
    }

    if(save){
      ggsave2(
        paste0('tl19-07_barchart_facetTrans_facetSubfam_',
               time_vec,
               file_name_end,
               '.png'),
        plot = g,
        dpi = 300,
        width = 28,
        height = 8,
        units = 'cm'
      )
    }

    grid.newpage()
    grid.draw(g)

}

# transmissionSplit_cir_plot(
#   sort(unique(cir_plot$time))[1],
#     threshold_min = 100,
#     threshold_max = Inf,
#     point_size = 1,
#     text_size = 4,
#     file_name_end = '_Over100TPM',
#     save = FALSE)

 lapply(
    sort(unique(cir_plot$time)),
    transmissionSplit_cir_plot,
    threshold_min = 100,
    threshold_max = Inf,
    point_size = 1,
    text_size = 5,
    file_name_end = '_Over100TPM',
    save = TRUE
  )
```

```{r pirs_side-by-side_figure4_RMT-SBP-completelySeparate}

cir_plot <- twentyfour_ct_comb$tpm %>% 
  filter(Geneid %in% cir_housekeeping_id) %>%
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(
    stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
    transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1],
    diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3],
    time = str_split_fixed(sample, pattern = '_', n = 3)[,2],
    subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)]
  ) %>% 
  mutate(subfam = ifelse(is.na(subfam), 'ref', subfam)) %>% 
  filter(!(time %in% c('sch.culture', 
                       'Unknown'))) %>% 
  mutate(subfam = str_replace(subfam, 
                              'ancestral',
                              'a')) %>% 
  mutate(
    subfam = interaction(
      transmission,
      forcats::fct_shift(
        relevel(
          as.factor(subfam),
          ref = 'ref')
      ),
      sep = '_',
      lex.order = TRUE
    )
  )
#Need to add the transmission name to the subfam so that they are ordered correctly.
# Need to move 'ref' factor to the start of the subfam, then shift the factor levels to the 'left' (so now
# ref is at the end) ensuring that the 'ref' genes are at the end of the graph for each transmission type.

cir_plot$Geneid <- ifelse(
  cir_plot$Geneid %in% housekeeping_genes[1,],
  colnames(housekeeping_genes)[ match(cir_plot$Geneid, housekeeping_genes[1,]) ],
  cir_plot$Geneid
)

limits_cir <- c(0, 
                mround(max(cir_plot$tpm), 
                       base = 100))

cir_plot_order <- cir_plot %>% 
  filter(time == '14h',
         transmission == 'rMT') %>% 
  arrange(desc(tpm))

cir_plot <- mutate(cir_plot, 
                   Geneid = factor(Geneid, 
                                   levels = unique(cir_plot_order$Geneid)))

transmissionSep_cir_plot <- function(time_vec = '14h',
                                     log_it = TRUE,
                                     save = TRUE,
                                     limits = limits_cir,
                                     point_size = 1,
                                     text_size = 10,
                                     file_name_end = '',
                                     threshold_min = 100,
                                     threshold_max = Inf){

  cir_plot_filter <- group_by(cir_plot, 
                       Geneid) %>% 
    filter(
      time == time_vec,
      !Geneid %in% cir_id | any(tpm >= threshold_min),
      !Geneid %in% cir_id | all(tpm <= threshold_max)
    )
  #Filter for pirs expressed above or below the thresholds, have the OR so that the ref genes aren't removed.
  # e.g. Either 'Not a cir' OR 'Any TPM is above the minimum'.

  # cir_plot_filter %>% group_by(Geneid) %>% summarise(n())
  if(log_it){
    cir_plot_filter$tpm <- log2(1+cir_plot_filter$tpm)
    limits = log2(1+limits)
    breaks <- round(seq.int(0, 
                            max(limits), 
                            length.out = round(max(limits)) + 1, 
                            digits = 0), 
                    digits = 0)
  }else{
    breaks <- seq.int(0, max(limits), length.out = 11)
  }

  cir_plot_filter$Geneid <- factor(
    str_remove(cir_plot_filter$Geneid, 
               pattern = 'PCHAS_'),
    levels = str_remove(levels(cir_plot_filter$Geneid), 
                        pattern = 'PCHAS_')
  )

  lapply(unique(cir_plot_filter$transmission), 
         function(transmission_method){

           cir_plot_filter <- filter(cir_plot_filter, 
                              transmission %in% transmission_method)

           y <- ggplot(
             data = cir_plot_filter,
             mapping = aes(
               x = Geneid,
               y = tpm
             )
           ) +
             stat_summary(fun.y = median, 
                          fun.ymin = median, 
                          fun.ymax = median,
                          geom = "bar",
                          fill = 'lightblue',
                          width = 0.75, 
                          size = 0.25, 
                          position = 'dodge') +
             geom_point(stat = 'identity', 
                        position = position_dodge(width = 0.75),
                        size = point_size) +
             scale_y_continuous(breaks = breaks,
                                limits = limits) +
             facet_grid(.~subfam, 
                        scales = "free_x",
                        space = "free_x",
                        labeller = labeller(subfam = function(string) {
                          str_remove_all(string, 
                                         pattern = '(rMT|SBP)_')
                        })) +
             theme_classic() +
             ylab('log(TPM + 1)')+
             ggtitle(paste(unique(cir_plot_filter$time)))

           if(grepl(time_vec, pattern = '20|23')) {
             y <- y + theme(
               axis.text.x = element_text(size = text_size,
                                          angle = 90,
                                          vjust = 0.5)
             )
           } else {
             y <- y + theme(
               axis.text.x = element_blank(),
               axis.ticks.x = element_blank()
             ) 
           }

           #This bit is to change the colour, and text colour, of the facet boxes

           g <- ggplot_gtable(ggplot_build(y))

           # g$widths[5] = 2*g$widths[5]

           strip_t <- which(grepl('strip-t', g$layout$name))
           subfam_vec <- sort(unique(cir_plot_filter$subfam)) %>% str_remove_all(pattern = '(rMT|SBP)_')
           fills <- case_when(str_detect(subfam_vec, 
                                         pattern = 'a') ~ clade_col['ancestral'],
                              subfam_vec %in% names(clade_col) ~ clade_col[paste(subfam_vec)],
                              str_detect(subfam_vec, 
                                         pattern = 'ref') ~ 'white')
           k <- 1
           for (i in strip_t) {
             j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
             g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
             k <- k+1
           }

           #subfam_vec also corresponds to which panel the different clades are, so for the small ones we can use this to
           # manually make them a bit wider.

           short_panels <- which(str_detect(subfam_vec, pattern = 'a|S1'))

           for(i in short_panels){
             num <- g$layout$l[grep(paste0('panel-1-', i, '$'), g$layout$name)]
             g$widths[num] = 1.75*g$widths[num]
             #1.75 times the width of the right panels.
             # print(num)
           }

           if(save){
             ggsave2(
               paste0('tl19-07_barchart_',
                      transmission_method,
                      '_',
                      'facetSubfam_',
                      time_vec,
                      file_name_end,
                      '.png'),
               plot = g,
               dpi = 300,
               width = 28,
               height = 8,
               units = 'cm'
             )
           }

           grid.newpage()
           grid.draw(g)
         })

}

# transmissionSep_cir_plot(
#   sort(unique(cir_plot$time))[1],
#   threshold_min = 50,
#   threshold_max = Inf,
#   point_size = 1,
#   text_size = 6,
#   file_name_end = '_Over50TPM',
#   save = FALSE)

lapply(
  sort(unique(cir_plot$time)),
  transmissionSep_cir_plot,
  threshold_min = 50,
  threshold_max = Inf,
  point_size = 1,
  text_size = 7.5,
  file_name_end = '_Over50TPM',
  save = TRUE
)
```

```{r ancestral_expression_barchart}

# housekeeping_genes <- data.frame('HSP70' = 'PCHAS_0721000',
#                              'MSP1' = 'PCHAS_0831300',
#                              'AMA1' = 'PCHAS_0931000')

cir_plot <- twentyfour_ct_comb$tpm %>% 
  filter(Geneid %in% c('PCHAS_0101200', 
                       'PCHAS_0931000')) %>%
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(
    stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
    transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1],
    diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3],
    time = str_split_fixed(sample, pattern = '_', n = 3)[,2],
    subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)]
  ) %>% 
  mutate(subfam = ifelse(is.na(subfam), 'ref', subfam)) %>% 
  filter(!(time %in% c('sch.culture', 
                       'Unknown')),
         transmission == 'rMT') %>% 
  mutate(subfam = str_replace(subfam, 
                              'ancestral',
                              'a')) %>% 
  mutate(
    subfam = interaction(
      transmission,
      forcats::fct_shift(
        relevel(
          as.factor(subfam),
          ref = 'ref')
      ),
      sep = '_',
      lex.order = TRUE
    )
  )

anc <- ggplot(data = cir_plot,
       aes(x = time,
           y = tpm,
           fill = Geneid)) +
  stat_summary(fun.y = median, 
               fun.ymin = median, 
               fun.ymax = median,
               geom = "bar", 
               width = 0.75, 
               size = 0.25, 
               position = 'dodge') +
  geom_point(stat = 'identity', 
             position = position_dodge(width = 0.75)) +
  scale_y_continuous(n.breaks = 11) +
   scale_fill_manual(values = cbPalette[c(5,8)],
                     labels = c('ancestral pir', 
                                'AMA1')) +
  xlab('Time point') + ylab('TPM') +
  theme_classic()

ggsave(filename = 'tl19-07_barchart_tpm_RMT_ancestralpirAMA1.png',
       plot = anc,
       dpi = 300,
       width = 20,
       height = 10,
       units = 'cm')

```

```{r rmt_pirs_higher_than_sbp_are_expressed_more}
cir_rmt_sbp_hi <- twentyfour_ct_comb$tpm %>% 
  filter(Geneid %in% cir_id) %>%
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(
    stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
    transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1],
    diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3],
    time = str_split_fixed(sample, pattern = '_', n = 3)[,2],
    subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)]
  ) %>% 
  filter(time == '14h') %>% 
  group_by(transmission, time, Geneid) %>% 
  summarise(tpm_mean = mean(tpm)) %>% 
  dcast(Geneid ~ transmission, margins = c('tpm_mean')) %>% 
  mutate(diff = ifelse(rMT - SBP > 0, 'RMT', 'SBP')) %>% 
  melt(id.vars = c('Geneid', 'diff'), 
       variable.name = 'transmission',
       value.name = 'tpm_mean')

cir_rmt_sbp_hi %>% 
  group_by(diff) %>% 
  summarise(tpm_mean = mean(tpm_mean))

ggplot(cir_rmt_sbp_hi,
       aes(x = diff,
           y = log2(tpm_mean+1),
           col = transmission)) +
  geom_point(position = 'jitter') + geom_boxplot()
```

# Supp figure - Pir expression of AAPLs and ChAPLs.

```{r Supp_fig_AapL-ChaAPl_expression}

  cir_plot <- twentyfour_ct_comb$tpm %>% 
    filter(Geneid %in% cir_id) %>%
    melt(id.vars = 'Geneid', 
         variable.name = 'sample', 
         value.name = 'tpm') %>% 
    mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
           transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1],
           diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3],
           time = str_split_fixed(sample, pattern = '_', n = 3)[,2],
           subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)],
           chapl_aapl = factor(cir_chapl_aapl_info$chapl_aapl[match(Geneid, 
                                                                    cir_chapl_aapl_info$Geneid)],
                               levels = sort(unique(cir_chapl_aapl_info$chapl_aapl[match(Geneid, 
                                                                                         cir_chapl_aapl_info$Geneid)])))) %>% 
    filter(!(time %in% c('sch', 
                         'Unknown')),
           !is.na(chapl_aapl) ) %>% 
    group_by(Geneid) %>% 
    # filter(
    #   # all(tpm < 100),
    #   #      any(tpm > 50)
    #   subfam == 'S7'
    #   ) %>% 
    arrange(desc(tpm))

  cir_plot_order <- cir_plot %>% 
    filter(time == '14h',
           transmission == 'SBP') %>% 
    arrange(desc(tpm))

  cir_plot <- mutate(cir_plot, 
                     Geneid = factor(Geneid, 
                                     levels = unique(cir_plot_order$Geneid)))
  cir_plot_chapl_aapl_supp_info <- cir_plot

  max_aapl_chapl <- ceiling(max(log2(cir_plot$tpm+1)))

  breaks = seq(0, max_aapl_chapl, by = 1)
    limits = c(0, max_aapl_chapl)
    point_size = 0.25

  lapply(
    sort(unique(cir_plot$time)),
    function(time_vec){

    cir_plot <- filter(cir_plot, 
                       time == time_vec)

      cir_plot$tpm <- log2(1+cir_plot$tpm)

    cir_plot$Geneid <- factor(
      str_remove(cir_plot$Geneid, 
                 pattern = 'PCHAS_'),
      levels = str_remove(levels(cir_plot$Geneid), 
                          pattern = 'PCHAS_')
    )

    y <- ggplot(
      data = cir_plot,
      mapping = aes(
        x = Geneid,
        fill = transmission,
        y = tpm
      )
    ) +
    stat_summary(fun.y = median, 
                 fun.ymin = median, 
                 fun.ymax = median,
                 geom = "bar", 
                 width = 0.75, 
                 size = 0.25, 
                 position = 'dodge') +
    geom_point(stat = 'identity', 
               position = position_dodge(width = 0.75),
               size = point_size) +
    scale_y_continuous(breaks = breaks,
                       limits = limits) +
      scale_fill_manual(values = c('skyblue','red')) +
      ylab('log(TPM +1)') +
    facet_grid(.~chapl_aapl, 
             scales = "free_x",
             space = "free_x") +
    theme_classic() +
    theme(
      axis.text.x = element_text(angle = 90,
                                 size = 5)
    ) +
    ggtitle(paste(unique(cir_plot$time)))

    ggsave2(
      paste0('tl20-05_side-by-side_barchart_chapl_aapl_',
             time_vec,
             '.png'),
      plot = y,
      dpi = 300,
      width = 25,
      height = 10,
      units = 'cm'
    )
  return(y)
})
```

# Chromosome 6 expression, split by ChAPL designation (mostly a ChAPL)

```{r Chr6-ChaAPl_expression}

  cir_plot <- twentyfour_ct_comb$tpm %>% 
    filter(Geneid %in% cir_id) %>%
    melt(id.vars = 'Geneid', 
         variable.name = 'sample', 
         value.name = 'tpm') %>% 
    mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
           transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1],
           diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3],
           time = str_split_fixed(sample, pattern = '_', n = 3)[,2],
           subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)],
           chrom = str_extract(Geneid, pattern = '(?<=PCHAS_)[:digit:]{2}'),
           chapl_aapl = factor(cir_chapl_aapl_info$chapl_aapl[match(Geneid, 
                                                                    cir_chapl_aapl_info$Geneid)],
                               levels = sort(unique(cir_chapl_aapl_info$chapl_aapl[match(Geneid, 
                                                                                         cir_chapl_aapl_info$Geneid)])))) %>% 
    filter(!(time %in% c('sch', 
                         'Unknown')),
             chrom == '06') %>% 
    group_by(Geneid) %>% 
    # filter(
    #   # all(tpm < 100),
    #   #      any(tpm > 50)
    #   subfam == 'S7'
    #   ) %>% 
    arrange(desc(tpm))

  cir_plot_order <- cir_plot %>% 
    filter(time == '14h',
           transmission == 'SBP') %>% 
    arrange(desc(tpm))

  cir_plot <- mutate(cir_plot, 
                     Geneid = factor(Geneid, 
                                     levels = unique(cir_plot_order$Geneid)))

  breaks = c(7,6,5,4, 3, 2, 1, 0)
    limits = c(0,7)
    point_size = 0.25

  lapply(
    sort(unique(cir_plot$time)),
    function(time_vec){

    cir_plot <- filter(cir_plot, 
                       time == time_vec)

      cir_plot$tpm <- log2(1+cir_plot$tpm)

    cir_plot$Geneid <- factor(
      str_remove(cir_plot$Geneid, 
                 pattern = 'PCHAS_'),
      levels = str_remove(levels(cir_plot$Geneid), 
                          pattern = 'PCHAS_')
    )

    y <- ggplot(
      data = cir_plot,
      mapping = aes(
        x = Geneid,
        fill = transmission,
        y = tpm
      )
    ) +
    stat_summary(fun.y = median, 
                 fun.ymin = median, 
                 fun.ymax = median,
                 geom = "bar", 
                 width = 0.75, 
                 size = 0.25, 
                 position = 'dodge') +
    geom_point(stat = 'identity', 
               position = position_dodge(width = 0.75),
               size = point_size) +
    scale_y_continuous(breaks = breaks,
                       limits = limits) +
    facet_grid(.~chapl_aapl, 
             scales = "free_x",
             space = "free_x") +
    theme_classic() +
    theme(
      axis.text.x = element_text(angle = 90,
                                 size = 5)
    ) +
    ggtitle(paste(unique(cir_plot$time)))

    ggsave2(
      paste0('tl20-05_side-by-side_barchart_chr6_chapl_',
             time_vec,
             '.png'),
      plot = y,
      dpi = 300,
      width = 25,
      height = 10,
      units = 'cm'
    )
  return(y)
})
```

## Bar chart of S and L clade expression across time.

```{r SorL_barchart}
tpm_data_SorL <- twentyfour_ct_comb$tpm %>% 
  filter(Geneid %in% cir_id) %>% 
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, 
                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
    mutate(stage = times_stages_table$time[match(stage, times_stages_table$r_stage_diff_count)]) %>% 
  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
  filter(stage != 'sch.culture') %>% 
  mutate(clade = cir_clade_info$clade[match(Geneid, 
                                            cir_clade_info$Geneid)]) %>% 
  mutate(SorL = str_extract(clade, 
                            pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'),
         # stage =  factor(stage, levels = ordering(stage, order_vector))
         ) %>% 
  group_by(transmission,
           stage, 
           SorL) %>% 
  dplyr::summarise(total_tpm = sum(tpm))

y <- ggplot(data = tpm_data_SorL, 
              aes(
                x = transmission,
                y = total_tpm, 
                fill = SorL,
                group = transmission
              )) +
    geom_bar(position = position_stack(),
             stat = 'identity',
             colour = 'black',
             size = 0.1) +
    scale_fill_manual(
      values = SorL_col,
      guide = guide_legend(reverse = TRUE)
    ) +
    scale_color_manual('black') + 
    # labs(fill = stack_label) +
    # scale_y_continuous(breaks = max_tpm) +
    theme_classic() +
    scale_size(range = c(0,3)) +
    ylab(
      bquote('total '~italic(.('pir'))~' TPM')
    ) + 
    facet_grid(.~stage, 
               scales = "free_x",
               space = "free_x") +
    theme(
      axis.title.x=element_blank(),
      # axis.text.x=element_blank(),
      axis.text.y = element_text(size = 5),
      axis.title.y = element_text(size = 8),
      legend.title = element_text(size = 7),
      legend.text = element_text(size = 5),
      legend.key.size = unit(3,'mm'),
      legend.key.width = unit(2,'mm'),
      strip.text.x = element_text(size=7, 
                                  face = 'bold', 
                                  angle = 0),
      strip.background = element_rect(
        color="black", 
        size=0.1, 
        linetype="solid"
      )
    ) 
y

ggsave2('tl20-05_clades_totalTPM_barchart.pdf',
        plot = y,
        dpi = 300,
        width = 15,
        height = 10,
        units = 'cm')

```

## Clades proportion bar chart

```{r clades_prop_barchart}
tpm_data_clade <- twentyfour_ct_comb$tpm %>% 
  filter(Geneid %in% cir_id) %>% 
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, 
                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
  mutate(stage = times_stages_table$time[match(stage, times_stages_table$r_stage_diff_count)]) %>% 
  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
  filter(stage != 'sch.culture') %>% 
  mutate(clade = cir_clade_info$clade[match(Geneid, 
                                            cir_clade_info$Geneid)]) %>% 
  # mutate(stage =  factor(stage, 
  #                        levels = ordering(stage, 
  #                                          order_vector))) %>% 
  group_by(transmission,
           stage, 
           clade) %>% 
  dplyr::summarise(total_tpm = sum(tpm)) %>% 
  arrange(stage)

y <- ggplot(data = tpm_data_clade, 
              aes(
                x = stage,
                y = total_tpm, 
                fill = clade,
                group = transmission
              )) +
    geom_bar(position = 'fill',
             stat = 'identity',
             colour = 'black',
             size = 0.1) +
    scale_fill_manual(
      breaks = sort(unique(tpm_data_clade$clade)),
      values = clade_col,
      guide = guide_legend(reverse = TRUE) 
    ) +
    scale_color_manual('black') + 
    # labs(fill = stack_label) +
    # scale_y_continuous(breaks = max_tpm) +
    theme_classic() +
    scale_size(range = c(0,3)) +
    ylab(
      bquote('proportion of total '~italic(.('pir'))~' TPM')
    ) + 
    facet_grid(.~transmission, 
               scales = "free_x",
               space = "free_x") +
    theme(
      axis.title.x=element_blank(),
      axis.text.x=element_text(size = 7, 
                                 angle = 90),
      axis.text.y = element_text(size = 5),
      axis.title.y = element_text(size = 8),
      legend.title = element_text(size = 7),
      legend.text = element_text(size = 5),
      legend.key.size = unit(3,'mm'),
      legend.key.width = unit(2,'mm'),
      strip.text.x = element_text(size=7, 
                                  face = 'bold', 
                                  angle = 0),
      strip.background = element_rect(
        color="black", 
        size=0.1, 
        linetype="solid"
      )
    ) 
y

ggsave2('tl20-05_subclades_propTPM_barchart.png',
        plot = y,
        dpi = 300,
        width = 15,
        height = 10,
        units = 'cm')

```

```{r add_berghei_SorL_prop}
#Add berghei data
berghei_SorL_file <- list.files('../Manuscript_pirAnalysis_Github/Pberghei/', 
                                recursive = TRUE, 
                                pattern = 'tl20-05_SorL-subfamily_pir.csv', 
                                full.names = TRUE)
berghei_SorL <- read_csv(berghei_SorL_file) %>% 
  filter(stage %in% c('Asex.SchizA',
                      'Asex.SchizB',
                      'Asex.Ring',
                      'Asex.Troph',
                      'Asex.Mixed')) 
berghei_SorL_avg <- berghei_SorL %>% 
  group_by(stage,
           SorL) %>% 
  dplyr::summarise(total_tpm = sum(total_tpm)) %>% 
  mutate(SorL = str_extract(SorL,
                            pattern = 'S|L|ancestral'),
         transmission = 'berghei')

```

```{r}
y <- ggplot(data = rbind(
  tpm_data_SorL,
  berghei_SorL_avg
), 
aes(
  x = stage,
  y = total_tpm, 
  fill = SorL,
  # group = transmission
)) +
  geom_bar(position = 'fill',
           stat = 'identity',
           colour = 'black',
           size = 0.1) +
  scale_fill_manual(
    values = SorL_col,
    guide = guide_legend(reverse = FALSE)
  ) +
  scale_color_manual('black') + 
  # labs(fill = stack_label) +
  # scale_y_continuous(breaks = max_tpm) +
  theme_classic() +
  scale_size(range = c(0,3)) +
  ylab(
    bquote('proportion of total '~italic(.('pir'))~' TPM')
  ) + 
  facet_grid(.~transmission, 
             scales = "free_x",
             space = "free_x") +
  labs(fill = 'sub-family') +
  theme(
    axis.title.x=element_blank(),
    axis.text.x=element_text(angle = 90),
    axis.text.y = element_text(size = 5),
    axis.title.y = element_text(size = 8),
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 5),
    legend.key.size = unit(3,'mm'),
    legend.key.width = unit(2,'mm'),
    strip.text.x = element_text(size=7, 
                                face = 'bold', 
                                angle = 0),
    strip.background = element_rect(
      color="black", 
      size=0.1, 
      linetype="solid"
    )
  ) 
y

ggsave2('tl20-05_SorL_totalTPM_facet-transmission_prop_barchart.pdf',
        plot = y,
        dpi = 300,
        width = 15,
        height = 10,
        units = 'cm')
```

```{r SorL_barchart_transmission-facet}
tpm_data_SorL <- twentyfour_ct_comb$tpm %>% 
  filter(Geneid %in% cir_id) %>% 
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, 
                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
    mutate(stage = times_stages_table$time[match(stage, times_stages_table$r_stage_diff_count)]) %>% 
  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
  filter(stage != 'sch.culture') %>% 
  mutate(clade = cir_clade_info$clade[match(Geneid, 
                                            cir_clade_info$Geneid)]) %>% 
  mutate(SorL = str_extract(clade, 
                            pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'),
         # stage =  factor(stage, levels = ordering(stage, order_vector))
         ) %>% 
  group_by(transmission,
           stage, 
           SorL) %>% 
  dplyr::summarise(total_tpm = sum(tpm)) %>% 
  mutate(SorL = factor(SorL, 
                       levels = sort(unique(SorL))))

y <- ggplot(data = tpm_data_SorL, 
              aes(
                x = stage,
                y = total_tpm, 
                fill = SorL,
                # group = transmission
              )) +
  geom_bar(position = position_stack(),
           stat = 'identity',
           colour = 'black',
           size = 0.1) +
  scale_fill_manual(
    values = SorL_col,
    guide = guide_legend(reverse = FALSE)
  ) +
  scale_color_manual('black') + 
  # labs(fill = stack_label) +
  # scale_y_continuous(breaks = max_tpm) +
  theme_classic() +
  scale_size(range = c(0,3)) +
  ylab(
    bquote('total '~italic(.('pir'))~' TPM')
  ) + 
  facet_grid(.~transmission, 
             scales = "free_x",
             space = "free_x") +
  labs(fill = 'sub-family') +
    theme(
      axis.title.x=element_blank(),
      # axis.text.x=element_blank(),
      axis.text.y = element_text(size = 5),
      axis.title.y = element_text(size = 8),
      legend.title = element_text(size = 7),
      legend.text = element_text(size = 5),
      legend.key.size = unit(3,'mm'),
      legend.key.width = unit(2,'mm'),
      strip.text.x = element_text(size=7, 
                                  face = 'bold', 
                                  angle = 0),
      strip.background = element_rect(
        color="black", 
        size=0.1, 
        linetype="solid"
      )
    ) 
y

ggsave2('tl20-05_SorL_totalTPM_facet-transmission_barchart.pdf',
        plot = y,
        dpi = 300,
        width = 15,
        height = 10,
        units = 'cm')

```

```{r proportion_data}

tpm_data_twentyfour <- twentyfour_ct_comb$tpm %>% 
  filter(Geneid %in% cir_id) %>% 
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, 
                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
  filter(stage != 'sch.culture') %>% 
  mutate(clade = cir_clade_info$clade[match(Geneid, 
                                            cir_clade_info$Geneid)],
         chrom = str_extract(Geneid, 
                             pattern = '(?<=PCHAS_)[[:digit:]]{2}')
         ) %>% 
  mutate(
    SorL = str_extract(clade, 
                       pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'),
    # stage =  factor(stage, 
    #                 levels = ordering(stage, order_vector))
  ) %>% 
  group_by(transmission,
           stage, 
           Geneid,
           clade,
           chrom,
           SorL) %>% 
  dplyr::summarise(tpm = mean(tpm))

```

```{r chrom_prop_bar_chart}
tpm_data_chrom <- tpm_data_twentyfour %>%
  mutate(stage = times_stages_table$time[match(stage,
                                               times_stages_table$r_stage_diff_count)]) %>%
  group_by(transmission,
           stage,
           chrom) %>%
  dplyr::summarise(total_tpm = sum(tpm))

y <- ggplot(data = tpm_data_chrom) +
  geom_bar_pattern(
    aes(
      x = stage,
      y = total_tpm, 
      fill = chrom,
      group = transmission,
      pattern = chrom
    ),
    position = 'fill',
    stat = 'identity',
    colour = 'black',
    size = 0.1,
    pattern_density = 0.05,
    pattern_spacing = 0.02,
    pattern_key_scale_factor = 0.5
  ) +
  scale_pattern_manual(
    values = rep(c('none', 
                   'circle'), 
                 each = length(c(cbPalette, 
                                 'red', 
                                 'white')), 
                 length.out = length(unique(tpm_data_chrom$chrom))),
    labels = unique(tpm_data_chrom$chrom),
    guide = guide_legend(reverse = TRUE),
    name = 'chromosome'
  ) +
  scale_fill_manual(
    breaks = unique(tpm_data_chrom$chrom),
    values = rep_len(c(cbPalette, 
                       'red', 
                       'white'),
                     length.out = length(unique(tpm_data_chrom$chrom))),
    labels =  unique(tpm_data_chrom$chrom),
    guide = guide_legend(reverse = TRUE),
    name = 'chromosome'
  ) +
  scale_color_manual('black') + 
  # labs(fill = stack_label) +
  # scale_y_continuous(breaks = max_tpm) +
  theme_classic() +
    scale_size(range = c(0,3)) +
    ylab(
      bquote('proportion of '~italic(.('pir'))~' TPM')
    ) + 
    facet_grid(.~transmission, 
               scales = "free_x",
               space = "free_x") +
    theme(
      axis.title.x=element_blank(),
      axis.text.x=element_text(size = 10, 
                                 angle = 0),
      axis.text.y = element_text(size = 5),
      axis.title.y = element_text(size = 8),
      legend.title = element_text(size = 7),
      legend.text = element_text(size = 5),
      legend.key.size = unit(3,'mm'),
      legend.key.width = unit(2,'mm'),
      strip.text.x = element_text(size=10, 
                                  face = 'bold', 
                                  angle = 0),
      strip.background = element_rect(
        color="black", 
        size=0.1, 
        linetype="solid"
      )
    ) 
y

ggsave2('tl19-05_chroms_propTPM_barchart.png',
        plot = y,
        dpi = 330,
        width = 15,
        height = 10,
        units = 'cm')

```

### 2D plot of PC1 v PC2 - Supp figure

```{r pca_calculation, results = TRUE}
twentyfour_tpm <- twentyfour_ct_comb_noschz$tpm
rownames_geneid <- twentyfour_tpm$Geneid
twentyfour_tpm_forpca <- select(twentyfour_tpm, -Geneid)
rownames(twentyfour_tpm_forpca) <- rownames_geneid
twentyfour_tpm_pca <- prcomp(twentyfour_tpm_forpca, center = TRUE, scale. = TRUE)
#Need to look into what exactly scale and center are doing
```

```{r 2d_pca, results = TRUE}
twentyfour_tpm_pca_var <- get_pca_var(twentyfour_tpm_pca)
twentyfour_tpm_pca_coords <- data.frame(twentyfour_tpm_pca_var$coord)

twentyfour_tpm_pca_coords <- data.frame(sample = rownames(twentyfour_tpm_pca_var$coord),
                                        PC1 = twentyfour_tpm_pca_coords$Dim.1,
                                        PC3 = twentyfour_tpm_pca_coords$Dim.3,
                                        PC2 = twentyfour_tpm_pca_coords$Dim.2) %>% 
  dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
  dplyr::mutate(time = str_split_fixed(stage, pattern = '_', n = 2)[,2]) %>% 
  dplyr::mutate(transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1])

colornames <- viridis(length(unique(twentyfour_tpm_pca_coords$time)))
colornames <- setNames(colornames, 
                       sort(unique(twentyfour_tpm_pca_coords$time)))

pca_importance <- as.data.frame(summary(twentyfour_tpm_pca)$importance)
pc1_importance <- round(pca_importance$PC1[2]*100, digits = 1)
pc2_importance <- round(pca_importance$PC2[2]*100, digits = 1)
pc3_importance <- round(pca_importance$PC3[2]*100, digits = 1)

(
  plot <- ggplot(twentyfour_tpm_pca_coords,
                 aes(
                   x = PC1,
                   y = PC2,
                   col = time,
                   shape = transmission
                 )) +
    geom_point(size = 3,
               alpha = 0.8) +
    scale_color_manual(values = colornames) +
    scale_shape_manual(
      values = setNames(
        c(17,19),
        unique(twentyfour_tpm_pca_coords$transmission)
      )
    ) +
    xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
    ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
    guides(shape=guide_legend(ncol=2),
           col = guide_legend(ncol=2)) +
    theme_classic() +
    theme(legend.key.size = unit(4, 'mm'),
          legend.text = element_text(size = 7))
)
ggsave2(filename = 'tl19-07_chabaudi24h_supp_figure_pca_nopseudo.png',
        plot = plot,
        dpi = 300,
        width = 15,
        height = 10,
        units = 'cm')

```

# z scores

```{r read_z}
### Read z-scores

twentyfour_z <- list(
  'rMT' = list(zscores = read_csv('tl19-07_chabaudi24h_rMT.csv'),
               tpm = twentyfour_avgtpm_noschz$rMT$tpm,
               experiment = 'rMT'),
  'SBP' = list(zscores = read_csv('tl19-07_chabaudi24h_SBP.csv'),
               tpm = twentyfour_avgtpm_noschz$SBP$tpm,
               experiment = 'SBP')
)

###
```

# Phaseograms of the data. {.tabset}

```{r row_order_phaseograms}

#Getting the row order for each phaseogram for each experiment
# Want the data ordered by the highest zscore for each gene in each stage.
phaseogram_row_order <- lapply(twentyfour_z, 
                               function(list_elements){
                                 zscore_df <- list_elements$zscores
                                 zscore_ordered <- melt(zscore_df, 
                                                        id = 'Geneid',
                                                        variable.name = 'stage',
                                                        value.name = 'zscore') %>% 
                                   group_by(Geneid) %>% 
                                   top_n(zscore, n = 1) %>% 
                                   arrange(stage, 
                                           desc(zscore)) %>%
                                   distinct(Geneid, 
                                            .keep_all = TRUE) %>%  #Keep only the ones that are uniquely highest expressed in one sample
                                   filter(zscore > 0) #For some reason some remain with zero zscore
                                 # zscore_ordered <- zscore_ordered[order(
                                 #   match(
                                 #     zscore_ordered$stage, 
                                 #     order_vector
                                 #   )
                                 # ),]
                                 return(c(list_elements, 
                                          list(zscore_order = zscore_ordered)))
                               })

phaseogram_time_order <- lapply(names(phaseogram_row_order), 
                                function(list_element){
                                  transmission_list <- getElement(phaseogram_row_order,
                                                                  list_element)
                                  order_by_time <- transmission_list$zscore_order %>% 
                                    mutate(stage = times_stages_table$time[match(stage,
                                                                                 times_stages_table$r_stage_diff_count)]) %>% 
                                    arrange(stage)
                                  return(list(zscore_order = order_by_time))
                                }
)
names(phaseogram_time_order) <- names(phaseogram_row_order)

```

~~Change the lnTPM graph to a heatmap of max/min TPM a la Brugat 2017.~~

~~Add annotation for the clades,~~ or possibly separate the heatmap by clade.

Add threshold option.

```{r function_draw_heatmap_zscore}

draw_heatmap_zscore <- function(zscore_list,
                                title = '',
                                gene_ordering_vector = NULL,
                                filter_genes_vector = NULL,
                                extra_title = NULL,
                                annotation_clade = NULL,
                                annotation_col,
                                threshold_tpm = NULL,
                                use_times = FALSE,
                                use_order_vector = TRUE,
                                tpm_mat = TRUE,
                                tpm_breaks = NULL,
                                ...){
  if(is.null(tpm_breaks) & tpm_mat | is.numeric(threshold_tpm)){
    #Need the generate the breaks in the data for the heatmap legend if not provided.
    # Threshold_tpm also needs this in order to filter lowly expressed genes.
    max_tpm <- lapply(zscore_list, 
                      function(list_element){
                        #list_element <- zscore_list[[1]]
                        heatmap_data <- as.data.frame(list_element$zscores)
                        title <- list_element$experiment

                        if(!'Geneid' %in% colnames(heatmap_data)){
                          stop('No Geneid column in the heatmap data.frame')
                        }

                        if(!is.null(filter_genes_vector)){
                          heatmap_data <- as.data.frame(
                            filter(heatmap_data, 
                                   Geneid %in% filter_genes_vector)
                          )
                        }

                        rownames(heatmap_data) <- as.character(heatmap_data$Geneid)

                        heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
                        #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!

                        if(!is.null(gene_ordering_vector)){
                          #Remove genes not in the main data frame but in the ordering vector, 
                          # and then order based on the vector
                          remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
                          heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
                          heatmap_data <- heatmap_data[na.omit(
                            match(gene_ordering_vector, 
                                  rownames(heatmap_data))
                          ),,
                          drop=FALSE]
                        }

                        geneids <- rownames(heatmap_data)

                        #Get max/min TPM scores for the side annotation heatmap                       
                        tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
                                                          list_element$tpm$Geneid),] %>% 
                          select(-Geneid) %>% 
                          as.matrix %>% 
                          log1p
                        rownames(tpm_mat) <- geneids

                        tpm_min <- rowMins(tpm_mat)
                        names(tpm_min) <- geneids
                        tpm_max <- rowMaxs(tpm_mat)
                        names(tpm_max) <- geneids
                        return(list(tpm_max = tpm_max,
                                    tpm_min = tpm_min))
                      })

    tpm_max_all <- mround(
      max(
        unlist(
          lapply(max_tpm, 
                 getElement,
                 'tpm_max')
        )
      ), 
      base = 1)

    tpm_breaks <- round(
      c(0,
        tpm_max_all/4,
        tpm_max_all/2,
        tpm_max_all*(3/4),
        tpm_max_all),
      digits = 2
    )
  }

  #Generate the colour function from the TPM legend breaks for the TPM annotation
  if(tpm_mat){
    tpm_col_fun <- colorRamp2(
      breaks = tpm_breaks,
      colors = viridis(length(tpm_breaks))
    )
  }

  if(is.numeric(threshold_tpm)){
    #Apply a threshold to the TPM in order to remove low expressed genes.
    # Remember to add log1p here! 
    # Threshold uses the max value, so it ony needs to be above thresh 
    #   in one stage avg in one transmission method.
    #Will alter the gene_filter vector in order to only select the genes above threshold

    gene_maxs <- bind_rows(
      lapply(max_tpm, 
             getElement,
             'tpm_max')
    )

    remove_genes <- names(gene_maxs)[apply(
      gene_maxs, 
      2, 
      function(maxs){
        all(maxs < log2(1+threshold_tpm))
      }
    )]

    filter_genes_vector <- filter_genes_vector[!filter_genes_vector %in% remove_genes]

  }

  if(is.character(annotation_clade)){
    order_cir <- gene_ordering_vector[gene_ordering_vector %in% filter_genes_vector]
    clade_info <- getElement(cir_clade_info,
                             annotation_clade)[match(order_cir, 
                                                     cir_clade_info$Geneid)]
    names(clade_info) <- str_split_fixed(order_cir, 
                                         pattern = '_', 
                                         n = 2)[,2]
    annotation_col <- annotation_col
  }

  heatmaps <- lapply(zscore_list, 
                     function(list_element){
                       #list_element <- zscore_list[[1]]
                       heatmap_data <- as.data.frame(list_element$zscores)
                       title <- list_element$experiment

                       if(!'Geneid' %in% colnames(heatmap_data)){
                         stop('No Geneid column in the heatmap data.frame')
                       }

                       if(!is.null(filter_genes_vector)){
                         heatmap_data <- as.data.frame(
                           filter(heatmap_data, 
                                  Geneid %in% filter_genes_vector)
                         )
                       }

                       rownames(heatmap_data) <- as.character(heatmap_data$Geneid)

                       heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
                       #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!

                       if(!is.null(gene_ordering_vector)){
                         #Remove genes not in the main data frame but in the ordering vector,
                         #  and then order based on the vector
                         remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
                         heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
                         heatmap_data <- heatmap_data[na.omit(
                           match(gene_ordering_vector, 
                                 rownames(heatmap_data))
                         ),,
                         drop=FALSE]
                       }

                       if(tpm_mat){
                         #Get max/min TPM scores for the side annotation heatmap                       
                         tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
                                                           list_element$tpm$Geneid),] %>%
                           select(-Geneid) %>% 
                           as.matrix %>% 
                           log1p

                         tpm_min <- rowMins(tpm_mat)
                         tpm_max <- rowMaxs(tpm_mat)

                         names(tpm_min) = rownames(heatmap_data)
                         names(tpm_max) = rownames(heatmap_data)

                         tpm_ha <- rowAnnotation(
                           min.max = anno_simple(cbind(tpm_min, 
                                                       tpm_max), 
                                                 col = tpm_col_fun,
                                                 gp = gpar(col = 'black',
                                                           lwd = 0.1),
                                                 border = TRUE,
                                                 simple_anno_size = unit(0.2, "cm")),
                           show_annotation_name = FALSE)
                       }

                       #Draw the left annotation is needed, only for the first element.

                       if(
                         identical(list_element,
                                   dplyr::first(zscore_list)) & is.character(annotation_clade)
                       ){
                         clade_ha <- rowAnnotation(
                           'subfamily' = anno_simple(
                             clade_info,
                             col = annotation_col
                           )
                         )
                       }

                       #Change rownames to remove PCHAS_
                       rownames(heatmap_data) = str_split_fixed(rownames(heatmap_data), 
                                                                pattern = '_', 
                                                                n = 2)[,2]

                       if(use_times){
                         colnames(heatmap_data) <- times_stages_table$time[match(colnames(heatmap_data),
                                                                                 times_stages_table$r_stage_diff_count)]
                       }

                       if(use_order_vector){
                         #Order the columns by ordering_vector
                         heatmap_data <- heatmap_data[,order(match(colnames(heatmap_data), 
                                                                   order_vector)),
                                                      drop = FALSE]
                       }else{
                         heatmap_data <- heatmap_data[,order(colnames(heatmap_data)),
                                                      drop = FALSE] 
                       }

                       h = Heatmap(heatmap_data,
                                   name = paste0('z-score',
                                                 extra_title),
                                   col = c(cbPalette[6],
                                           'white', 
                                           cbPalette[7]),
                                   row_names_gp = gpar(fontsize = ifelse(
                                     nrow(heatmap_data) > 13, 
                                     12*(13/nrow(heatmap_data)), 
                                     12)
                                   ),
                                   column_names_gp = gpar(fontsize = 10),
                                   row_dend_width = unit(3, "cm"),
                                   right_annotation = if(tpm_mat){
                                     tpm_ha
                                   }else{
                                     NULL
                                   },
                                   left_annotation = 
                                     if(identical(list_element,
                                                  dplyr::first(zscore_list)) & is.character(annotation_clade)){
                                       clade_ha
                                     }else{
                                       NULL
                                     }, 
                                   #Only want the clade annotation when it is first heatmap
                                   width = unit(4.5, 'cm'),
                                   cluster_columns = FALSE,
                                   height = unit(7,'cm'),
                                   row_title = NULL,
                                   cluster_rows = FALSE,
                                   column_title = paste(title),
                                   heatmap_legend_param = list(
                                     at = c(3,2,1,0,-1,-2,-3),
                                     labels_gp = gpar(fontsize = 10),
                                     title_gp = gpar(fontsize = 10)
                                   ),
                                   ...
                       )
                       return(h)
                     })

  ht_list <- Reduce('+', heatmaps) #To add the list elements together

  if(tpm_mat){
    #Draw the legend
    tpm_lgd <- Legend(
      title = "Min/Max log(TPM + 1)", 
      col = tpm_col_fun, 
      at = tpm_breaks, 
      labels = tpm_breaks,
      title_gp = gpar(fontsize = 7)
    )
  }

  if(is.character(annotation_clade)){
    clade_lgd <- Legend(
      title = 'sub-family',
      at = names(annotation_col),
      legend_gp = gpar(fill = annotation_col)
    )
  }

  draw(ht_list, 
       row_title_side = 'left',
       column_title_side = 'bottom',
       heatmap_legend_side = 'left',
       #Only include the annotation legends for the annotations we actually have.
       annotation_legend_list = if(is.character(annotation_clade) & tpm_mat){
         list(tpm_lgd,
              clade_lgd)
       }else if(tpm_mat & !is.character(annotation_clade) ){
         list(tpm_lgd)
       } else if(!tpm_mat & is.character(annotation_clade) ){
         list(clade_lgd)
       })
}
```

## Phaseograms of the entire genome.

```{r phaseograms_whole_genome}
png(filename = 'tl19-07_phaseogram_rMTorder_allgenes_filter1TPM.png',
    units = 'cm',
    height = 20,
    width = 22,
    res = 330)
draw_heatmap_zscore(zscore_list = twentyfour_z,
                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
                    use_times = TRUE,
                    use_order_vector = FALSE,
                    show_row_names = FALSE,
                    tpm_mat = FALSE,
                    threshold_tpm = 1)
dev.off()
```

## Phaseograms of the _cir_ genes with clade annotations.

```{r pir_gene_phaseogram_breaks}
  max_tpm <- lapply(twentyfour_z, 
                    function(list_element){
                      #list_element <- zscore_list[[1]]
                      heatmap_data <- as.data.frame(list_element$zscores)
                      title <- list_element$experiment

                      if(!'Geneid' %in% colnames(heatmap_data)){
                        stop('No Geneid column in the heatmap data.frame')
                      }

                            heatmap_data <- as.data.frame(
                              filter(heatmap_data, 
                                     Geneid %in% cir_id)
                            )

                          rownames(heatmap_data) <- as.character(heatmap_data$Geneid)

                          heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
                          #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
                          # 
                          # if(!is.null(gene_ordering_vector)){
                          #   #Remove genes not in the main data frame but in the ordering vector, 
                          #   # and then order based on the vector
                          #   remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
                          #   heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
                          #   heatmap_data <- heatmap_data[na.omit(
                          #     match(gene_ordering_vector, 
                          #           rownames(heatmap_data))
                          #   ),,
                          #   drop=FALSE]
                          # }

                          geneids <- rownames(heatmap_data)

                          #Get max/min TPM scores for the side annotation heatmap                       
                          tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
                                                            list_element$tpm$Geneid),] %>% 
                            select(-Geneid) %>% 
                            as.matrix %>% 
                            log1p
                          rownames(tpm_mat) <- geneids

                          tpm_min <- rowMins(tpm_mat)
                          names(tpm_min) <- geneids
                          tpm_max <- rowMaxs(tpm_mat)
                          names(tpm_max) <- geneids
                          return(list(tpm_max = tpm_max,
                                      tpm_min = tpm_min))
                        })

  tpm_max_all <- mround(
    max(
      unlist(
        lapply(max_tpm, 
               getElement,
               'tpm_max')
      )
    ), 
    base = 1)

  tpm_breaks <- round(
    c(0,
      tpm_max_all/4,
      tpm_max_all/2,
      tpm_max_all*(3/4),
      tpm_max_all),
    digits = 2
  )
```

```{r phaseograms_cir_order_each}

#Generate the colour function from the TPM legend breaks for the TPM annotation

tpm_col_fun <- colorRamp2(
  breaks = tpm_breaks,
  colors = viridis(length(tpm_breaks))
)

#Apply a threshold to the TPM in order to remove low expressed genes.
# Remember to add log1p here! 
# Threshold uses the max value, so it ony needs to be above thresh 
#   in one stage avg in one transmission method.
#Will alter the gene_filter vector in order to only select the genes above threshold

gene_maxs <- bind_rows(
  lapply(max_tpm, 
         getElement,
         'tpm_max')
)

remove_genes <- names(gene_maxs)[apply(
  gene_maxs, 
  2, 
  function(maxs){
    all(maxs < log2(1+1))
  }
)]

filter_genes_vector <- cir_id[!cir_id %in% remove_genes]

heatmaps <- lapply(names(twentyfour_z), 
                   function(list_name){
                     list_element <- getElement(twentyfour_z, 
                                                list_name)
                     #list_name <- names(twentyfour_z)[[1]]
                     heatmap_data <- as.data.frame(list_element$zscores)
                     title <- list_element$experiment

                     heatmap_data <- as.data.frame(
                       filter(heatmap_data, 
                              Geneid %in% filter_genes_vector)
                     )

                     rownames(heatmap_data) <- as.character(heatmap_data$Geneid)

                     heatmap_data <- heatmap_data[, 
                                                  !colnames(heatmap_data) %in% 'Geneid', 
                                                  drop = TRUE]
                     #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!

                     #Remove genes not in the main data frame but in the ordering vector,
                     #  and then order based on the vector
                     gene_ordering_vector <- getElement(phaseogram_time_order,
                                                        list_name)$zscore_order$Geneid

                     order_cir <- gene_ordering_vector[gene_ordering_vector %in% filter_genes_vector]
                     clade_info <- getElement(cir_clade_info,
                                              'SorL')[match(order_cir, 
                                                            cir_clade_info$Geneid)]
                     names(clade_info) <- str_split_fixed(order_cir, 
                                                          pattern = '_', 
                                                          n = 2)[,2]
                     annotation_col <- SorL_col

                     remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
                     heatmap_data <- heatmap_data[remove_index,
                                                  ,
                                                  drop = FALSE]
                     heatmap_data <- heatmap_data[na.omit(
                       match(gene_ordering_vector, 
                             rownames(heatmap_data))
                     ),
                     ,
                     drop=FALSE]

                     #Get max/min TPM scores for the side annotation heatmap                       
                     tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
                                                       list_element$tpm$Geneid),] %>%
                       select(-Geneid) %>% 
                       as.matrix %>% 
                       log1p

                     tpm_min <- rowMins(tpm_mat)
                     tpm_max <- rowMaxs(tpm_mat)

                     names(tpm_min) = rownames(heatmap_data)
                     names(tpm_max) = rownames(heatmap_data)

                     tpm_ha <- rowAnnotation(
                       min.max = anno_simple(cbind(tpm_min, 
                                                   tpm_max), 
                                             col = tpm_col_fun,
                                             gp = gpar(col = 'black',
                                                       lwd = 0.1),
                                             border = TRUE,
                                             simple_anno_size = unit(0.2, "cm")),
                       show_annotation_name = FALSE)

                     #Clade annotation

                     clade_ha <- rowAnnotation(
                       'sub-family' = anno_simple(
                         clade_info,
                         col = annotation_col,
                       ),
                       annotation_name_gp = gpar(fontsize = 10)
                     )

                     #Change rownames to remove PCHAS_
                     rownames(heatmap_data) = str_split_fixed(rownames(heatmap_data), 
                                                              pattern = '_', 
                                                              n = 2)[,2]

                     colnames(heatmap_data) <- times_stages_table$time[match(colnames(heatmap_data),
                                                                             times_stages_table$r_stage_diff_count)]

                     heatmap_data <- heatmap_data[,order(colnames(heatmap_data)),
                                                  drop = FALSE]

                     h = Heatmap(heatmap_data,
                                 name = paste0('z-score'),
                                 col = c(cbPalette[6],
                                         'white', 
                                         cbPalette[7]),
                                 row_names_gp = gpar(fontsize = ifelse(
                                   nrow(heatmap_data) > 13, 
                                   12*(13/nrow(heatmap_data)), 
                                   12)
                                 ),
                                 column_names_gp = gpar(fontsize = 10),
                                 row_dend_width = unit(3, "cm"),
                                 right_annotation = tpm_ha,
                                 left_annotation = clade_ha,
                                 width = unit(4.5, 'cm'),
                                 cluster_columns = FALSE,
                                 height = unit(7,'cm'),
                                 row_title = NULL,
                                 cluster_rows = FALSE,
                                 column_title = paste(title),
                                 heatmap_legend_param = list(
                                   at = c(3,2,1,0,-1,-2,-3),
                                   labels_gp = gpar(fontsize = 10),
                                     title_gp = gpar(fontsize = 10)
                                 ),
                                 show_row_names = FALSE
                     )
                     return(h)
                   })

ht_list <- Reduce('+', heatmaps) #To add the list elements together

#Draw the legend
tpm_lgd <- Legend(
  title = "Min/Max log(TPM + 1)", 
  col = tpm_col_fun, 
  at = tpm_breaks, 
  labels = tpm_breaks,
  title_gp = gpar(fontsize = 7)
)

clade_lgd <- Legend(
  title = 'sub-family',
  at = names(SorL_col),
  legend_gp = gpar(fill = SorL_col)
)

png(filename = 'tl19-07_phaseogram_each-in-order_cir_clades.png',
    units = 'cm',
    height = 20,
    width = 22,
    res = 200)

  draw(ht_list, 
       row_title_side = 'left',
       column_title_side = 'bottom',
       heatmap_legend_side = 'left',
       annotation_legend_list = 
         list(tpm_lgd,
              clade_lgd),
       ht_gap = unit(.5,'cm')
       )

  dev.off()

```

```{r phaseograms_cir}
png(filename = 'tl19-07_phaseogram_rMTorder_cir_clades_filter1TPM.png',
    units = 'cm',
    height = 20,
    width = 22,
    res = 330)
draw_heatmap_zscore(zscore_list = twentyfour_z,
                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
                    filter_genes_vector = cir_id,
                    annotation_clade = 'clade',
                    annotation_col = clade_col,
                    threshold_tpm = 1,
                    use_times = TRUE,
                    use_order_vector = FALSE,
                    tpm_mat = FALSE)
dev.off()

png(filename = 'tl19-07_phaseogram_rMTorder_cir_SorL_filter1TPM.png',
    units = 'cm',
    height = 20,
    width = 22,
    res = 330)
draw_heatmap_zscore(zscore_list = twentyfour_z,
                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
                    filter_genes_vector = cir_id,
                    annotation_clade = 'SorL',
                    annotation_col = SorL_col,
                    threshold_tpm = 1,
                    use_times = TRUE,
                    use_order_vector = FALSE,
                    show_row_names = FALSE,
                    tpm_mat = FALSE)
dev.off()
draw_heatmap_zscore(zscore_list = twentyfour_z,
                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
                    filter_genes_vector = cir_id,
                    annotation_clade = 'SorL',
                    annotation_col = SorL_col,
                    threshold_tpm = 1,
                    use_times = TRUE,
                    use_order_vector = FALSE,
                    show_row_names = FALSE,
                    tpm_mat = FALSE)

png(filename = 'tl19-07_phaseogram_SBPorder_cir_SorL_filter1TPM.png',
    units = 'cm',
    height = 20,
    width = 22,
    res = 330)
draw_heatmap_zscore(zscore_list = twentyfour_z,
                    gene_ordering_vector = phaseogram_time_order$SBP$zscore_order$Geneid,
                    filter_genes_vector = cir_id,
                    annotation_clade = 'SorL',
                    annotation_col = SorL_col,
                    threshold_tpm = 1,
                    use_times = TRUE,
                    use_order_vector = FALSE,
                    show_row_names = FALSE,
                    tpm_breaks = tpm_breaks,
                    tpm_mat = FALSE)
dev.off()
draw_heatmap_zscore(zscore_list = twentyfour_z,
                    gene_ordering_vector = phaseogram_time_order$SBP$zscore_order$Geneid,
                    filter_genes_vector = cir_id,
                    annotation_clade = 'SorL',
                    annotation_col = SorL_col,
                    threshold_tpm = 1,
                    use_times = TRUE,
                    use_order_vector = FALSE,
                    show_row_names = FALSE,
                    tpm_breaks = tpm_breaks,
                    tpm_mat = FALSE)

```
## Figure 3b - Highest expressed all _pir_ heatmap

```{r highest_expressed_stage_heatmap_fig3b}

num_SorL <- cir_info %>% 
  group_by(SorL) %>% 
  summarise(num = n_distinct(Gene.ID))

labels_SorL <- setNames(paste0(num_SorL$SorL, ' (n = ', num_SorL$num, ')'),
                        num_SorL$SorL)

# Get order of pir genes by highest expressed genes in each sample.

combined_data_pir_ht_list <- lapply(names(twentyfour_avgtpm_noschz), 
       function(list_name){

         combined_data_pir <- getElement(twentyfour_avgtpm_noschz, 
                                     list_name)$tpm %>% 
           filter(Geneid %in% cir_id) %>% 
           column_to_rownames(var = 'Geneid')

         #Ordering that a more understandable order

         combined_data_pir <- combined_data_pir[,c(ordering(colnames(combined_data_pir)[-1], 
                                                            order_vector)),
                                                drop = FALSE]

         combined_data_ordered <- tibble(Geneid = names(apply(combined_data_pir, 1, mean)),
                                         means = apply(combined_data_pir, 1, mean),
                                         subfam = cir_clade_info$SorL[match(names(apply(combined_data_pir, 1, mean)), 
                                                                            cir_clade_info$Geneid)]) %>% 
           filter(means > 1) %>% 
           arrange(subfam, 
                   desc(means)) %>% 
           .$Geneid

         #Order the heatmap data by the order defined above (SorL, then mean expression).
         combined_data_pir_ht <- combined_data_pir[
           order(
             match(rownames(combined_data_pir),
                   combined_data_ordered),
             na.last = NA #Remove NAs
           ),
         ]

         #Change the colnames to the times
         colnames(combined_data_pir_ht) <- times_stages_table$time[match(colnames(combined_data_pir_ht),

                                                                         times_stages_table$r_stage_diff_count)]

         #Order by time point
         combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
                                                      drop = FALSE] 
         #Add a column for the SorL
         combined_data_pir_ht$SorL <- cir_clade_info$SorL[match(names(apply(combined_data_pir_ht, 1, mean)), 
                                                                cir_clade_info$Geneid)]

         return(combined_data_pir_ht)
       })
names(combined_data_pir_ht_list) <- names(twentyfour_avgtpm_noschz)

maxcol <- ceiling(
  max(
    unlist(
      lapply(combined_data_pir_ht_list,
             function(list_element){
               max(
                 log2(1+list_element[!colnames(list_element) %in% 'SorL'])
               )
             }
      )
    )
  )
)

lapply(names(combined_data_pir_ht_list),
       function(list_name_to_order){

         combined_data_pir_ht <- getElement(combined_data_pir_ht_list,
                                            list_name_to_order)

         pir_highest_tpm_order <- rownames(combined_data_pir_ht)

         #Sub-family annotation
         clade_info <- combined_data_pir_ht$SorL
         annotation_col <- SorL_col

         #Subfamily annotation

         clade_ha <- rowAnnotation(
           'sub-family' = anno_simple(
             clade_info,
             col = annotation_col,
           ),
           annotation_name_gp = gpar(fontsize = 10)
         )

         anc_ha <- rowAnnotation(foo = anno_mark(at =  na.omit(match(clade_info, 'ancestral')) , 
                                                 labels = 'ancestral',
                                                 labels_gp = gpar(fontsize = 7),
                                                 link_width = unit(3, "mm")))

         #lapply across the datasets now, for each order
         ht_list <- lapply(names(combined_data_pir_ht_list),
                           function(list_name){

                             combined_data_pir_ht <- getElement(twentyfour_avgtpm_noschz, 
                                                                list_name)$tpm %>% 
                               column_to_rownames(var = 'Geneid')

                             #Get the times instead of the stage names.
                             colnames(combined_data_pir_ht) <- times_stages_table$time[match(colnames(combined_data_pir_ht),

                                                                                             times_stages_table$r_stage_diff_count)]

                             combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
                                                                          drop = FALSE]

                             Heatmap(
                               log2(1+combined_data_pir_ht)[pir_highest_tpm_order,],
                               cluster_rows = FALSE,
                               cluster_columns = FALSE,
                               show_row_names = FALSE,
                               name = 'log(TPM+1)', 
                               col = colorRamp2(c(0,
                                                  maxcol/4,
                                                  maxcol/2,
                                                  3*maxcol/4,
                                                  maxcol),
                                                viridis(5)),
                               na_col = 'white', 
                               border = TRUE,
                               border_gp = gpar(lwd = 0.1),
                               column_names_gp = gpar(fontsize = 10),
                               row_dend_width = unit(3, "cm"),
                               width = unit(4.5, 'cm'),
                               height = unit(7,'cm'),
                               row_title = NULL,
                               left_annotation = if(
                                 list_name %in% dplyr::first(names(combined_data_pir_ht_list))
                               ){
                                 clade_ha
                               }else{
                                 NULL
                               },
                               right_annotation = if(
                                 list_name %in% dplyr::last(names(combined_data_pir_ht_list))
                               ){
                                 anc_ha
                               }else{
                                 NULL
                               },
                               # row_split =  cir_clade_info$SorL[match(pir_highest_tpm_order, 
                               #                                        cir_clade_info$Geneid)],
                               column_title = paste(list_name),
                               heatmap_legend_param = list(
                                 labels_gp = gpar(fontsize = 10),
                                 title_gp = gpar(fontsize = 10,
                                                 fontface = 'bold'),
                                 at = c(0,
                                        maxcol/4,
                                        maxcol/2,
                                        3*maxcol/4,
                                        maxcol)
                               )
                             )
                           }
         )
         ht_list <- Reduce(`+`, ht_list)
         clade_lgd <- Legend(
           title = 'sub-family',
           at = labels_SorL,
           legend_gp = gpar(fill = SorL_col)
         )

         png(filename = paste0('tl19-07_combined_pirs_ordered_highest_stage_expression_heatmap_TPM_',
                               list_name_to_order, 'order',
                               '.png'),
             width = 20,
             height = 10,
             units = 'cm',
             res = 330)

         draw(ht_list, 
              row_title_side = 'left',
              column_title_side = 'bottom',
              heatmap_legend_side = 'right',
              annotation_legend_side = 'right',
              annotation_legend_list = list(clade_lgd),
              ht_gap = unit(.35,'cm'),
              merge_legend = TRUE
         )

         dev.off()
         return(
           draw(ht_list, 
                row_title_side = 'left',
                column_title_side = 'bottom',
                heatmap_legend_side = 'right',
                annotation_legend_side = 'right',
                annotation_legend_list = list(clade_lgd),
                ht_gap = unit(.35,'cm'),
                merge_legend = TRUE
           ))
       })

```

# ```{r figure_3b_suppinfo}
# 
# supp_info_fig_3b <- lapply(names(combined_data_pir_ht_list),
#        function(list_name){
#          
#          combined_data_pir_ht <- getElement(twentyfour_avgtpm_noschz, 
#                                             list_name)$tpm %>% 
#            column_to_rownames(var = 'Geneid')
#          
#          #Get the times instead of the stage names.
#          colnames(combined_data_pir_ht) <- times_stages_table$time[match(colnames(combined_data_pir_ht),
#                                                                          
#                                                                          times_stages_table$r_stage_diff_count)]
#          
#          combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
#                                                       drop = FALSE] 
#          
#          colnames(combined_data_pir_ht) <- paste0(list_name, 
#                                                   '_', 
#                                                   colnames(combined_data_pir_ht))
#          
#          log2(1+combined_data_pir_ht)[pir_highest_tpm_order,]
#        }) %>% bind_cols
# 
# supp_info_fig_3b$SorL <- cir_clade_info$SorL[match(names(apply(supp_info_fig_3b, 1, mean)), 
#                                                                 cir_clade_info$Geneid)]
# 
# ```

# Lower pir expression vs rest of genome

```{r low_pir_expression}
set.seed(4534)

gene_expression_df <- twentyfour_ct_comb$tpm %>% 
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, 
                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
  filter(stage != 'sch.culture') %>% 
  mutate(pir_gene = ifelse(Geneid %in% cir_id,
                           'pir',
                           'non-pir')) %>% 
  group_by(stage, 
           pir_gene,
           transmission) %>% 
  slice_sample(n = length(cir_id)) %>%
  summarise(mean_tpm = mean(tpm),
            med_tpm = round(median(tpm),
                            digits = 6))

gene_expression_df %>% 
  group_by(transmission, 
           pir_gene) %>% 
  slice_min(mean_tpm, n =1)

gene_expression_df %>% 
  group_by(transmission, 
           pir_gene) %>% 
  slice_max(mean_tpm, n =1)

y <- ggplot(gene_expression_df,
       aes(x = pir_gene,
           y = log2(1+mean_tpm))) +
  facet_grid(~transmission) +
  ggtitle(expression(paste(italic('P. chabaudi'), ' 24h cycle'))) +
  xlab('Gene group')+
  ylab('log(Median TPM + 1)')+
  # geom_violin()+
  geom_dotplot(binaxis = 'y',
               stackdir = 'center',
               # binwidth = 1,
               # dotsize = 2,
               method = 'dotdensity') +
  geom_path(aes(group = stage,
                col = stage)) +
  theme_classic()
y

ggsave2('tl19-07_pir-v-nonpir_mean_TPM_barchart.pdf',
        plot = y,
        dpi = 300,
        width = 15,
        height = 10,
        units = 'cm')
```

```{r write_excel_data_sheets}

twentyfour_avgtpm_noschz_timecolnames <- lapply(names(twentyfour_avgtpm_noschz), 
                                                function(list_name){
                                                  list_element <- getElement(twentyfour_avgtpm_noschz, list_name)
                                                  colnames(list_element$tpm)[-1] <- times_stages_table$time[match(colnames(list_element$tpm)[-1],
                                                                                                              times_stages_table$r_stage_diff_count)]
                                                  list_element$tpm <- list_element$tpm[c('Geneid',sort(colnames(list_element$tpm)[-1]))]
                                                  return(list_element)
                                                })
names(twentyfour_avgtpm_noschz_timecolnames) <- names(twentyfour_avgtpm_noschz)

twentyfour_z_timecolnames <- lapply(names(twentyfour_z), 
                                    function(list_name){
                                      list_element <- getElement(twentyfour_z, list_name)
                                      colnames(list_element$zscores)[-1] <- times_stages_table$time[match(colnames(list_element$zscores)[-1],
                                                                                                          times_stages_table$r_stage_diff_count)]
                                      list_element$zscores <- list_element$zscores[c('Geneid',sort(colnames(list_element$zscores)[-1]))]
                                      return(list_element)
                                    })
names(twentyfour_z_timecolnames) <- names(twentyfour_z)

excel_data <- list(
  pir_gene_info = cir_info,
  chab_TPM_bioreps = twentyfour_ct_comb_noschz$tpm[,c('Geneid',sort(colnames(twentyfour_ct_comb_noschz$tpm)[-1]))] %>% 
    select(matches('Geneid|rMT')) %>% 
    dplyr::rename_with(~ str_remove_all(., pattern = 'rMT_'), everything()),
  chab_TPM_avg = twentyfour_avgtpm_noschz_timecolnames$rMT$tpm,
  # chab_TPM_avg_SBP = twentyfour_avgtpm_noschz_timecolnames$SBP$tpm,
  # chab_logTPM_SorL_rMTorder = supp_info_fig_3b,
  chab_TPM_ChaplAapl = select(cir_plot_chapl_aapl_supp_info, c(Geneid, time, transmission, chapl_aapl, sample, tpm)) %>% 
    filter(transmission == 'rMT') %>% 
    select(-transmission) %>% 
    filter(time != 'sch.culture') %>% 
    mutate(sample = str_remove_all(sample, pattern = 'rMT_')) %>% 
    arrange(Geneid, time),
  # chab_TPM_SorLSums = tpm_data_SorL,
  # chab_TPM_SubfamSums = tpm_data_clade,
  # chab_TPM_ChromSums = tpm_data_chrom,
  chab_zscore = twentyfour_z_timecolnames$rMT$zscores
  # chab_zscore_SBP = twentyfour_z_timecolnames$SBP$zscores
)

write_xlsx(excel_data,
           path = 'Manuscript_SuppInfo_chabaudi24h.xlsx')

```