### 2022-07-27 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_phaseGenes.Rmd
Start moving phase/GO enrichment code to its own Rmd file

---
title: "Manuscript_Github_Pchabaudi_phaseGene"
author: "Timothy Little"
date: "27/07/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      results = TRUE, 
                      message = FALSE, 
                      warning = FALSE)

knitr::opts_knit$set(root.dir = 
                       ".")
#Note that the above doesn't give an error if the dir doesn't exist

options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
```

```{r loading_packages, include = FALSE, message = FALSE}
# Load the required packages

library('readr')
library('dplyr')
library('ggplot2')
library("RColorBrewer")
library('data.table')
library('tibble')
library('stringr')
library('kableExtra')
library('viridis')
library('reshape2')
library('ComplexHeatmap')
library('circlize') # for circular plots and colour
library("factoextra")
library("MatrixGenerics") # for rowMins and rowMax
library("topGO")
library("FoldGO")

#Colour blind friendly palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

#Want the defaults for arrange, mutate and summarise to be from dplyr
arrange <- dplyr::arrange
mutate <- dplyr::mutate
summarise <- dplyr::summarise
select <- dplyr::select

#Global save var - set to TRUE if you want to save figures.
global_save_var <- FALSE
```

```{r sourcing_functions}
source('../rfunctions/uniquify_fun.R')
source('../rfunctions/ordering_fun.R')
source('../rfunctions/tpm_fun.R')
source('../rfunctions/usefulFunctions_mround_my-max_firstup_fun.R')
source('../rfunctions/draw_heatmap_zscore.R')
```

```{r ordering_vec}
order_vector <- c("Early.Rings",
                  "Early.Mid.Rings",
                  "Late.Mid.Rings",
                  "Late.Rings",
                  "Rings.Troph.conversion",
                  "Early.Trophs",
                  "Mid.Trophs",
                  "Late.Trophs")
```

The _cir_ gene names were downloaded using a search of PlasmoDB _P. chabaudi_ genome. _cir_ genes are the names of _pir_ genes in _P chabaudi_.

```{r cir_id, message=FALSE, results='hide'}

# Getting the cir gene names and transcript lengths needed for calculating the transcript-per-million

chabaudi_gene_info <- read.csv('../standard_files/PlasmoDB-47_beta_Pchabaudichabaudi_GenesByTaxon_Summary.csv',
                              header = TRUE,
                              stringsAsFactors = FALSE)

chabaudi_transcript_lengths <- select(chabaudi_gene_info, c(Gene.ID, Transcript.Length)) %>% 
  arrange(Gene.ID)

chabaudi_gene_info <- dplyr::rename(chabaudi_gene_info, 
                             Geneid = Gene.ID,
                             Description = Product.Description)

```

```{r cir_info}

cir_info <- read_delim('../standard_files/Pchabaudi_pir_info.txt',
           delim = '\t')
cir_id <- cir_info$Geneid

cir_clade_info <- read_delim('../standard_files/Pchabaudi_pir_clades.txt',
            delim = '\t')
```

```{r times_stages_labels}
#Use the RMT stage differential count in order to name the samples, but keep time to retain order
times_stages_table <- data.frame(time = c('02h', '05h', '08h', '11h', '14h', '17h','20h','23h'),
                                 stage_diff_count = c('Late-Rings',
                                                      'Rings-Troph-conversion',
                                                      'Early-Trophs',
                                                      'Mid-Trophs',
                                                      'Late-Trophs',
                                                      'Early-Rings',
                                                      'Early-Mid-Rings',
                                                      'Late-Mid-Rings'))
# R will convert the dashes to '.' at some point so let's have this notation here also
times_stages_table$r_stage_diff_count <- str_replace_all(times_stages_table$stage_diff_count, 
                                                         pattern = '-', 
                                                         replacement = '\\.')
```

```{r read_TPM}
#Need to use the calculateTPM script first.

chab_tpm.df <- read_csv('data/Pchabaudi_TPM_all.csv')

chab_tpm_avg.df <- read_csv('data/Pchabaudi_TPM_average.csv')

chab_tpm_avg.list <- list(
  'RMT' = list(
    tpm = chab_tpm_avg.df[,grepl(colnames(chab_tpm_avg.df), 
                                pattern = 'rMT|Geneid')],
    experiment = 'RMT'
  ),
  'SBP' = list(
    tpm = chab_tpm_avg.df[,grepl(colnames(chab_tpm_avg.df), 
                                pattern = 'SBP|Geneid')],
    experiment = 'SBP'
  )
)

```

```{r import_z}
chab_z.df <- inner_join(chab_z.list[[1]]$zscores, 
                        chab_z.list[[2]]$zscores, 
                        by = 'Geneid')
chab_z.df <- read.csv(file = 'data/Pchabaudi_z.csv')

chab_z.list <- list('rMT' = list('zscores' = chab_z.df[,grepl(colnames(chab_z.df),
                                                                             pattern = 'Geneid|rMT')]),
                    'SBP' = list('zscores' = chab_z.df[,grepl(colnames(chab_z.df),
                                                                             pattern = 'Geneid|SBP')]))

order_z_rMT <- read.table(file = 'data/Pchabaudi_z_phaseOrder_rMT.list')[,1]
order_z_SBP <- read.table(file = 'data/Pchabaudi_z_phaseOrder_SBP.list')[,1]

```

```{r colours_clades_phase}
SorL_col <- setNames(
  c(    '#F0E442',
        '#56B4E9', #L
        '#E69F00', #S
        '#999999'),
  sort(
    unique(cir_info$SorL)
  )
)

SorL_col_num <- SorL_col
names(SorL_col_num) <- sapply(names(SorL_col), 
                              function(clade_name) {
                                paste0(
                                  clade_name, 
                                  ' (n = ',
                                  sum(cir_info$SorL == clade_name),
                                  ')'
                                )
                              })

clade_col <- setNames(
  c('#F0E442',
    '#56B4E9', #L1
    '#009E73',
    '#0072B2',
    '#E69F00', #S1
    '#D55E00',
    '#CC79A7',
    '#999999'),
  sort(
    unique(cir_info$subfam)
  )
)

clade_col_num <- clade_col
names(clade_col_num) <- sapply(names(clade_col), 
                               function(clade_name) {
                                 paste0(
                                   clade_name, 
                                   ' (n = ',
                                   sum(cir_info$subfam == clade_name),
                                   ')'
                                 )
                               })

```

```{r gene_sync}

#Which genes are out of sync between rMT and SBP?

#Melt the z-score dataframe into long form, then slice out the maximum z-score value.
# Filter out any gene which has multiple max values for simplicity.
# Summarise this with the numbers that each stage (the max stage) corresponds to.

z_max <- lapply(chab_z.list, function(x) {
  x$zscores %>% 
  reshape2::melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'zscore') %>% 
  group_by(Geneid) %>% 
  slice_max(order_by = zscore, n = 1) %>% 
  filter(n() == 1) %>% 
  summarise(sample_max = times_stages_table$num[
    str_detect(sample, 
               times_stages_table$time)
    ])
})

#Bind the two together with inner_join so that it only includes genes in both.
# Then calculate the peak difference by subtracting the difference between the sample numbers.
# Since this is a cycle, no point of the cycle is theoretically 
#  more than 4 time points away from the next one. 
#  e.g. 23h and 02h (8 and 1) are only 1 time point apart, not seven.
#  Hence we subtract values above four from eight
z_max_diff <- inner_join(z_max[[1]], z_max[[2]], by = 'Geneid') %>% 
  mutate(diff = ifelse(abs(sample_max.x - sample_max.y) > 4,
                       8 - abs(sample_max.x - sample_max.y),
                       abs(sample_max.x - sample_max.y)))

#Probably unnecessarily complicated code to make a barplot of how many genes 
# are peaking at different timepoints between rMT and SBP
barplot(group_by(z_max_diff, diff) %>% summarise(total = n()) %>% .$total,
        names.arg = paste0(sort(unique(z_max_diff$diff))*3, 'h'))

#And for the pir genes
barplot(group_by(z_max_diff, diff) %>% 
          filter(Geneid %in% cir_id) %>% 
          summarise(total = n()) %>% 
          .$total,
        names.arg = paste0(sort(unique(z_max_diff$diff))*3, 'h'))

```

```{r phase_go_enrich}
#Get the gene list for each group of genes in and out of phase.
gene.list <- chabaudi_gene_info$Geneid
z_max_diff.list <- lapply(seq(0,4), 
                          function(diff_num) {
                            diff.list <- filter(z_max_diff, diff == diff_num) %>% 
                              .$Geneid
                            #Get 1 for on the list and 0 for not.
                            gene_diff.list <- as.integer(gene.list %in% diff.list)
                            names(gene_diff.list) <- gene.list
                            return(gene_diff.list)
                          })
names(z_max_diff.list) <- seq(0,4)

#Read in the GO-Annotation file using GAFReader from FoldGO.
# getAnnotation to obtain a list of the GO terms and associated genes.
# inverseList (from topGO) to turn this into list of geneID with associated GO terms.
gaf <- GAFReader("../standard_files/PlasmoDB-48_Pchabaudichabaudi_GO.gaf")
go_annotation <- inverseList(getAnnotation(gaf))

# How many genes for each group do not have any GO annotations?

lapply(z_max_diff.list,
       function(gene_diff.list){
         total_noGO <- sum(!names(gene_diff.list[gene_diff.list == 1]) %in% 
                             names(go_annotation))
         prop_noGO <- total_noGO/sum(gene_diff.list == 1)
         return(paste0(total_noGO, ' genes without annotation = ',
                       round(prop_noGO*100, digits = 2), '%'))
       })

```

```{r}
#Load for the showSigOfNodes network function.
library('Rgraphviz')

#Make a topGOdata object 
GOdata <- new("topGOdata", 
              ontology = "BP", 
              geneSel = function(x){x == 1},
              allGenes = z_max_diff.list$`1`,
              nodeSize = 5,
              annot = annFUN.gene2GO, 
              gene2GO = go_annotation)
resultFis <- runTest(GOdata, 
                     algorithm = "classic", 
                     statistic = "fisher")
plot.new()
hist(score(resultFis), 50, xlab = "p-values")
dev.off()
showSigOfNodes(GOdata,
               score(resultFis),
               firstSigNodes = 5,
               useInfo = 'all')
goRes <- GenTable(GOdata, 
         classic = resultFis, 
         orderBy = "classic", 
         ranksOf = "classic",
         topNodes = 25) %>% 
  filter(classic < 0.05)

#Get first (most significant) enriched GO id 
goID <- goRes[4, "GO.ID"]
#Show distribution of GO term genes compared to null distribution (all other genes)
print(showGroupDensity(GOdata, goID, ranks = TRUE))

#Unlist the lapply below which goes through each element of the list and 
# returns TRUE is the go_id is listed for that gene.
#This makes a TRUE/FALSE index for the list/names of the list which can get 
# the genes associated with that GO id.
goID_genes <- names(go_annotation)[unlist(
  lapply(go_annotation, 
         function(go_ids){
           any(goID %in% go_ids)
         })
)]

draw_heatmap_zscore(zscore_list = chab_z.list,
                    filter_genes_vector = names(z_max_diff.list$`1`[z_max_diff.list$`2` == 1]),
                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
                    threshold_tpm = 1,
                    use_times = TRUE,
                    use_order_vector = FALSE,
                    show_row_names = FALSE,
                    tpm_mat = TRUE)
```

```{r}

lapply(seq(0,4), 
       function(phase){
         GOdata <- new("topGOdata", 
                       ontology = "BP", 
                       geneSel = function(x){x == 1},
                       allGenes = getElement(z_max_diff.list, paste(phase)),
                       nodeSize = 5,
                       annot = annFUN.gene2GO, 
                       gene2GO = go_annotation)
         resultFis <- runTest(GOdata, 
                              algorithm = "classic", 
                              statistic = "fisher")
         draw_heatmap_zscore(zscore_list = chab_z.list,
                             filter_genes_vector = names(
                               getElement(z_max_diff.list, paste(phase))[
                                 getElement(z_max_diff.list, paste(phase)) == 1
                               ]
                             ),
                             gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
                             threshold_tpm = 1,
                             use_times = TRUE,
                    use_order_vector = FALSE,
                    show_row_names = FALSE,
                    tpm_mat = TRUE)
         GenTable(GOdata, 
                  classic = resultFis, 
                  orderBy = "classic", 
                  ranksOf = "classic",
                  topNodes = 25) %>% 
           filter(classic < 0.05)
       })

```