### 2022-08-28 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_qusage.Rmd
Look at L4s ChAPL and AAPLs separate

---
title: "Manuscript_Github_Pchabaudi_qusage"
author: "Timothy Little"
date: "29/07/2022"
output: html_document
---

Qusage analysis of gene set enrichment (e.g. pirs, L-forms, ChAPL loci) between the transmission methods and timepoints of the asexual life cycle.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      results = TRUE, 
                      message = FALSE, 
                      warning = FALSE)

knitr::opts_knit$set(root.dir = 
                       ".")
#Note that the above doesn't give an error if the dir doesn't exist

options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)

#Set random seed for reproducible randomisation
set.seed(34576)
```

```{r loading_packages, include = FALSE, message = FALSE}
# Load the required packages

library('readr')
library('dplyr')
library('ggplot2')
library("RColorBrewer")
library('data.table')
library('tibble')
library('stringr')
library('kableExtra')
library('viridis')
library('reshape2')
library('ComplexHeatmap')
library('circlize') # for circular plots and colour
library("factoextra")
library("MatrixGenerics") # for rowMins and rowMax
library('qusage')

#Colour blind friendly palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

#Want the defaults for arrange, mutate and summarise to be from dplyr
arrange <- dplyr::arrange
mutate <- dplyr::mutate
summarise <- dplyr::summarise
select <- dplyr::select

#Global save var - set to TRUE if you want to save figures.
global_save_var <- FALSE
```

```{r sourcing_functions}
source('../rfunctions/uniquify_fun.R')
source('../rfunctions/ordering_fun.R')
source('../rfunctions/tpm_fun.R')
source('../rfunctions/usefulFunctions_mround_my-max_firstup_fun.R')
source('../rfunctions/draw_heatmap_zscore.R')
```

```{r ordering_vec}
order_vector <- c("Early.Rings",
                  "Early.Mid.Rings",
                  "Late.Mid.Rings",
                  "Late.Rings",
                  "Rings.Troph.conversion",
                  "Early.Trophs",
                  "Mid.Trophs",
                  "Late.Trophs")
```

```{r cir_id, message=FALSE, results='hide'}
# The _cir_ gene names were downloaded using a search of PlasmoDB _P. chabaudi_ genome. _cir_ genes are the names of _pir_ genes in _P chabaudi_.
# Getting the cir gene names and transcript lengths needed for calculating the transcript-per-million

chabaudi_gene_info <- read.csv('../standard_files/PlasmoDB-47_beta_Pchabaudichabaudi_GenesByTaxon_Summary.csv',
                              header = TRUE,
                              stringsAsFactors = FALSE)

chabaudi_transcript_lengths <- select(chabaudi_gene_info, c(Gene.ID, Transcript.Length)) %>% 
  arrange(Gene.ID)

chabaudi_gene_info <- dplyr::rename(chabaudi_gene_info, 
                             Geneid = Gene.ID,
                             Description = Product.Description)

```

```{r cir_info}

cir_info <- read_delim('../standard_files/Pchabaudi_pir_info.txt',
           delim = '\t')
cir_id <- cir_info$Geneid

cir_clade_info <- read_delim('../standard_files/Pchabaudi_pir_clades.txt',
            delim = '\t')
```

```{r times_stages_labels}
#Use the RMT stage differential count in order to name the samples, but keep time to retain order
times_stages_table <- data.frame(time = c('02h', '05h', '08h', '11h', '14h', '17h','20h','23h'),
                                 stage_diff_count = c('Late-Rings',
                                                      'Rings-Troph-conversion',
                                                      'Early-Trophs',
                                                      'Mid-Trophs',
                                                      'Late-Trophs',
                                                      'Early-Rings',
                                                      'Early-Mid-Rings',
                                                      'Late-Mid-Rings'))
# R will convert the dashes to '.' at some point so let's have this notation here also
times_stages_table$r_stage_diff_count <- str_replace_all(times_stages_table$stage_diff_count, 
                                                         pattern = '-', 
                                                         replacement = '\\.')
#Add a number from 1 to 8 for each stage, useful for phase calculations
times_stages_table$num <- 1:8
```

```{r colours_clades_phase}
SorL_col <- setNames(
  c(    '#F0E442',
        '#56B4E9', #L
        '#E69F00', #S
        '#999999'),
  sort(
    unique(cir_info$SorL)
  )
)

SorL_col_num <- SorL_col
names(SorL_col_num) <- sapply(names(SorL_col), 
                              function(clade_name) {
                                paste0(
                                  clade_name, 
                                  ' (n = ',
                                  sum(cir_info$SorL == clade_name),
                                  ')'
                                )
                              })

clade_col <- setNames(
  c('#F0E442',
    '#56B4E9', #L1
    '#009E73',
    '#0072B2',
    '#E69F00', #S1
    '#D55E00',
    '#CC79A7',
    '#999999'),
  sort(
    unique(cir_info$subfam)
  )
)

clade_col_num <- clade_col
names(clade_col_num) <- sapply(names(clade_col), 
                               function(clade_name) {
                                 paste0(
                                   clade_name, 
                                   ' (n = ',
                                   sum(cir_info$subfam == clade_name),
                                   ')'
                                 )
                               })

chapl_aapl_name_col <- setNames(
  c('#F0E442',
    '#56B4E9', #L1
    '#009E73',
    '#0072B2',
    '#E69F00', #S1
    '#D55E00',
    '#CC79A7',
    '#999999'),
  sort(
    unique(cir_info$chapl_aapl_name)
  )
)

chapl_aapl_name_col_num <- chapl_aapl_name_col
names(chapl_aapl_name_col_num) <- sapply(
  names(chapl_aapl_name_col), 
  function(clade_name) {
    paste0(
      clade_name, 
      ' (n = ',
      sum(cir_info$chapl_aapl_name == clade_name, 
          na.rm = TRUE),
      ')'
    )
  })
```

```{r read_TPM}
#Need to use the calculateTPM script first.

chab_tpm.df <- read_csv('data/Pchabaudi_TPM_all.csv')

chab_tpm_avg.df <- read_csv('data/Pchabaudi_TPM_average.csv')

# chab_tpm_avg.list <- list(
#   'RMT' = list(
#     tpm = chab_tpm_avg.df[,grepl(colnames(chab_tpm_avg.df), 
#                                 pattern = 'rMT|Geneid')],
#     experiment = 'RMT'
#   ),
#   'SBP' = list(
#     tpm = chab_tpm_avg.df[,grepl(colnames(chab_tpm_avg.df), 
#                                 pattern = 'SBP|Geneid')],
#     experiment = 'SBP'
#   )
# )

```

```{r df}
#Filter low expressed genes (greater than 10 tpm in at least one sample), convert to a matrix, add 1 for the log transformation, then transform to log2
chab_tpm.df <-column_to_rownames(chab_tpm.df, var = 'Geneid')
chab_tpm.df <- chab_tpm.df[apply(chab_tpm.df, 1, function(x) {any(x >= 10)}),]
chab_tpm_log2plus1.mat <- chab_tpm.df %>% 
  as.matrix %>% 
  `+`(.+1) %>% 
  log(base = 2)
chab_tpm_log2plus1.mat[1:5,1:5]
```

```{r generate_labels}
#Use make.names to remove the '-' from the labels
colnames(chab_tpm_log2plus1.mat) <- make.names(colnames(chab_tpm_log2plus1.mat))
```

Removing genesets with < 3 members.

```{r generate_genesets}
pir_geneset <- list(pirs = cir_info$Geneid)

subfam_list <- lapply(unique(cir_info$subfam), 
                      function(clade_nam) filter(cir_info, 
                                                 subfam == clade_nam)$Geneid)
names(subfam_list) <- unique(cir_info$subfam)
subfam_list <- subfam_list[unlist(lapply(subfam_list, 
                                         function(x) length(x) > 3))]

cir_info <- mutate(cir_info, 
                   subfam_loci = paste(subfam,
                                       chapl_aapl, 
                                       sep = '_'))
L4_loci_list <- lapply(c('L4_ChAPL', 'L4_AAPL'), 
                  function(clade_nam) filter(cir_info, 
                                             subfam_loci == clade_nam)$Geneid)
names(L4_loci_list) <- c('L4_ChAPL', 'L4_AAPL')
L4_loci_list <- L4_loci_list[unlist(lapply(L4_loci_list, 
                                         function(x) length(x) > 3))]

ShortLong_list <- 
  lapply(unique(cir_info$SorL), 
         function(SorL_nam) filter(cir_info, 
                                   SorL == SorL_nam)$Geneid)
names(ShortLong_list) <- unique(cir_info$SorL)
ShortLong_list <- ShortLong_list[unlist(lapply(ShortLong_list, 
                                               function(x) length(x) > 3))]
order_path.names <- ordering(c('pirs', unique(cir_info$SorL), sort(unique(cir_info$subfam))), 
                             order_vector = c('pir','L','S'))

locus_list <- 
  lapply(unique(cir_info$chapl_aapl_name), 
         function(chapl_aapl_name_nam) filter(cir_info, 
                                   chapl_aapl_name == chapl_aapl_name_nam)$Geneid)
names(locus_list) <- unique(cir_info$chapl_aapl_name)
locus_list <- locus_list[unlist(lapply(locus_list, 
                                               function(x) length(x) > 3))]

```

Compare the gene lists between the same stages in different transmission

```{r qusage_stagesBetweenTransmission}

comparisons_stagesTrans <- paste0(paste('rMT', 
                                        times_stages_table$time, 
                                        sep = '_'),
                                  '-',
                                  paste('SBP', 
                                        times_stages_table$time, 
                                        sep = '_'))

comp_stagesTrans.mat <- chab_tpm_log2plus1.mat[
  ,!grepl(colnames(chab_tpm_log2plus1.mat), pattern = 'sch.culture')
]
# For the life cycle labels then extract the transmission + timepoint label
lifeCycle_labels <- str_extract(colnames(comp_stagesTrans.mat),
                                pattern = '^(rMT|SBP)_([[:digit:]]{2}h|sch.culture)')

all_comparisons <- lapply(comparisons_stagesTrans, 
                          function(comparison){
                            print(comparison)
                            list_comp <- lapply(
                              list(pir_geneset,
                                   ShortLong_list,
                                   subfam_list,
                                   L4_loci_list,
                                   locus_list), 
                              function(gene_list){
                                qs.results.subfam <- qusage(comp_stagesTrans.mat,
                                                            lifeCycle_labels,
                                                            comparison,
                                                            gene_list,
                                                            n.points = 2^18)
                                list(results = qs.results.subfam, 
                                     qstable = qsTable(qs.results.subfam) %>% 
                                       mutate(contrast = qs.results.subfam$contrast))
                              })
                            names(list_comp) <- c('pir', 'shortlong', 'subfam', 'locus')
                            list_comp$combined <- bind_rows(lapply(list_comp, 
                                                                   getElement, 
                                                                   'qstable'))
                            return(list_comp)
                          })
names(all_comparisons) <- comparisons_stagesTrans

all_comparisons.df <- bind_rows(lapply(all_comparisons, 
                                       function(x) bind_rows(lapply(x, 
                                                                    function(y) y$qstable)))) %>% 
  arrange(pathway.name)

stagesTrans_ht.mat <- filter(all_comparisons.df, FDR < 0.05) %>% 
  mutate(pathway.name = factor(pathway.name, 
                               levels = ordering(pathway.name, 
                                                 order_vector = c('pirs', 
                                                                  'L', 
                                                                  'S', 
                                                                  'AAPL', 
                                                                  'ChAPL')))) %>% 
  arrange(pathway.name) %>% 
  melt(id.vars = c('contrast', 'pathway.name'),
       measure.vars  = c('log.fold.change')) %>% 
  dcast(contrast ~ pathway.name, 
        value.var = 'value') %>% 
  column_to_rownames('contrast') %>% 
  as.matrix

maxcol <- my_max(stagesTrans_ht.mat)
#Multiply by minus one to use my_max to get the most negative value
mincol <- my_max(stagesTrans_ht.mat*-1)*-1

#Transpose the matrix so that the stages (comparison) are on the x axis like other figures.
ht <- Heatmap(t(stagesTrans_ht.mat),
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        cluster_column_slices = FALSE,
        na_col = 'white',
        #Use factors to keep the ordering that we want.
        row_split = factor(str_extract(colnames(stagesTrans_ht.mat),
                                   pattern = 'pirs|^L|^S|AAPL|ChAPL'),
                              levels = unique(str_extract(colnames(stagesTrans_ht.mat),
                                   pattern = 'pirs|^L|^S|AAPL|ChAPL'))),
        row_order = colnames(stagesTrans_ht.mat),
        row_gap = unit(.5,"cm"),
        row_names_gp = gpar(fontsize = 9),
        border = TRUE,
        #Using this colour scheme because mincol is a lot less negative 
        # than maxcol is positive
        col = colorRamp2(c(mincol,
                           0,
                           maxcol/4,
                           maxcol/2,
                           maxcol),
                         c('dodgerblue1', 'white', 'yellow', 'orange', 'red')),
        heatmap_legend_param = list(title = 'LogFoldChange'))

if(global_save_var){
  pdf(file = 'plots/stageTrans_logFoldheatmap.pdf',
      height = 8,
      width = 6)
  draw(ht)
  dev.off()
}
```

Compare between stages within the same transmission method.

```{r qusage_transStage}
ordered_stages <- ordering(unique(lifeCycle_labels),
         order_vector = times_stages_table$time)
#Remove sch.culture from the comparisons for now.
ordered_stages <- ordered_stages[!ordered_stages == 'sch.culture']
comparisons_transStage.list <- lapply(c('rMT','SBP'), function(transmission){
  #Getting the comparisons across the time points for each transmission
  ordered_stages_trans <- str_subset(ordered_stages,
                               pattern = transmission)
  comparisons <- c(
    unlist(
      lapply(seq_along(ordered_stages_trans),
             function(i){ 
               if(i < length(ordered_stages_trans)){ 
                 paste0(ordered_stages_trans[i+1],'-',ordered_stages_trans[i]) 
               } else {
                 #When at the end, want it to cycle back round
                 paste0(ordered_stages_trans[1],'-',ordered_stages_trans[i]) 
               } 
             }
      )
    )
  )
})
names(comparisons_transStage.list) <- c('rMT', 'SBP')

# For the life cycle labels then extract the transmission + timepoint label
lifeCycle_labels <- str_extract(colnames(comp_stagesTrans.mat),
                                pattern = '^(rMT|SBP)_([[:digit:]]{2}h|sch.culture)')

all_comparisons_transStage <- 
  lapply(c('rMT', 'SBP'), 
         function(transmission){
           #Extract the comparisons for this transmission method
           comparisons_transStage <- comparisons_transStage.list[[transmission]]
           #Subset the dataset based on the transmission and removing sch.cultures
           comp_transStage.mat <- chab_tpm_log2plus1.mat[
             ,str_detect(colnames(chab_tpm_log2plus1.mat), pattern = paste0(transmission,'_(?!sch.culture)'))
           ]
           #Get appropriate life cycle labels from this.
           lifeCycle_labels <- str_extract(colnames(comp_transStage.mat),
                                pattern = '^(rMT|SBP)_([[:digit:]]{2}h|sch.culture)')
           results_transStage.list <- lapply(comparisons_transStage, 
                  function(comparison){
                    print(comparison)
                    list_comp <- lapply(
                      list(pir_geneset,
                           ShortLong_list,
                           subfam_list,
                           L4_loci_list,
                           locus_list), 
                      function(gene_list){
                        qs.results.subfam <- qusage(comp_transStage.mat,
                                                    lifeCycle_labels,
                                                    comparison,
                                                    gene_list,
                                                    n.points = 2^18)
                        list(results = qs.results.subfam, 
                             qstable = qsTable(qs.results.subfam) %>% 
                               mutate(contrast = qs.results.subfam$contrast))
                      })
                    names(list_comp) <- c('pir', 'shortlong', 'subfam', 'locus')
                    list_comp$combined <- bind_rows(lapply(list_comp, 
                                                           getElement, 
                                                           'qstable'))
                    return(list_comp)
                  })
           names(results_transStage.list) <- comparisons_transStage
           return(results_transStage.list)
         })
names(all_comparisons_transStage) <- c('rMT', 'SBP')

all_comparisons_transStage.df <- bind_rows(lapply(all_comparisons_transStage, 
                                           function(z) bind_rows(
                                             lapply(z, 
                                                    function(x) bind_rows(
                                                      lapply(x, 
                                                             function(y) y$qstable)))))) %>% 
  arrange(pathway.name)

```

```{r heatmap_df_transStage}
transStages_ht.mat <- filter(all_comparisons_transStage.df, FDR < 0.05) %>% 
  mutate(pathway.name = factor(pathway.name, 
                               levels = ordering(pathway.name, 
                                                 order_vector = c('pirs', 
                                                                  'L', 
                                                                  'S', 
                                                                  'AAPL', 
                                                                  'ChAPL')))) %>% 
  arrange(pathway.name) %>% 
  melt(id.vars = c('contrast', 'pathway.name'),
       measure.vars  = c('log.fold.change')) %>% 
  dcast(contrast ~ pathway.name, 
        value.var = 'value') %>% 
  column_to_rownames('contrast') %>% 
  as.matrix

maxcol <- my_max(transStages_ht.mat)
#Multiply by minus one to use my_max to get the most negative value
mincol <- my_max(transStages_ht.mat*-1)*-1

#Transpose the matrix so that the stages (comparison) are on the x axis like other figures.
# Change the colnames of the transposed matrix so that they're just the times.
# Will split the heatmap by rMT and SBP when drawing the heatmap.
transStages_ht.mat.t <- t(transStages_ht.mat)
colnames(transStages_ht.mat.t) <- str_remove_all(colnames(transStages_ht.mat.t),
                                                 pattern = '(rMT|SBP)_')

ht <- Heatmap(transStages_ht.mat.t,
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        cluster_column_slices = FALSE,
        na_col = 'white',
        #Use factors to keep the ordering that we want.
        row_split = factor(str_extract(colnames(transStages_ht.mat),
                                   pattern = 'pirs|^L|^S|AAPL|ChAPL'),
                              levels = unique(str_extract(colnames(transStages_ht.mat),
                                   pattern = 'pirs|^L|^S|AAPL|ChAPL'))),
        row_order = colnames(transStages_ht.mat),
        row_gap = unit(.5,"cm"),
        row_names_gp = gpar(fontsize = 9),
        column_split = factor(str_extract(rownames(transStages_ht.mat),
                                   pattern = 'rMT|SBP'),
                              levels = unique(str_extract(rownames(transStages_ht.mat),
                                   pattern = 'rMT|SBP'))),
        column_gap = unit(.3, "cm"),
        border = TRUE,
        #Using this colour scheme because mincol is a lot less negative 
        # than maxcol is positive
        col = colorRamp2(c(mincol,
                           0,
                           maxcol/4,
                           maxcol/2,
                           maxcol),
                         c('blue', 'white', 'yellow', 'orange', 'red')),
        heatmap_legend_param = list(title = 'LogFoldChange'))
if(global_save_var){
  pdf(file = 'plots/stagesWithinTrans_logFoldheatmap.pdf',
      height = 8,
      width = 6)
  draw(ht)
  dev.off()
}
```