### 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_VvxDeNovoAssemblyAnalysis_master.Rmd

---
title: "tl22-01_VivaxPirsInTheDeNovoAssemblies Master"
output: 
  html_document:
    theme: cosmo
    code_download: true
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
params:
  assembly_name: 'SRX2884276_SRR5646671_scspades_0.95cdhitest'
  assembly_file: '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/cdhitest/Rangel20_PvTCF/Rangel20_PvTCF_0.95_cdhitest.fasta'
  orf_file: '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/Rangel20_PvTCF_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir.fasta'
  blast_output_file: '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/blast/blast_to_Proteins/scspades/Rangel20_PvTCF/Rangel20_PvTCF_scspades_0.95cdhitest_hmmerPfam-Plasmodium_Vir_blast-VvxProt.outfmt6'
  new_table: FALSE
---

# Intro

This script analyses the results of a _de novo_ assembly of Plasmodium vivax RNAseq data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      results = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
# getwd()
```

```{r loading_packages, include = FALSE, message = FALSE}
# Load the required packages
library('readr')
library('dplyr', 
        lib.loc = "/camp/home/littlet/R/x86_64-pc-linux-gnu-library/4.0")
library('stringr')
library('purrr')
library('IRanges')
library('kableExtra')
library('plotly')
library('seqinr')
library('rmarkdown')
```

```{r function_ordering}
ordering <- function(to_order, order_vector, fromLast = TRUE){
  order = unlist(sapply(order_vector, 
                        function(x){unique(str_subset(to_order, 
                                                      pattern = paste(x)))}))
  order_missing = unique(to_order[!to_order %in% order])
  unique(c(order, order_missing), fromLast = fromLast)
}
```

```{r load_data}
pir.info <- read.csv('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/github_project/tl22-01_vivaxDeNovo/PlasmoDBv56_VvxPirs_P01_Sal1.csv', 
                     header = TRUE)
```

```{r filter_nident}
#Set a filter for bp length to assume is a valid match
filter_nident <- 100
```

```{r assembly_name_expt}
#Extract the first two numbers of the name e.g. ERX2560125_ERR2541707
accession_name <- str_extract(params$assembly_name, 
                              pattern = '(SR|ER).{8,9}_(SR|ER).{8,9}(?=_)')
#If this works (not NA and only two elements) then find the expt name in the table
if(!is.na(accession_name) & length(unlist(strsplit(accession_name, '_'))) == 2){
  accession_table <- read.csv('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/samples_to_cat.csv')
  expt_name <- accession_table[grepl(accession_table$fastq_1, pattern = accession_name),][,1]
} else {
  #Otherwise assume that the expt name is in the assembly_name
  expt_name <- str_extract(params$assembly_name, 
                           pattern = '[:upper:][:lower:]+[:digit:]{2}')
}
if(rlang::is_empty(expt_name)){
  stop(paste0('No experiment name for accession ', accession_name))
}
```

Filtering matches by a minimum of `r filter_nident`bp to be a valid match.

```{r read_blast_results}
# list.files('../blast', pattern = 'outfmt')

any_blast <- system(paste0('wc -l ', 
              params$blast_output_file), 
       intern = TRUE)
any_blast <- !grepl(any_blast, pattern = '^0 ')
if(any_blast){
outfmt6_headers <- 
  'qseqid sseqid pident length mismatch gapopen qstart qend sstart send evalue bitscore qcovs qcovhsp slen nident' %>% 
  strsplit(split = ' ') %>% 
  unlist

blast <- read.table(params$blast_output_file,
                    col.names = outfmt6_headers) %>% 
  mutate(
         scovhsp = round( (nident/slen * 100), digits = 2),
         scovs = round(((abs(send - sstart) + 1)/slen *100), digits = 2),
         evalue = format(evalue, scientific = TRUE)
  ) %>% 
  #Filter for the pir genes.
  # The Sal1 and Zhu genes are already only the virs so I grep for them first in ifelse.
  # Then, for the P01, I extract the gene name and search in the pir.info list.
  filter(nident > filter_nident,
         ifelse(grepl(sseqid, pattern = 'PVX|comp'),
                TRUE,
                str_extract_all(sseqid, 
                                pattern = '.+(?=\\.1)') %in% pir.info$Gene.ID)) %>%
  group_by(qseqid) %>% 
  mutate(num_genes_match = n_distinct(sseqid))

# Filter for the top hits for each de novo transcript by e-value.

blast_top <- group_by(blast, qseqid) %>% 
  slice_min(evalue, n = 1)

#Filter for the de novo transcripts which only match a single known pir.
# Also removing idents below 100 because these are probably poor matches.

blast_uniq <- filter(blast, num_genes_match == 1)

#sanity test
# all(blast$scovs >= blast$scovhsp)
} else {
  blast <- 0
  print('There are no BLAST hits to the P01, Sal1 or Zhu16 de novo pirs')
}
```

# ORF analysis

```{r orf_fasta}
any_orfs <- system(paste0('wc -l ', params$orf_file), intern = TRUE)
any_orfs <- !grepl(any_orfs, pattern = '^0 ')
if(any_orfs){
  orf_fasta <- read.fasta(params$orf_file,
                          seqtype = c("AA"),
                          seqonly = FALSE)
} else {
  orf_fasta <- 0
  print("There are no pir gene ORFs detected by the criteria used.")
}
```

```{r orf_lengths, eval = any_orfs}

getLength(orf_fasta) %>% log %>% hist(breaks = 1000, 
                                      main = 'log ORF lengths with 150aa line')
abline(v = log(150), col = 'blue')

```

# Number of hits for each de novo transcript with a BLAST hit.

Are there many de novo transcripts which match multiple different known transcripts? Are most of them unique?

```{r number_of_hits_per_deNovoTranscript, eval = any_blast}
blast %>% 
  group_by(qseqid) %>% 
  summarise(num = n_distinct(sseqid)) %>% 
  .$num %>% 
  hist(main = 'Histogram of BLAST hits per de novo transcript (50 max)',
       breaks = 50)
```

#Genes with the most number of uniquely mapping de novo transcripts.

```{r number_of_uniq_transcripts_per_gene, eval = any_blast}
blast %>% 
  filter(num_genes_match== 1) %>% 
  group_by(sseqid) %>% 
  summarise(num_of_uniq_deNovoTranscripts = n_distinct(qseqid),
            max_id = max(pident),
            min_id = min(pident),
            max_nident = max(nident),
            min_nident = min(nident),
            max_gap = max(gapopen),
            min_gap = min(gapopen),
            max_qcovh = max(qcovhsp),
            min_qcovh = min(qcovhsp),
            max_scovh = max(scovhsp),
            min_scovh = min(scovhsp)) %>% 
  arrange(desc(num_of_uniq_deNovoTranscripts)) %>% 
  kbl() %>% kable_styling() %>% scroll_box(width = "100%", height = "400px")
```

# Coverage of the subject (known transcripts) and query (assembled de novo transcripts) sequences.

Using only the de novo transcripts that uniquely match one known pir transcript.

For our purposes the subject coverage is probably most important.

```{r scov_v_qcov, eval = any_blast}
fig <- plot_ly(
  blast_uniq,
  x = ~qcovhsp,
  y = ~scovhsp,
  text = ~paste(sseqid)
) %>% 
  layout(xaxis = list(title = "De novo transcript coverage"),
         yaxis = list(title = "Known transcript coverage"))
fig
```

Number of unique genes found by BLAST: `r ifelse(any_blast, unique(blast$sseqid) %>% length,0)`.

# Known transcripts with >95% coverage from one de novo transcript.

How many of the known transcripts are covered >=95% by de novo transcripts?

num_match_all and num_match_95 show how many other known transcripts these de novo transcripts match in total or >=95% respectively.

```{r transcripts_covering_most_of_known_transcripts, eval = any_blast}
transcripts_covered_by_denovo <- blast %>% 
  group_by(qseqid) %>%
  mutate(num_match_all = n_distinct(sseqid)) %>%
  group_by(sseqid) %>% 
  filter(scovhsp >= 95) %>% 
  group_by(qseqid) %>%
  mutate(num_match_95 = n_distinct(sseqid)) %>% 
  arrange(sseqid)
transcripts_covered_by_denovo %>% 
  select(-qstart,-qend,-sstart,-send,-evalue,-bitscore,-qcovs,-slen,-sseqid,-nident,-scovs) %>% 
  kbl() %>% kable_styling() %>% scroll_box(width = "100%", height = "300px")
```

Showing the BLAST alignments of these >=95% coverage de novo transcripts in a graph, and all the matches they have. 
The "n =" number after the transcript name shows how many other BLAST hits the de novo transcript had in total.
Size of each point shows the length of the alignment.

```{r transcripts_covering_most_of_known_transcripts_graph, eval = any_blast}
blast %>% 
  filter(qseqid %in% unique(transcripts_covered_by_denovo$qseqid)) %>%
  group_by(qseqid) %>%
  mutate(num_match_all = n_distinct(sseqid)) %>% 
  mutate(qseqid = paste0(qseqid, '; n = ', num_match_all),
         over95 = ifelse(scovhsp >= 95, 
                         '>=95% coverage', 
                         '<95% coverage')) %>% 
  plot_ly(x = ~qcovhsp,
          y = ~scovhsp,
          text = ~paste(sseqid, '\n', 
                        length, 'bp', '\n',
                        scovhsp, '% coverage of known'),
          color = ~qseqid,
          size = ~length,
          symbol = ~over95,
          symbols = c('triangle-up', 'circle'),
          sizes = c(2,100),
          legendgroup = ~qseqid) %>% 
  layout(xaxis = list(title = "De novo transcript coverage"),
         yaxis = list(title = "Known transcript coverage"))
```

Alignments with >=95% known coverage but <50% de novo transcript coverage.

```{r low_qcovhsp, eval = any_blast}
if(transcripts_covered_by_denovo %>% 
   filter(qcovhsp < 50) %>% 
   nrow() > 0){
  transcripts_covered_by_denovo %>% 
    filter(qcovhsp < 50) %>% 
    kbl() %>% kable_styling() %>% scroll_box(width = "100%", height = "300px")
}
```

# All the BLAST matches

Fix: Colours for low numbers of hits.

```{r transcripts_coverage_allHits_transcripts_graph, eval = any_blast}
blast %>% 
  group_by(qseqid) %>%
  mutate(num_match_all = n_distinct(sseqid)) %>% 
  mutate(qseqid = paste0(qseqid, '; n = ', num_match_all),
         over95 = ifelse(scovhsp >= 95, 
                         '>=95% coverage', 
                         '<95% coverage')) %>% 
  plot_ly(x = ~qcovhsp,
          y = ~scovhsp,
          legendgroup = ~qseqid) %>% 
  add_markers(text = ~paste(sseqid, '\n', 
                            length, 'aa', '\n',
                            scovhsp, '% coverage of known'),
              color = ~qseqid,
              size = ~length,
              symbol = ~over95,
              symbols = c('triangle-up', 'circle'),
              colors = "Set1",
              sizes = c(2,100)) %>% 
  layout(xaxis = list(title = "De novo transcript coverage"),
         yaxis = list(title = "Known transcript coverage"))
```

# BLAST hit coverages for the top alignments.

```{r transcripts_coverage_topBLAST_transcripts_graph, eval = any_blast}
blast %>% 
  group_by(qseqid) %>%
  mutate(num_match_all = n_distinct(sseqid)) %>% 
  mutate(qseqid = paste0(qseqid, '; n = ', num_match_all),
         over95 = ifelse(scovhsp >= 95, 
                         '>=95% coverage', 
                         '<95% coverage')) %>% 
  group_by(sseqid) %>% 
  slice_min(evalue, n = 1) %>% 
  plot_ly(x = ~qcovhsp,
          y = ~scovhsp,
          legendgroup = ~qseqid) %>% 
  add_markers(text = ~paste(sseqid, '\n', 
                            length, 'aa', '\n',
                            scovhsp, '% coverage of known'),
              color = ~qseqid,
              size = ~length,
              symbol = ~over95,
              symbols = c('triangle-up', 'circle'),
              colors = "Set3",
              sizes = c(2,100)) %>% 
  layout(xaxis = list(title = "De novo transcript coverage"),
         yaxis = list(title = "Known transcript coverage"))
```

# Novel ORFs for ORFs not matching known virs.

Novel ORFs are described as those that don't match any known pir >=95% identity.

```{r orf_novel, eval = any_blast}
num_orf_novel_95 <- blast %>% 
  filter(scovhsp < 95) %>% 
  .$qseqid %>% 
  unique %>% 
  length

mRNA_novel_95 <- blast %>% 
  filter(scovhsp < 95) %>% 
  .$qseqid %>% 
  str_extract('(?<=NODE_)[[:digit:]]+(?=_)') %>% 
  unique

num_mRNA_novel_95 <-  length(mRNA_novel_95)

blast %>% 
  filter(scovhsp < 95) %>% 
  .$qseqid %>% 
  unique %>% 
  str_extract(pattern = '(?<=length_)[:digit:]+(?=_)') %>% 
  as.numeric %>% 
  hist(breaks = 50, main = "Novel ORF lengths (<95% id match known pirs)")
```

## Novel ORFs matched by percentage coverage threshold.

```{r orf_coverage_by_percentage, eval = any_blast}

orfs_covered_by_denovo_diffcov <- lapply(seq(0,100,by = 5),
                                                function(covrg) {
                                                  blast %>% 
                                                    group_by(sseqid) %>% 
                                                    filter(scovhsp < covrg) %>%
                                                    .$qseqid %>% unique %>% length
                                                })
names(orfs_covered_by_denovo_diffcov) <- seq(0,100,by = 5)
plot(y = unlist(orfs_covered_by_denovo_diffcov), 
     x = seq(0,100,by = 5),
     # ylim = c(0,max(unlist(orfs_covered_by_denovo_diffcov))),
     # main = 'For all de novo transcripts',
     ylab = 'Number of novel ORFs',
     xlab = 'Percentage of coverage matching a known pir',
     yaxt = 'none')
axis(2, at = seq(0,nrow(pir.info),by = 100))
# abline(a = 25, b = 0)

```

# BLAST hit coverages for the top alignments to the Zhu 2016 de novo virs.

```{r transcripts_coverage_topBLAST_Zhu16transcripts_graph, eval = any_blast}
blast %>% 
  group_by(qseqid) %>%
  mutate(num_match_all = n_distinct(sseqid)) %>% 
  mutate(qseqid = paste0(qseqid, '; n = ', num_match_all),
         over95 = ifelse(scovhsp >= 95, 
                         '>=95% coverage', 
                         '<95% coverage')) %>% 
  filter(grepl(sseqid, 
               pattern = 'comp')) %>% 
  group_by(sseqid) %>% 
  slice_min(evalue, n = 1) %>% 
  plot_ly(x = ~qcovhsp,
          y = ~scovhsp,
          text = ~paste(sseqid, '\n', 
                        length, 'aa', '\n',
                        scovhsp, '% coverage of known'),
          color = ~qseqid,
          size = ~length,
          symbol = ~over95,
          symbols = c('triangle-up', 'circle'),
          sizes = c(2,100),
          legendgroup = ~qseqid) %>% 
  layout(xaxis = list(title = "De novo transcript coverage"),
         yaxis = list(title = "Known transcript coverage"))
```

# Transcripts matched by percentage coverage.

```{r transcript_coverage_by_percentage, eval = any_blast}

transcripts_covered_by_denovo_diffcov <- lapply(seq(0,100,by = 5),
                                                function(covrg) {
                                                  blast %>% 
                                                    group_by(sseqid) %>% 
                                                    filter(scovhsp >= covrg) %>%
                                                    .$sseqid %>% unique %>% length
                                                })
names(transcripts_covered_by_denovo_diffcov) <- seq(0,100,by = 5)
plot(y = unlist(transcripts_covered_by_denovo_diffcov), 
     x = seq(0,100,by = 5),
     ylim = c(0,nrow(pir.info)),
     main = 'For all de novo transcripts',
     ylab = 'Number of known transcripts covered',
     xlab = 'Percentage of coverage by individual transcripts',
     yaxt = 'none')
axis(2, at = seq(0,nrow(pir.info),by = 100))
# abline(a = 25, b = 0)

```

What data is worth saving? 
* Number of ORFs found.
* Number of ORFs over certain lengths.

```{r orfs_to_save}
number_transcripts <- system(paste0('cat ',
                                    params$assembly_file, 
                                    ' | grep ">" | wc -l'),
                             intern = TRUE)
if(any_orfs){
  number_orfs_unique_mRNA <- str_extract(getName(orf_fasta),
                                         pattern = '(?<=NODE_)[[:digit:]]+(?=_)') %>% 
    unique() %>% 
    length

  orf_lengths <- getLength(orf_fasta)

  orfs_unique_mRNA_Over200aaLength <- str_extract(getName(orf_fasta[orf_lengths >= 200]),
                                                  pattern = '(?<=NODE_)[[:digit:]]+(?=_)') %>% 
    unique()

  number_orfs_unique_mRNA_Over200aaLength <-  length(orfs_unique_mRNA_Over200aaLength)

  num_ORFs_over200aaLength = sum(orf_lengths>=200)
  num_ORFs_novel_nomatch95known = num_orf_novel_95
  num_uniq_mRNA_ORFs_predict = number_orfs_unique_mRNA
  num_uniq_mRNA_ORFs_over200aaLength = number_orfs_unique_mRNA_Over200aaLength

} else {
  number_orfs_unique_mRNA <- orf_lengths <- orfs_unique_mRNA_Over200aaLength <- number_orfs_unique_mRNA_Over200aaLength <- num_ORFs_over200aaLength <- num_ORFs_novel_nomatch95known <- num_uniq_mRNA_ORFs_predict <- num_uniq_mRNA_ORFs_over200aaLength <- 0
}
```

```{r blast_to_save}
if(any_blast){
  mRNA_novel_95_over200aa <- mRNA_novel_95[mRNA_novel_95 %in% orfs_unique_mRNA_Over200aaLength]

  num_mRNA_novel_95_over200aa <- length(mRNA_novel_95_over200aa)
  num_uniq_mRNA_ORF_novel_nomatch95known = num_mRNA_novel_95
  knwn_trnscrpt_cvg_over75 = transcripts_covered_by_denovo_diffcov[['75']]
  ancestral_over75 = blast %>% 
    group_by(qseqid) %>%
    mutate(num_match_all = n_distinct(sseqid)) %>%
    group_by(sseqid) %>% 
    filter(scovhsp >= 75, 
           grepl(sseqid, 
                 pattern = 'PVP01_1147700|PVX_113230')) %>% 
    nrow
  knwn_trnscrpt_cvg_over95 = length(unique(transcripts_covered_by_denovo$sseqid))
  zhu16_trnscrpt_cvg_over95 = length(unique(str_subset(transcripts_covered_by_denovo$sseqid, 
                                                       pattern = 'comp')))
  Sal1_trnscrpt_cvg_over95 = length(unique(str_subset(transcripts_covered_by_denovo$sseqid, 
                                                      pattern = 'PVX')))
  P01_trnscrpt_cvg_over95 = length(unique(str_subset(transcripts_covered_by_denovo$sseqid, 
                                                     pattern = 'PVP01')))
  ancestral_over95 = ifelse(
    any(grepl(transcripts_covered_by_denovo$sseqid,
              pattern = 'PVP01_1147700|PVX_113230')),
    paste(transcripts_covered_by_denovo$sseqid[
      grepl(transcripts_covered_by_denovo$sseqid,
            pattern = 'PVP01_1147700|PVX_113230')
    ], sep = '; '),
    0)
  mean_mismatches = mean(transcripts_covered_by_denovo$mismatch)
  mean_gaps = mean(transcripts_covered_by_denovo$gapopen)
  med_mismatches = median(transcripts_covered_by_denovo$mismatch)
  med_gaps = median(transcripts_covered_by_denovo$gapopen)
}else{
  mRNA_novel_95_over200aa <- num_mRNA_novel_95_over200aa <- num_uniq_mRNA_ORF_novel_nomatch95known <- knwn_trnscrpt_cvg_over75 <- ancestral_over75 <- knwn_trnscrpt_cvg_over95 <- zhu16_trnscrpt_cvg_over95 <- Sal1_trnscrpt_cvg_over95 <- P01_trnscrpt_cvg_over95 <- ancestral_over95 <- mean_mismatches <- mean_gaps <- med_mismatches <- med_gaps <- 0
}

number_orfs <- system(paste0('cat ',
                             params$orf_file,
                             ' | grep ">" | wc -l'), 
                      intern = TRUE)
```

```{r data_save}

df_save <- data.frame("experiment" = expt_name,
                      "assembly" = params$assembly_name,
                      "num_uniq_mRNA_pirHMMerORFs_nomatch95known_over200aa" = num_mRNA_novel_95_over200aa,
                      "num_transcripts" = number_transcripts,
                      "num_ORFs_predict" = number_orfs,
                      "num_ORFs_over200aaLength" = num_ORFs_over200aaLength,
                      "num_ORFs_novel_nomatch95known" = num_ORFs_novel_nomatch95known,
                      "num_uniq_mRNA_ORFs_predict" = num_uniq_mRNA_ORFs_predict,
                      "num_uniq_mRNA_ORFs_over200aaLength" = num_uniq_mRNA_ORFs_over200aaLength,
                      "num_uniq_mRNA_ORF_novel_nomatch95known" = num_uniq_mRNA_ORF_novel_nomatch95known,
                      "knwn_trnscrpt_cvg_over75" = knwn_trnscrpt_cvg_over75,
                      "ancestral_over75" = ancestral_over75,
                      "knwn_trnscrpt_cvg_over95" = knwn_trnscrpt_cvg_over95,
                      "zhu16_trnscrpt_cvg_over95" = zhu16_trnscrpt_cvg_over95,
                      "Sal1_trnscrpt_cvg_over95" = Sal1_trnscrpt_cvg_over95,
                      "P01_trnscrpt_cvg_over95" = P01_trnscrpt_cvg_over95,
                      "ancestral_over95" = ancestral_over95,
                      "mean_mismatches" = mean_mismatches,
                      "mean_gaps" = mean_gaps,
                      "med_mismatches" = med_mismatches,
                      "med_gaps" = med_gaps,
                      "assembly_file" = params$assembly_file,
                      "orf_file" = params$orf_file,
                      "blast_file" = params$blast_output_file
)

if(!file.exists('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/tl22-01_assembly_info.txt') | params$new_table){
  #If table file doesn't exist then create it.
  write.table(df_save,
              file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/tl22-01_assembly_info.txt',
              sep = '\t',
              col.names = TRUE,
              append = FALSE,
              row.names = FALSE)
} else if (params$assembly_name %in% read.delim('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/tl22-01_assembly_info.txt')$assembly){
  #If assembly_name is already in table then overwrite it.
  assembly_info <- read.delim('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/tl22-01_assembly_info.txt')
  assembly_info[assembly_info$assembly %in% params$assembly_name,] <- df_save[1,]
  write.table(assembly_info,
              file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/tl22-01_assembly_info.txt',
              sep = '\t',
              row.names = FALSE)
} else {
  #Append to the existing table is both above are false
  write.table(df_save,
              file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/tl22-01_assembly_info.txt',
              sep = '\t',
              col.names = FALSE,
              append = TRUE,
              row.names = FALSE)
}

df_save %>% 
  kbl() %>% kable_styling() %>% scroll_box(width = "100%", height = "400px")

```