### 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_VvxDeNovo_getorfFind1_hmmerMatches_blast2VvxProt.Rmd

---
title: "tl22-01 BLASTing the HMMer matches from Plasmodium_Vir (using ORFs from getorf -find 1 option) to Plasmodium vivax P01 v56 proteins"
output: 
  html_document:
    theme: cosmo
    code_download: true
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
params:
  blast_output_file: '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/blast/blast_to_Proteins/scspades_getorfTableZero/SRX697844_SRR1571714/SRX697844_SRR1571714_scspades_getorfTableZero_hmmerPfam-Plasmodium_Vir_bast-VvxProt.outfmt6'
---

# Intro

This script is looks at the BLAST results of the HMMer matches with the Plasmodium_Vir among the ORFs from the Plasmodium vivax RNAseq scspades assembly. BLASTed to the Plasmodium vivax P01 PlasmoDB v56 proteins.

# To-do:
* BLASTs to non-pirs?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      results = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
```

```{r loading_packages, include = FALSE, message = FALSE}
# Load the required packages
library('readr')
library('dplyr', 
        lib.loc = "/camp/home/littlet/R/x86_64-pc-linux-gnu-library/4.0")
library('stringr')
library('purrr')
library('IRanges')
library('kableExtra')
library('plotly')
```

```{r function_ordering}
ordering <- function(to_order, order_vector, fromLast = TRUE){
  order = unlist(sapply(order_vector, 
                        function(x){unique(str_subset(to_order, 
                                                      pattern = paste(x)))}))
  order_missing = unique(to_order[!to_order %in% order])
  unique(c(order, order_missing), fromLast = fromLast)
}
```

```{r function_plotRanges}
#' Plot genomic ranges from the IRanges object or GRanges object.
# from https://rdrr.io/github/JokingHero/rxtras/src/R/plot_wrappers.R

plotRanges <- function(x, xlim = x, main = deparse(substitute(x)),  col = "black", sep = 0.5, ...){
  if (!requireNamespace("IRanges", quietly = TRUE)) {
    stop("IRanges needed for this function to work. Please install it.",
         call. = FALSE)
  }
  height <- 1
  if (is(xlim, "Ranges"))
      xlim <- c(min(start(xlim)), max(end(xlim)))
  bins <- disjointBins(IRanges(start(x), end(x) + 1))
  plot.new()
  plot.window(xlim, c(0, max(bins) * (height + sep)))
  ybottom <- bins * (sep + height) - height
  rect(start(x) - 0.5, ybottom, end(x) + 0.5, ybottom + height, col = col, ...)
  title(main)
  axis(1)
}
```

```{r load_data}
pir.info <- read.csv('PlasmoDBv56_VvxPirs_P01_Sal1.csv', 
                     header = TRUE)
```

```{r read_blast_results}
#Set a filter for bp length to assume is a valid match
filter_nident <- 100

outfmt6_headers <- 
  'qseqid sseqid pident length mismatch gapopen qstart qend sstart send evalue bitscore qcovs qcovhsp slen nident' %>% 
  strsplit(split = ' ') %>% 
  unlist

blast_all <- read.table(params$blast_output_file,
                        col.names = outfmt6_headers) %>% 
  mutate(
    sseqid = unlist(
      lapply(
        str_extract_all(sseqid, 
                        pattern = '.+(?=\\.1)'), 
        paste, 
        collapse = '_')
    ),
    scovhsp = round( (nident/slen * 100), digits = 2),
    scovs = round(((abs(send - sstart) + 1)/slen *100), digits = 2),
    evalue = format(evalue, scientific = TRUE)
  ) %>% 
  filter(nident > filter_nident) %>%
  group_by(qseqid) %>% 
  mutate(num_genes_match = n_distinct(sseqid))

#Filter for the pir genes.
blast <- filter(blast_all, 
                sseqid %in% pir.info$Gene.ID) %>%
  group_by(qseqid) %>% 
  mutate(num_genes_match = n_distinct(sseqid))

#Non-pir genes
blast_nonpir <- filter(blast_all, 
                       !sseqid %in% pir.info$Gene.ID) %>%
  group_by(qseqid) %>% 
  mutate(num_genes_match = n_distinct(sseqid))

# Filter for the top hits for each de novo transcript by e-value.

blast_top <- group_by(blast, qseqid) %>% 
  slice_min(evalue, n = 1)

#Filter for the de novo transcripts which only match a single known pir.
# Also removing idents below 100 because these are probably poor matches.

blast_uniq <- filter(blast, num_genes_match == 1)

#sanity test
# all(blast$scovs >= blast$scovhsp)
```

Filtering matches by a minimum of `r filter_nident`bp to be a valid match.

# Number of hits for each de novo transcript with a BLAST hit.

Are there many de novo transcripts which match multiple different known transcripts? Are most of them unique?

```{r number_of_hits_per_deNovoTranscript}
blast %>% 
  group_by(qseqid) %>% 
  summarise(num = n_distinct(sseqid)) %>% 
  .$num %>% 
  hist(main = 'Histogram of BLAST hits per de novo transcript (50 max)',
       breaks = 50)
```

#Genes with the most number of uniquely mapping de novo transcripts.

```{r number_of_uniq_transcripts_per_gene}
blast %>% 
  filter(num_genes_match== 1) %>% 
  group_by(sseqid) %>% 
  summarise(num_of_uniq_deNovoTranscripts = n_distinct(qseqid),
            max_id = max(pident),
            min_id = min(pident),
            max_nident = max(nident),
            min_nident = min(nident),
            max_gap = max(gapopen),
            min_gap = min(gapopen),
            max_qcovh = max(qcovhsp),
            min_qcovh = min(qcovhsp),
            max_scovh = max(scovhsp),
            min_scovh = min(scovhsp)) %>% 
  arrange(desc(num_of_uniq_deNovoTranscripts)) %>% 
  kbl() %>% kable_styling() %>% scroll_box(width = "100%", height = "400px")
```

# Known transcripts with >95% coverage from one de novo transcript.

How many of the known transcripts are covered >=95% by de novo transcripts?

num_match_all and num_match_95 show how many other known transcripts these de novo transcripts match in total or >=95% respectively.

```{r transcripts_covering_most_of_known_transcripts}
transcripts_covered_by_denovo <- blast %>% 
  group_by(qseqid) %>%
  mutate(num_match_all = n_distinct(sseqid)) %>%
  group_by(sseqid) %>% 
  filter(scovhsp >= 95) %>% 
  group_by(qseqid) %>%
  mutate(num_match_95 = n_distinct(sseqid)) %>% 
  arrange(sseqid)
transcripts_covered_by_denovo %>% 
  select(-qstart,-qend,-sstart,-send,-evalue,-bitscore,-qcovs,-slen,-sseqid,-nident,-scovs) %>% 
  kbl() %>% kable_styling() %>% scroll_box(width = "100%", height = "300px")
```

Showing the BLAST alignments of these >=95% coverage de novo transcripts in a graph, and all the matches they have. 
The "n =" number after the transcript name shows how many other BLAST hits the de novo transcript had in total.
Size of each point shows the length of the alignment.

Note: PVP01_1147700 = ancestral

```{r transcripts_covering_most_of_known_transcripts_graph}
blast %>% 
  filter(qseqid %in% unique(transcripts_covered_by_denovo$qseqid)) %>%
  group_by(qseqid) %>%
  mutate(num_match_all = n_distinct(sseqid)) %>% 
  mutate(qseqid = paste0(qseqid, '; n = ', num_match_all),
         over95 = ifelse(scovhsp >= 95, 
                         '>=95% coverage', 
                         '<95% coverage')) %>% 
  plot_ly(x = ~qcovhsp,
          y = ~scovhsp,
          text = ~paste(sseqid, '\n', 
                        length, 'aa', '\n',
                        scovhsp, '% coverage of known'),
          color = ~qseqid,
          size = ~length,
          symbol = ~over95,
          symbols = c('triangle-up', 'circle'),
          sizes = c(2,100),
          legendgroup = ~qseqid) %>% 
  layout(xaxis = list(title = "De novo transcript coverage"),
         yaxis = list(title = "Known transcript coverage"))
```

Alignments with >=95% known coverage but <50% de novo transcript coverage.

```{r low_qcovhsp}
if(transcripts_covered_by_denovo %>% 
   filter(qcovhsp < 50) %>% 
   nrow() > 0){
  transcripts_covered_by_denovo %>% 
    filter(qcovhsp < 50) %>% 
    kbl() %>% kable_styling() %>% scroll_box(width = "100%", height = "300px")
}
```

# All the BLAST matches

```{r transcripts_coverage_allHits_transcripts_graph}
blast %>% 
  group_by(qseqid) %>%
  mutate(num_match_all = n_distinct(sseqid)) %>% 
  mutate(qseqid = paste0(qseqid, '; n = ', num_match_all),
         over95 = ifelse(scovhsp >= 95, 
                         '>=95% coverage', 
                         '<95% coverage')) %>% 
  plot_ly(x = ~qcovhsp,
          y = ~scovhsp,
          text = ~paste(sseqid, '\n', 
                        length, 'aa', '\n',
                        scovhsp, '% coverage of known'),
          color = ~qseqid,
          size = ~length,
          symbol = ~over95,
          symbols = c('triangle-up', 'circle'),
          sizes = c(2,100),
          legendgroup = ~qseqid) %>% 
  layout(xaxis = list(title = "De novo transcript coverage"),
         yaxis = list(title = "Known transcript coverage"))
```

# BLAST hit coverages for the top alignments.

```{r transcripts_coverage_topBLAST_transcripts_graph}
blast %>% 
  group_by(qseqid) %>%
  mutate(num_match_all = n_distinct(sseqid)) %>% 
  mutate(qseqid = paste0(qseqid, '; n = ', num_match_all),
         over95 = ifelse(scovhsp >= 95, 
                         '>=95% coverage', 
                         '<95% coverage')) %>% 
  group_by(sseqid) %>% 
  slice_min(evalue, n = 1) %>% 
  plot_ly(x = ~qcovhsp,
          y = ~scovhsp,
          text = ~paste(sseqid, '\n', 
                        length, 'aa', '\n',
                        scovhsp, '% coverage of known'),
          color = ~qseqid,
          size = ~length,
          symbol = ~over95,
          symbols = c('triangle-up', 'circle'),
          sizes = c(2,100),
          legendgroup = ~qseqid) %>% 
  layout(xaxis = list(title = "De novo transcript coverage"),
         yaxis = list(title = "Known transcript coverage"))
```

# Transcripts matched by percentage coverage.

```{r transcript_coverage_by_percentage}

transcripts_covered_by_denovo_diffcov <- lapply(seq(0,100,by = 5),
                                                function(covrg) {
                                                  blast %>% 
                                                    group_by(sseqid) %>% 
                                                    filter(scovhsp >= covrg) %>%
                                                    .$sseqid %>% unique %>% length
                                                })
names(transcripts_covered_by_denovo_diffcov) <- seq(0,100,by = 5)
plot(y = unlist(transcripts_covered_by_denovo_diffcov), 
     x = seq(0,100,by = 5),
     ylim = c(0,nrow(pir.info)),
     main = 'For pir de novo transcripts',
     ylab = 'Number of known transcripts covered',
     xlab = 'Percentage of coverage by individual transcripts',
     yaxt = 'none')
axis(2, at = seq(0,nrow(pir.info),by = 100))
# abline(a = 25, b = 0)

```

# Important instant details:

Filtered matches by a minimum of `r filter_nident`bp to be a valid match.

There are `r length(unique(transcripts_covered_by_denovo$sseqid))` known transcripts covered by >=95% of at least one individual _de novo_ transcript. There is a mean of `r mean(transcripts_covered_by_denovo$mismatch)` mismatches and `r mean(transcripts_covered_by_denovo$gapopen)` gaps and a median of `r median(transcripts_covered_by_denovo$mismatch)` mismatches and `r median(transcripts_covered_by_denovo$gapopen)` gaps in these alignments (Note that there are probably some de novo transcripts that align >=95% to the same known transcript). Of these `r length(unique(transcripts_covered_by_denovo$sseqid[transcripts_covered_by_denovo$num_match_all == 1]))`
 match only that known transcript and no others by the BLAST search. `r length(unique(filter(ungroup(transcripts_covered_by_denovo), qcovhsp < 50)$sseqid))` of the genes with >=95% known transcript alignments cover less than 50% of the de novo transcript.

See section [Known transcripts with >95% coverage from one de novo transcript.].

 With a threshold of 75% coverage `r transcripts_covered_by_denovo_diffcov['75']` known transcripts are represented by the de novo transcripts.

See section [Transcripts matched by percentage coverage.].