### 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_analysingNetworkStats.Rmd

---
title: "tl22-01_analysingNetworkStats"
output: html_document
---

Going to adapt this script to analyse multiple networks.

What do we want from the optimal network?:

* More sequences included among the larger clusters of the network.
* Sal1 sub-families are rarely mixed in the same clusters.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      results = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
# getwd()
```

```{r load_packages}
library(ggplot2)
library(stringr)
library(dplyr)
library(data.table)
library(seqinr)
safe_colorblind_palette <- c("#88CCEE", "#CC6677", "#DDCC77", "#117733", "#332288", "#AA4499", 
                             "#44AA99", "#999933", "#882255", "#6699CC", "#661100", "#888888")
global_save_var <- FALSE
```

```{r function_alternateReverse}
alternateReverse <- function(input.vec){
  #Alternate the indexes of a vector, with one alternation in reverse order.
  # Good for colour systems so the end of a gradient is shifted between very different colours.
  output.vec <- c(rbind(input.vec[seq_along(input.vec)%%2==1], 
                        rev(input.vec[seq_along(input.vec)%%2==0])))
  if(length(output.vec) != length(input.vec)){
    #If input.vec is odd length rbind will add an extra value at the end.
    # So the length will be longer, this removes the last, erroneous element.
    return(output.vec[-last(output.vec)])
  }else{
    return(output.vec)
  }
}
```

```{r function_uniquify}
#Make unique identifiers from a vector of names by appending '_X' where X is the number of the 'replicate'.
uniquify <- function(vector_names, first_one = TRUE){
  new_vector <- c()
  df_num <- data.frame(names = unique(vector_names), rep = 1)
  if(first_one){
    duplicated_names <- unique(vector_names[duplicated(vector_names)])
  }else{
    duplicated_names <- c()
  }
  for(i in 1:length(vector_names)){
    if(vector_names[i] %in% duplicated_names|!first_one){
      name <- paste(vector_names[i],
                    df_num$rep[match(vector_names[i], df_num$names)],
                    sep = '_')
      #Increase rep counter
      df_num$rep[match(vector_names[i], df_num$names)] <- df_num$rep[match(vector_names[i], df_num$names)] + 1
      new_vector[i] <- name
    }else{
      new_vector[i] <- vector_names[i]
    }
  }
  new_vector
}
```

```{r read_data}
accession_table <- read.csv('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/samples_to_cat.csv')
assembly.df <- read.delim('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/tl22-01_assembly_info.txt')
subfam.df <- read.delim('/camp/lab/langhornej/working/HPC/littlet/tl22-02_VirSubfamNetwork/pir_info/lopez13_vir_subfam.txt', header = FALSE)
colnames(subfam.df) <- c('Geneid','subfam')
expt.df <- read.delim('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/scripts/all-cat.list', header = FALSE, sep = '')
colnames(expt.df) <- c('Accession', 'expt')
expt_num.df <- read.csv('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/github_project/tl22-01_vivaxDeNovo/expt_codes.csv')
```

```{r MCL_parameters}
# list.files('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/tpmOver1', 
  #          pattern = 'mci.+list') %>% 
  # head %>% 
  # str_extract_all(pattern = '(?<=_)[[:digit:]]+(?=pid)|[[:digit:]]+(?=lngth)|(?<=mci\\.)I.+(?=\\.list)')

network_file_dir <- '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/tpmOver1'

#Search for all the MCI lists of the clustering, extract the relevant bits which shows the % id, alignment length and inflation factor.
network_parameters.list <- list.files(network_file_dir, 
           pattern = 'all-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1_all2allBlast.mci.I.+list') %>% 
  str_extract(pattern = 'I.+(?=\\.list)') %>% 
  # str_replace(pattern = 'NoSame2Same\\.mci\\.', replacement = '_') %>% 
  unique

#Get the MCI cluster values

mcl.list <- lapply(network_parameters.list,
                   function(network_parameters){
                     network_parameters.file <- list.files(network_file_dir,
                                                           pattern = paste0(
                                                             'all-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1_all2allBlast.mci.',
                                                             network_parameters, 
                                                             '.list'
                                                           ),
                                                           full.names = TRUE)
                     if(length(network_parameters.file) > 1){
                       stop(paste0('More than 1 possible file found for one set of network parameters for ',
                                   network_parameters))
                     }

        #Read_table saves the number of the line of the sequences (each line is a cluster) as a separate column
        network_parameters.table <- read.table(network_parameters.file)
        colnames(network_parameters.table) <- c('label', 'mcl_cluster')

        return(network_parameters.table)
       })
names(mcl.list) <- network_parameters.list
```

```{r get_MCLclusters}
fasta <- read.fasta(file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/all-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1.fasta')

inflation_values <- unique(str_extract(names(mcl.list),
                                       pattern = '(?<=I).+'))

# sequence.df <- lapply(names(fasta),
#        function(sequence.name){
#          if(grepl(sequence.name,
#                   pattern = 'P01|PVX|comp')){
#            #For the known pirs, which have P01, PVX or comp in the sequence name.
#            seq_source <- case_when(
#              grepl(sequence.name, pattern = 'P01') ~ 'P01',
#              grepl(sequence.name, pattern = 'PVX') ~ 'Sal1',
#              grepl(sequence.name, pattern = 'comp') ~ 'Zhu16'
#            )
#            seq_expt <- 'known'
#            #Get the subfam number for the Sal1 sequences.
#            seq_subfam <- ifelse(seq_source == 'Sal1',
#                                 subfam.df$subfam[match(sequence.name,
#                                                        subfam.df$Geneid)],
#                                 seq_source)
#            #Need to set the Sal1 sequences which don't have sub-family assignments to just Sal1
#            #  They are NA at this point.
#            seq_subfam <- ifelse(is.na(seq_subfam) && seq_source == 'Sal1',
#                   'Sal1',
#                   seq_subfam)
# 
#          }else{
#            #Get source name of de novo pirs from the first section of the sequence name
#            seq_source <- str_extract(sequence.name,
#                                      pattern = ".+(?=_NODE.+)")
#            seq_subfam <- 'assembled'
#            seq_expt <- expt.df$expt[match(seq_source, expt.df$Accession)]
#          }
# 
#          #Get Id (just a number) from the index of the sequence.
# 
#          info.df <- data.frame(Id = which(names(fasta) == sequence.name),
#                                label = sequence.name,
#                                source = seq_source,
#                                subfam = seq_subfam,
#                                expt = seq_expt)
# 
#          #Get the 10pid25lngthNoSame2Same.mci.IX values as this uses the most data and is the MCL clustering done with the most data.
# # 
# # mcl_10pid25lngth <- lapply(inflation_values,
# #                            function(infl){
# #                              mcl.df <- getElement(mcl.list,
# #                                                   paste0('10pid25lngthNoSame2Same.mci.I',
# #                                                          infl))
# #                            })
# # names(mcl_10pid25lngth) <- inflation_values
# 
#          #Go through the mcl.list to get the mcl cluster number for this sequence in each cluster.
#          # Make a dataframe to bind_cols with the previous information df.
#          sequence_mcl.df <- lapply(names(mcl.list),
#                 function(network_parameters){
#                   mcl.df <- getElement(mcl.list, network_parameters)
#                   #If it is in this network (i.e. has a BLAST alignment), then get cluster number
#                   # from 10pid25lngthNoSame2Same.mci.IX.
#                   # If not in the network, give a cluster number of zero
#                   # mcl_10pid25lngth_netpara <- getElement(mcl_10pid25lngth,
#                   #                                        str_extract(network_parameters,
#                   #                                                    pattern = '(?<=I).+'))
#                   mcl_cluster_num <- ifelse(sequence.name %in% mcl.df$label,
#                                             mcl.df$mcl_cluster[match(sequence.name,
#                                                                                        mcl.df$label)],
#                                             0)
#                   mcl_cluster_num.df <- data.frame(mcl_cluster_num)
#                   colnames(mcl_cluster_num.df) <- network_parameters
#                   return(mcl_cluster_num.df)
#                 }) %>% bind_cols
# 
#          return(bind_cols(info.df, sequence_mcl.df))
#        }) %>% bind_rows
#

# 
# write.table(sequence.df,
#             file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/sequence-MCLclusters_tpmOver1.txt')

sequence.df <- read.table(file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/sequence-MCLclusters_tpmOver1.txt',
                          check.names = FALSE)

# # How many sequences have all zeros? Sequence.df should have same row num as fasta file length.
notInNetwork <- sequence.df[apply(sequence.df[,grepl(colnames(sequence.df), pattern  = '^I[[:digit:]]')], 
                     1, 
                     function(x) all(x == 0)),]$label
```

```{r blast_networks}
perc_ident.vec <- c(10,20,30,40,50,60,70,80,90)
length_aa.vec <- c(25,50,75,100,150,200,250,300)

# pid_lntgh.df <- lapply(perc_ident.vec, function(perc_ident){
#   lapply(length_aa.vec, function(length_aa){
#     #Read in BLAST file for these percentage id and length aa 
#     # Get a list of the unique sequences from the first two columns
#     blast.names <- read.delim(paste0(
#       '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/tpmOver1/all-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1_all2allBlast_',
#       perc_ident,'pid',
#       length_aa,'lngthNoSame2Same.blast'
#     ),
#     header = FALSE) %>% 
#       .[,c(1,2)] %>% 
#       unlist() %>% 
#       unique
#     #lapply across the sequences in the fasta to record whether they are present in this network (i.e. in the BLAST file)
#     in_network.df <- lapply(names(fasta), 
#            function(seq.id){
#              in_network <- ifelse(seq.id %in% blast.names,
#                                   1,
#                                   0)
#              return(data.frame(label = seq.id, 
#                                in_network = in_network,
#                                pid = perc_ident,
#                                lngth = length_aa))
#            }) %>% bind_rows
#   })
# }) %>% bind_rows

```

```{r long_form_sequence.df}
# sequence.df.melt <- melt(sequence.df, 
#                          id.vars = c('Id','label','subfam','source','expt'),
#                          variable.name = 'network_parameters',
#                          value.name = 'cluster_num') %>% 
#   group_by(network_parameters, cluster_num) %>% 
#   mutate(cluster_size = n_distinct(label)) %>% 
#   group_by(network_parameters) %>% 
#   mutate(total_clusters = n_distinct(cluster_num),
#          total_clusters_over2 = n_distinct(cluster_num[cluster_size > 2]),
#          inflation = str_extract(network_parameters, pattern = '(?<=I).+'))
# 
# # all(sequence.df.melt$total_clusters >= sequence.df.melt$total_clusters_over2)
# 
# #Join the BLAST parameters and whether or not they're part of the network. 
# # Set the cluster_num to zero for not being in the network.
# sequence.df.melt <- inner_join(pid_lntgh.df, sequence.df.melt, by = 'label') %>% 
#   mutate(cluster_num = cluster_num*in_network) %>% 
#   group_by(pid, lngth, inflation) %>% 
#   mutate(num_inNetwork = sum(cluster_num > 0)) %>% 
#   select(-in_network)

# write.table(sequence.df.melt,
#             file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/sequence-MCLclusters_tpmOver1_longForm.txt')
sequence.df.melt <- read.table(file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/sequence-MCLclusters_tpmOver1_longForm.txt')
```

```{r not_in_top100}
group_by(sequence.df.melt, pid, lngth, inflation) %>% 
  filter(cluster_num < 1 | cluster_num > 100) %>% 
  summarise(num = n_distinct(label)) %>% 
  arrange(num) 
```

```{r}
#Networks with large, un-subfam-mixed clusters
# Largest cluster is unmixed?
# Where is the highest I, J, B unmixed cluster?

cluster_subfam_mixing.df <- sequence.df.melt %>% 
    filter(source == 'Sal1',
           subfam %in% c('C','E','G','I','K','J','B'),
           cluster_num > 0) %>% 
  select(-Id, -source, -expt) %>% 
  group_by(across(-label)) %>% 
  #With the subfams I want to see which networks have the largest, unmixed, with all subfam in as few clusters as possible
  summarise(num_members = n_distinct(label)) %>% 
  group_by(pid, lngth, inflation, subfam) %>% 
  mutate(sf_total_groups = n_distinct(cluster_num)) %>% 
  group_by(across(-c(subfam, num_members, sf_total_groups))) %>% 
  summarise(subfam_members = paste(paste0(subfam, ' (', num_members, ' - ', sf_total_groups, ')'), collapse = ', '),
            max_prop = max(num_members)/sum(num_members)) %>% 
  arrange(desc(cluster_size), desc(num_inNetwork))
```

```{r plotly}
df <- group_by(cluster_subfam_mixing.df, pid, lngth, inflation, num_inNetwork) %>%
  mutate(num_clust = n_distinct(cluster_num)) %>% 
  filter(cluster_num %in% c(1:15),
         num_inNetwork > round(4468*0.9)) %>% 
  group_by(num_inNetwork, num_clust, pid, lngth, inflation, 
           total_clusters, total_clusters_over2) %>% 
  summarise(avg_maxProps = sum(max_prop)/n_distinct(cluster_num), 
            max_clustSize = max(cluster_size)) %>% 
  arrange(desc(avg_maxProps), num_clust)

library(plotly)
plot_ly(df, x = ~avg_maxProps, y = ~total_clusters, size = ~num_inNetwork,
        symbol = ~pid, color = ~lngth, text = ~paste0(pid,lngth,inflation))
ggplot(df,
       aes(x = avg_maxProps, 
              y = as.integer(total_clusters), 
              # size = inflation,
              shape = factor(lngth), 
              color = factor(inflation)
           )) +
  scale_color_manual(values = safe_colorblind_palette) +
  geom_point() +
  theme_classic() +
  guides(color = guide_legend(ncol = 2),
         size = guide_legend(ncol = 2),
         shape = guide_legend(ncol = 2)) +
  facet_wrap(facets = vars(pid))
```

```{r}
cluster_subfam_mixing.df %>% 
  filter(grepl(subfam_members, pattern = '^J'))

cluster_subfam_mixing.df %>% 
  filter(grepl(subfam_members, pattern = '^I'))

cluster_subfam_mixing.df %>% 
  filter(grepl(subfam_members, pattern = '^K')) %>% arrange(desc(num_inNetwork))

lapply(c(25, 50, 75, 100, 150),
       function(length){
         filter(cluster_subfam_mixing.df, 
                pid == 50,
                lngth == length,
                inflation == 1.4) %>% 
           select(!c(pid, lngth, inflation))
       })

filter(cluster_subfam_mixing.df, 
       pid == 30,
       lngth == 25,
       inflation == 1.4) %>% 
  select(!c(pid, lngth, inflation))

```

50pid, 50/25 lngth appears to be the best now.

```{r read_data}

# node.df <- filter(sequence.df.melt,
#        pid == 30,
#        lngth == c(25)) %>% 
#   mutate(network_parameters = paste0(pid,'pid',lngth,'lngthI',inflation)) %>% 
#   dcast(label + Id + source + subfam + expt ~ network_parameters, value.var = 'cluster_num')
# 
# write.csv(node.df,
#           '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/sequence-MCLclusters_tpmOver1_30pid25lngthI14.csv',
#           row.names = FALSE,
#           quote = FALSE)

node.df <- read.csv(
  '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/sequence-MCLclusters_tpmOver1_30pid25lngthI14.csv')

node.df[apply(node.df[,str_detect(colnames(node.df), pattern = 'pid.+lngth')], 
              1,
              function(x) all(x == 0)),] %>% 
  filter(!label %in% notInNetwork) %>% 
  select(label, subfam)

```

```{r define_subfam}

#Define the sub-families of the clusters using the assigned Sal1 members.
# This process works by going through the sub-families present in each cluster,
#   then defining the overall cluster based on the Sal1 present/not present.
#   Thenn filtering to ensure that there's only the one assignment for each cluster.

 #If there are no Sal1 sequences in the cluster say 'unassigned'

cluster_subfam.df <- node.df %>% 
  ungroup %>% 
 #Set non-Sal1 to 'unassigned', so that clusters only made up of them are named such.
  mutate(subfam = ifelse(subfam %in% c('assembled','P01','Zhu16'),
                         'unassigned',
                         subfam)) %>% 
  dplyr::group_by(X30pid25lngthI1.4, subfam) %>% 
  dplyr::summarise(num = n_distinct(label)) %>% 
  group_by(X30pid25lngthI1.4) %>% 
  mutate(total_subfam = n_distinct(subfam),
         total_pirs = sum(num)) %>% 
  group_by(X30pid25lngthI1.4, total_pirs) %>% 
  #Only keep 'unassigned'/'Sal1' if it is the only subfamily present (i.e. entire cluster is unassigned).
  # Unless only other subfam in cluster is unassigned/Sal1, then I want this to be unassigned.
  filter(!(subfam %in% c('unassigned', 'Sal1') & total_subfam > 1) | all(subfam %in% c('unassigned','Sal1'))) %>% 
  #Calculate the percentage the subfams constitute of each subfam to see if any groups are split between sub-families.
  mutate(subfam_perc = round(100*num/sum(num), digits = 2)) %>% 
  #Keep the subfam name if it makes up over 50% of the number of subfamily members in the cluster.
  # This will now become the subfam assignment of the cluster.
  # If there's no dominating subfam then assign it as mixed.
  mutate(mixed = ifelse(all(100*num/sum(num) <= 50), 'mixed', 'dom')) %>% 
  summarise(subfam_num = paste0(subfam, ' (n = ', num, ')', collapse = '; '),
            subfam_clust = case_when(
              #Need convert the genes in no clusters into it's own noCluster group. 
              # (not part of the BLAST network/no alignments)
              # convert to integer and character because being a factor makes it tricky
              as.integer(as.character(X30pid25lngthI1.4)) == 0 ~ 'noCluster',
              100*num/sum(num) > 50 ~ subfam,
              mixed == 'mixed' ~ 'mixed',
              TRUE ~ ''
            )
  ) %>% 
  #Remove the rows with an empty subfam_clust (grep for a letter of the alphabet) 
  filter(grepl(subfam_clust, pattern = '[[:alpha:]]')) %>% 
  #Use unique to remove the redundant rows this creates for the 'mixed' samples
  unique %>% 
  mutate(clust_num = X30pid25lngthI1.4)

#Uniquify for numbering the subfam
cluster_subfam.df$subfam <- uniquify(cluster_subfam.df$subfam_clust, 
                           first_one = TRUE)
# Create column subfam_assignment to make code later easier since node.melt already has 'subfam' column
cluster_subfam.df$subfam_assignment <- uniquify(cluster_subfam.df$subfam_clust, 
                           first_one = TRUE)

# write.csv(cluster_subfam.df,
#           file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/cluster_subfam_30pid25lngthI1.4.csv',
#           row.names = FALSE,
#           quote = FALSE)

```

```{r write_edge.csv}

# blast_30pid25lngth.df <- read.delim('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/tpmOver1/all-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1_all2allBlast_30pid25lngthNoSame2Same.blast',
#            header = FALSE)
# 
# node_labels.df <- blast_30pid25lngth.df %>%
#   rowwise() %>%
#   mutate(label1 = node.df$Id[node.df$label %in% V1],
#          label2 = node.df$Id[node.df$label %in% V2]) %>%
#   select(V1,V2,label1, label2)
# 
# node_labels_noDup.df <- node_labels.df[!duplicated(apply(node_labels.df,
#                                                          sort,
#                                                          MARGIN =1),
#                                                    MARGIN=2),]
# #Already appears to have same2same matches absent.
# 
# node_labels_noDup_toSave.df <- select(node_labels_noDup.df, label1, label2)
# colnames(node_labels_noDup_toSave.df) <- c('Source', 'Target')

# write.csv(node_labels_noDup_toSave.df,
#           file = '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/networks/tpmOver1/edges_30pid25lngthNoS2S.csv',
#           row.names = FALSE,
#           quote = FALSE)

```

#Want to show the stages of development in this.

```{r expt_name}

#Extract the first two numbers of the name e.g. ERX2560125_ERR2541707
stage_geog.df <- bind_rows(
  lapply(
    assembly.df$assembly, 
    function(assembly_name){
      accession_name <- str_extract(assembly_name, 
                                    pattern = '(SR|ER).{8,9}_(SR|ER).{8,9}(?=_)')
      #If this works (not NA and only two elements) then find the expt name in the table
      if(!is.na(accession_name) & length(unlist(strsplit(accession_name, '_'))) == 2){
        cat_sample_stage <- accession_table$Cat.Sample.stage.s.[grepl(accession_table$fastq_1, pattern = accession_name)]
        geog_source <- accession_table$Geography_source[grepl(accession_table$fastq_1, pattern = accession_name)]
      } else {
        #Get the first two elements of the name (separated by '_') - split then paste back together.
        expt_name <- paste(str_split_fixed(assembly_name, 
                                           pattern = '_', n = 3)[c(1,2)], collapse = '_')
        if(any(grepl(accession_table$Concatenated.file, 
                     pattern = expt_name))){
          #Otherwise this is probably a re-named concatenated file of samples, so
          # we need the 'Concatenated file' column.
          cat_sample_stage_all <- accession_table$Cat.Sample.stage.s.[grepl(accession_table$Concatenated.file, 
                                                                            pattern = expt_name)]
          geog_source_all <- accession_table$Geography_source[grepl(accession_table$Concatenated.file, 
                                                                    pattern = expt_name)]
        } else {
          #Instead of being sensible I had a different naming scheme for the un-concatenated Kim and Gunalan files.
          cat_sample_stage_all <- accession_table$Cat.Sample.stage.s.[grepl(accession_table$file_prefix, 
                                                                            pattern = expt_name)]
          geog_source_all <- accession_table$Geography_source[grepl(accession_table$file_prefix, 
                                                                    pattern = expt_name)]
        }

        cat_sample_stage <- unique(cat_sample_stage_all)
        geog_source <- unique(geog_source_all)

        if(rlang::is_empty(cat_sample_stage) | rlang::is_empty(geog_source)){
          stop(paste0('No life cycle stage or geographical source name for assembly ', assembly_name))
        }

        if(!length(cat_sample_stage) == 1){
          cat_sample_stage <- 'Mixed.stages' 
        }

        if(!length(geog_source) == 1){ 
          geog_source <- 'Mixed.sources'
        }

      }
      return(data.frame(assembly = assembly_name, stage = cat_sample_stage, geog_source = geog_source))
    })
)
```

# Spread of experiments

```{r expt_makeup_of_clusters}

node.melt <- node.df %>% 
  setDT %>% 
  melt(measure.vars = patterns('lngth'), 
       variable.name = 'mcl', 
       value = 'clust_num') %>% 
  mutate(mcl = as.character(mcl),
         expt = ifelse(expt == 'known',
                       paste0(source,'_known'),
                       expt_num.df$expt_code[match(expt, expt_num.df$expt)])) %>% 
  inner_join(., cluster_subfam.df, by = 'clust_num')

safe_colorblind_palette_expt <- c(expt_num.df$col[1:9],
                                  hcl.colors(3, palette = "Lajolla")[1:3])
names(safe_colorblind_palette_expt) <- c(expt_num.df$expt_code[1:9],
                                         unique(node.melt$expt)[
                                           grepl(unique(node.melt$expt), 
                                                 pattern = 'known')
                                         ])

lapply(unique(node.melt$mcl), 
       function(mcl_name){

         node.melt.df <- filter(node.melt, 
                                mcl == paste(mcl_name)) %>% 
           group_by(clust_num, mcl) %>% 
           mutate(clust_size = paste0(clust_num, ' (n=', 
                                     n_distinct(label), ')')) %>% 
           group_by(expt, clust_num, clust_size) %>% 
           filter(clust_num %in% 1:20) %>% 
           dplyr::summarize(num = n_distinct(label)) %>% 
           arrange(clust_num)

         node.melt.df$clust_size <- factor(node.melt.df$clust_size,
                                           levels = unique(node.melt.df$clust_size))
         node.melt.df$expt <- factor(
           node.melt.df$expt,
           levels = names(safe_colorblind_palette_expt)[
             names(safe_colorblind_palette_expt) %in% unique(node.melt.df$expt)
           ]
         )

         lapply(c('withKnown','onlyDeNovo'), 
                function(include){

                  if(include == 'onlyDeNovo'){
                    node.melt.df <- filter(node.melt.df, 
                                           !grepl(expt, pattern = 'known'))
                  } else {
                    #Removing the '_known' from the P01 etc names so that the legend fits better.
                    node.melt.df <- mutate(node.melt.df, 
                                           expt = str_remove(expt, pattern = '_known'))
                    names(safe_colorblind_palette_expt) <- str_remove(names(safe_colorblind_palette_expt),
                                                                      pattern = '_known')
                  }

                  plot <- ggplot(data = node.melt.df,
                                 aes(x = clust_size, 
                                     y = num, 
                                     fill = expt)) +
                    geom_col(position = 'fill') +
                    xlab('Cluster number') +
                    ylab('Proportion of assembled pirs\nin each cluster') +
                    scale_fill_manual(values = safe_colorblind_palette_expt[
                      names(safe_colorblind_palette_expt) %in% unique(node.melt.df$expt)
                    ],
                    guide = guide_legend(title = 'Experiment') )+
                    ggtitle(paste(mcl_name)) +
                    theme_classic() +
                    theme(axis.text.x = element_text(angle = 90,
                                                     vjust = 0.5))

                  if(global_save_var){
                    ggsave(plot = plot,
                           paste0('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/networks_exptComposition_',
                                  include,'_',mcl_name,'.pdf'),
                           device = 'pdf',
                           width = 7,
                           height = 5,
                           units = "in")
                  }

                  return(plot)
                })
       })
```

```{r expt_composition}
# How many clusters to include? e.g. 10 is the top ten largest clusters.
# Uses clust_num to subset this, so including 0 will include noCluster.
num_clusters <- 10
clusters_include <- filter(cluster_subfam.df, 
                           clust_num %in% 1:10)$subfam_assignment

#Get standardised colours 
top_subfam_col <- read.csv('top_subfam_col.csv')

node.melt.df <- filter(node.melt, 
                                mcl == 'X30pid25lngthI1.4',
                       expt != 'E05') %>% 
  group_by(subfam_assignment, mcl) %>% 
  mutate(clust_size = paste0(subfam_assignment, ' (n=', 
                             n_distinct(label), ')'),
         clust_numbers = n_distinct(label) ) %>% 
  group_by(expt) %>% 
  mutate(expt = paste0(expt, ' (n=', 
                       n_distinct(label), ')')) %>% 
  group_by(expt, subfam_assignment, clust_size, clust_numbers) %>% 
  filter(subfam_assignment %in% clusters_include) %>% 
  dplyr::summarize(num = n_distinct(label)) %>% 
  arrange(desc(clust_numbers))

node.melt.df <- mutate(node.melt.df, 
                       is_known = str_detect(expt, pattern = '_known'),
                       expt = str_remove(expt, pattern = '_known'))

node.melt.df$clust_size <- factor(node.melt.df$clust_size,
                                  levels = unique(node.melt.df$clust_size))
# node.melt.df$expt <- factor(
#   node.melt.df$expt,
#   levels = names(safe_colorblind_palette_expt)[
#     names(safe_colorblind_palette_expt) %in% unique(node.melt.df$expt)
#   ]
# )

color_palette <- c(
  hcl.colors(num_clusters, 
             palette = 'Spectral')[1:(num_clusters/2)],
  hcl.colors(num_clusters/2, 
             palette = 'viridis')
)
# %>% 
#   alternateReverse()

colours_subfam <- inner_join(node.melt.df, 
                             top_subfam_col, 
                             by = 'subfam_assignment')
top_subfam_col.vec <- setNames(unique(colours_subfam$colours),
                           unique(colours_subfam$clust_size))

( plot <- ggplot(data = node.melt.df,
                 aes(x = expt, 
                     y = num, 
                     fill = clust_size)) +
    geom_col(position = 'fill') +
    xlab('Experiment') +
    ylab(expression(paste('Proportion of ', 
                          italic(pir), 
                          's from each cluster'))) +
    scale_fill_manual(
      values = c( 
        top_subfam_col.vec
      ),
      guide = guide_legend(title = 'Cluster')
    ) +
    theme_classic() +
    facet_grid(~ is_known, scales = 'free_x', space = 'free') +
    theme(axis.text.x = element_text(angle = 90,
                                     vjust = 0.5),
          strip.text.x = element_blank(),
          strip.background = element_blank())
  )

if(global_save_var){
  ggsave(plot = plot,
         paste0('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/networks_clustComp_Expt.pdf'),
         device = 'pdf',
         width = 7,
         height = 8,
         units = "in")
}
```

```{r geog_source}
geog.df <- filter(node.melt, 
                       mcl == 'X30pid25lngthI1.4') %>%
  #Get the geographic source from the stage_geog.df. Need to extract the part of the assembly name without '_scapdes_0.95cdhitest'
  mutate(geog_source = stage_geog.df$geog_source[match(source, 
                                                       str_extract(stage_geog.df$assembly, 
                                                                           pattern = '.+(?=_scspades)'))] ) %>% 
  group_by(clust_num, mcl) %>% 
  mutate(clust_size = paste0(clust_num, ' (n=', 
                             n_distinct(label), ')')) %>% 
  group_by(expt) %>% 
  mutate(expt_size = n_distinct(label)) %>% 
  group_by(clust_num, clust_size, geog_source, expt, expt_size) %>% 
  filter(clust_num %in% 1:20,
         !is.na(geog_source)) %>% 
  dplyr::summarize(num = n_distinct(label),
                   num_norm = n_distinct(label)/expt_size
                   ) %>%
  arrange(clust_num) %>% 
  unique()
#For some reason calculating num_norm means we need to use unique

geog.df$clust_size <- factor(geog.df$clust_size,
                                           levels = unique(geog.df$clust_size))

ggplot(data = filter(geog.df, !clust_num %in% c(1,18)),
                                 aes(x = clust_size, 
                                     y = num_norm, 
                                     fill = geog_source)) +
                    geom_col(position = 'stack', col = 'black') +
                    xlab('Cluster number') +
                    ylab('Number of assembled pirs\nin each cluster normalized by total expt num') +
                    # scale_fill_manual(values = safe_colorblind_palette_expt[
                    #   names(safe_colorblind_palette_expt) %in% unique(node.melt.df$expt)
                    # ],
                    # guide = guide_legend(title = 'Experiment') )+
                    # ggtitle(paste(mcl_name)) +
                    theme_classic() +
                    theme(axis.text.x = element_text(angle = 90,
                                                     vjust = 0.5))
```

Generally the assembled pirs are pretty well spread out throughout the top 10 clusters.

## What's the largest cluster made up of 100% from one experiment?

```{r largest_100pc_cluster}
lapply(unique(node.melt$mcl), 
       function(mcl_name){
filter(node.melt, 
                mcl == paste(mcl_name)) %>% 
  group_by(clust_num) %>% 
  mutate(clust_num_num = sum(n_distinct(label))) %>% 
  group_by(expt, clust_num, clust_num_num, mcl) %>% 
  dplyr::summarize(prop = 100*round(n_distinct(label)/clust_num_num, digits = 3)) %>% 
   # Why do I need to call unique instead of summarise() doing it properly?
   unique %>% 
  arrange(desc(prop), desc(clust_num_num)) %>% 
  .[1:20,]
       })
```

The largest clusters 100% from one experiment are only five nodes in size. Cluster 25 (n=9) is entirely P01.

```{r clusters_dominated_by_expt}
lapply(unique(node.melt$mcl), 
       function(mcl_name){
         filter(node.melt, 
                mcl == paste(mcl_name)) %>% 
  group_by(clust_num) %>% 
  mutate(clust_num_num = sum(n_distinct(label))) %>% 
  filter(clust_num_num > 10) %>% 
  group_by(expt, clust_num, clust_num_num) %>% 
  dplyr::summarize(prop = 100*round(n_distinct(label)/clust_num_num, digits = 3)) %>% 
   # Why do I need to call unique instead of summarise() doing it properly?
   unique %>% 
  arrange(desc(prop), desc(clust_num_num)) %>% 
  .[1:20,]})
```

There are 14 clusters over ten nodes in size represented >=50% by only one experiment. Most of these are from the 'known' set of pre-existing pirs (P01, Sal1, Zhu16 original assembly).

## How much of each expt is found in the top N clusters?

```{r expt_num_across_clusters, eval = FALSE}
lapply(unique(node.melt$mcl), 
       function(mcl_name){
         node_mcl.df <- filter(node.melt, 
                               mcl == paste(mcl_name),
                               !grepl(expt, pattern = 'known')) %>% 
           mutate(expt = factor(expt)) %>% 
           group_by(expt) %>% 
           mutate(expt_num = sum(n_distinct(label)))
         num_clust <- length(unique(node_mcl.df$clust_num))
         expt_prop_across_clusters <- lapply(c(0:9,seq(10,num_clust + 5,1)),
                                             function(cluster_name){
                                               # Need to convert expt to a factor so that complete() later will add the missing values for large cluster number.
                                               # Also need a special case for when using 0 clusters, to get the lines to start at 100%.
                                               if(cluster_name != 0){node_mcl.df <- filter(node_mcl.df,
                                                                                     !clust_num %in% 1:cluster_name)} 
                                                 node_mcl.df %>%  
                                                 group_by(expt) %>% 
                                                 mutate(expt_num_noTen = sum(n_distinct(label))) %>% 
                                                 summarize(prop = 100*round(expt_num_noTen/expt_num, digits = 3)) %>% 
                                                 unique %>% 
                                                 ungroup %>%  
                                                 tidyr::complete(expt, fill = list(prop = 0)) %>% 
                                                 mutate(num_clusters = cluster_name)
                                             }) %>% bind_rows()

         plot <- ggplot(data = expt_prop_across_clusters,
                aes(y = prop, x = num_clusters, col = expt, linetype = expt)) +
           geom_step() + 
           scale_color_manual(values = safe_colorblind_palette_expt[
             names(safe_colorblind_palette_expt) %in% unique(expt_prop_across_clusters$expt)
           ],
           labels = names(safe_colorblind_palette_expt)[
             names(safe_colorblind_palette_expt) %in% unique(expt_prop_across_clusters$expt)
           ],
           guide = guide_legend(ncol = 2)) +
           scale_linetype_manual(values = c(1,1,1,1,1,3,3,2,2),
                                 labels = names(safe_colorblind_palette_expt)[
                                   names(safe_colorblind_palette_expt) %in% unique(expt_prop_across_clusters$expt)
                                 ]) +
           theme_classic()

         if(global_save_var){
           ggsave(plot = plot,
                  paste0('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/networks_exptStep_'
                         ,mcl_name,'.pdf'),
                  device = 'pdf',
                  width = 7,
                  height = 5,
                  units = "in")
         }

         return(plot)
       })

```

# Spread of sub-family with clusters

Note that some clusters are missing because no Sal1 is present within them e.g. 15.

```{r subfam_makeup_of_clusters}

cbf_1 <- c("#999999", "#E69F00", "#56B4E9", "#009E73", 
           "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
names(cbf_1) <- c("B","C","E","I","G","J","K", "unassigned")
cbf_1['unassigned']<- 'black'

lapply(unique(node.melt$mcl), 
       function(mcl_name){
         mcl_num_subfamNum.df <- filter(node.melt, 
                                        mcl == paste(mcl_name)) %>% 
           #Make clust_num numeric for correct ordering by number.
           mutate(clust_num = as.numeric(clust_num),
                  subfam = ifelse(subfam == 'Sal1', 
                                  'unassigned',
                                  subfam)) %>% 
           group_by(subfam, clust_num) %>% 
           filter(clust_num %in% 1:25, 
                  source == 'Sal1') %>%
           dplyr::summarize(num = n_distinct(label))

         plot <- ggplot(data = mcl_num_subfamNum.df,
                        aes(x = clust_num, y = num, fill = subfam)) +
           geom_col(position = 'stack') +
           scale_fill_manual(values = cbf_1,
                             guide = guide_legend(title = 'Sal1 Sub-family')) +
           #Set scale_x_cont here so that all cluster numbers are shown.
           scale_x_continuous(breaks = 1:25) +
           scale_y_continuous(n.breaks = 10) +
           xlab('Cluster number') +
           ylab('Number of pirs') +
           ggtitle(mcl_name) +
           theme_classic()

         if(global_save_var){
           ggsave(plot = plot,
                  paste0('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/networks_sal1Composition_',
                         mcl_name,'.pdf'),
                  device = 'pdf',
                  width = 7,
                  height = 5,
                  units = "in")
         }

           return(plot)
       })
```

```{r pir_sizes}

node_length.melt <- filter(node.melt, 
       mcl == 'X30pid25lngthI1.4') %>% 
  rowwise() %>% 
  mutate(orf_length = length(getElement(fasta, label)))

#Can remove all ORFs under 200aa to test how the short 'known' pirs impacts the results.
# Spoiler: They don't.
# %>% 
#   filter(orf_length >= 200)

#Select first 11 clusters and remove the short Zhu16 seqeunces.
# Order by median length and factor so that ggplot will show in this order.
node_length_clust1.11_noZhu16.melt <- filter(node_length.melt, 
                                             clust_num %in% 1:11, 
                                             expt != 'Zhu16_known') %>% 
  group_by(clust_num) %>% 
  mutate(orf_length_med = median(orf_length)) %>% 
  arrange(desc(orf_length_med)) %>% 
  ungroup() %>% 
  # ungroup() so that factor works properly and doesn't sort clust_num
  mutate(clust_num = factor(clust_num, 
                            levels = unique(clust_num)))

node_length_clust1.11_noZhu16.melt %>% 
  group_by(clust_num) %>% 
  summarise(med_length = median(orf_length))

(plot_orf <- ggplot(node_length_clust1.11_noZhu16.melt, 
                    aes(x = clust_num,
                        y= orf_length)) +
    geom_boxplot() +
    xlab('Cluster number') +
    ylab('Length of ORFs (aa)') +
    theme_classic() +
    theme(axis.text.x = element_text(size = 13),
          axis.text.y = element_text(size = 13)) )

if(global_save_var){
  ggsave(plot = plot_orf,
         '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/plots/networks_ORFlength.pdf',
         device = 'pdf',
         width = 7,
         height = 5,
         units = "in")
}

```

```{r orfLengths}
ggplot(filter(node_length.melt, clust_num %in% c(1,12:20), 
              expt != 'Zhu16_known'), 
       aes(x = factor(clust_num),
           y= orf_length)) +
  geom_boxplot()
ggplot(filter(node_length.melt, clust_num %in% 1:11), 
       aes(x = factor(clust_num),
           y= orf_length)) +
  geom_boxplot() + facet_wrap(~expt)

ggplot(filter(node_length.melt, clust_num %in% 1:11), 
       aes(x = factor(expt),
           y= orf_length)) +
  geom_boxplot()

lapply(unique(node_length.melt$expt), function(expt){
  lengths <- node_length.melt$orf_length[node_length.melt$expt == expt]
  if(length(lengths) > 2){
    hist(lengths, 
         breaks = 100,
         main = expt,
         xlim = c(0,900))
  }
})
lapply(1:11, function(clust_num){
  lengths <- node_length.melt$orf_length[node_length.melt$clust_num == clust_num]
  if(length(lengths) > 2){
        hist(lengths, 
         breaks = 100,
         main = paste0('Cluster Number ', clust_num),
         xlim = c(0,900))
  }
})
```

Cluster 8 particularly long?

Looks like there's a little group of longer pirs for some experiments/P01/Sal1.
Like over 800aa - pretty much all in Cluster 1 (E). 
I wonder if this is the two groups in the Gephi network.

Cluster 7 and 9 low variability, potentially shorter?

```{r kwtest_length}
#K-W test of if there is a signif difference between the clusters for length

kruskal.test( node_length.melt$orf_length, factor(node_length.melt$clust_num))
kruskal.test(node_length_clust1.11_noZhu16.melt$orf_length,
             factor(node_length_clust1.11_noZhu16.melt$clust_num))

clust_comp <- combn(1:11, m = 2, simplify = FALSE)

#Do the tests for each combination, combione into a data frame and use p.adjust multiple testing adjustment.
( length_tests.df <- lapply(c('greater', 'less'), 
                            function(wilcox_text_alt){
                              lapply(clust_comp, function(comparison){
  test.stat <- wilcox.test(orf_length ~ clust_num,
                           data = filter(node_length.melt,
                                         expt != 'Zhu16_known'),
                           subset = clust_num %in% comparison,
                           alternative = wilcox_text_alt)
  return(data.frame(comparison_1 = comparison[1],
                    comparison_2 = comparison[2],
                     stat = test.stat$statistic,
                     pval = test.stat$p.value,
                    wilcox_text.alt = wilcox_text_alt))
                              }) %>% bind_rows
}) %>% 
  bind_rows %>% 
  mutate(padj = round(p.adjust(pval), digits = 5),
         pval = round(pval, digits = 5)) %>% 
  arrange(desc(padj)) )

length_med.df <- node_length.melt %>% 
  group_by(clust_num) %>% 
  summarise(orf_length_med = median(orf_length))

length_tests.mat <- mutate(length_tests.df,
                           orf_ratio = ifelse(padj <= 0.05, 
                                              length_med.df$orf_length_med[
                                                match(comparison_1, 
                                                      length_med.df$clust_num)
                                              ]/length_med.df$orf_length_med[
                                                match(comparison_2, 
                                                      length_med.df$clust_num)
                                              ], 
                                              NA)) %>% 
  dcast(.,
        comparison_1 ~ comparison_2,
        value.var = 'orf_ratio')
rownames(length_tests.mat) <- length_tests.mat[,1]
length_tests.mat <- length_tests.mat[,-1]

# nams <- unique(unlist(clust_comp))
# 
# #Need to add the first and last rows due to combn not using 1 as combination_2 and 11 as combination_1.
# one.mat <- matrix(nrow = nrow(length_tests.mat),
#        dimnames = list(rep(NA, length.out = nrow(length_tests.mat)), first(nams))) 
# length_tests.mat <- cbind(one.mat, length_tests.mat)
# 
# last.mat <- matrix(ncol = ncol(length_tests.mat),
#        dimnames = list(last(nams), rep(NA, 
#                                        length.out = ncol(length_tests.mat))))
# colnames(last.mat) <- names(length_tests.mat)
# length_tests.mat <- rbind(length_tests.mat, last.mat)
# 
# library(ComplexHeatmap)
# library(viridis)
# #Make matrix symmetrical
# length_tests.mat[
#   lower.tri(length_tests.mat)
# ] <- 1/(t(length_tests.mat)[
#   lower.tri(length_tests.mat)
# ])
# 
# colnames(length_tests.mat) <- rownames(length_tests.mat) <- nams
# 
# Heatmap(length_tests.mat, 
#         cluster_columns = FALSE, 
#         cluster_rows = FALSE,
#         na_col = 'black',
#         col = circlize::colorRamp2(c(0, 1, 1.4), c('blue', 'white', 'red')))

length_tests.df %>%
  filter(if_any(matches('comparison'),
                ~grepl(.x,
                       pattern = '^1$|3|8|10'))) %>%
  arrange(desc(padj))

length_tests.df %>%
  filter(if_any(matches('comparison'),
                ~grepl(.x,
                       pattern = '7'))) %>%
  arrange(desc(padj))
length_tests.df %>%
  filter(if_any(matches('comparison'),
                ~grepl(.x,
                       pattern = '9'))) %>%
  arrange(desc(padj))
```

```{r where_ancestral}

outfmt6_headers <- 
  'qseqid sseqid pident length mismatch gapopen qstart qend sstart send evalue bitscore qcovs qcovhsp slen nident' %>% 
  strsplit(split = ' ') %>% 
  unlist

  #Read in BLAST file 
blast_30pid25lngth.df <- read.delim('/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/hmmer/scspades/G18_B20_C19_Z16_M19_R20_K1719_G19-cat/tpmOver1/all-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1_all2allBlast_30pid25lngthNoSame2Same.blast',
           col.names = outfmt6_headers) %>% 
  mutate(
    scovhsp = round( (nident/slen * 100), digits = 2),
    scovs = round(((abs(send - sstart) + 1)/slen *100), digits = 2),
    evalue = format(evalue, scientific = TRUE)
  )

#Ancestral: PVP01_1147700, PVX_113230
#Vir14:

node.melt %>% 
  filter(mcl == 'X30pid25lngthI1.4',
         label %in% c('PVP01_1147700', 'PVX_113230', 'PVX_108770'))

#All in cluster 4

ancestral_blast <- blast_30pid25lngth.df %>% 
  filter(qseqid %in% 
           c('PVP01_1147700', 'PVX_113230') | sseqid %in% 
           c('PVP01_1147700', 'PVX_113230'))
hist(ancestral_blast$pident, breaks = 100)
hist(ancestral_blast$scovhsp, breaks = 100)
hist(ancestral_blast$qcovhsp, breaks = 100)

ancestral_blast_filter <- filter(ancestral_blast,
                          qcovhsp > 50,
                          scovhsp > 50,
                          nident > 80) 
# ancestral_blast_filter <- filter(ancestral_blast, 
#                                  grepl(qseqid, pattern = 'Zhu') | grepl(sseqid, pattern = 'Zhu')) %>% 
#   arrange(pident)
ancestral_blasters <- unique(c(ancestral_blast_filter$qseqid, ancestral_blast_filter$sseqid))

node.df$ancestral_blast <- ifelse(node.df$label %in% ancestral_blasters,
                                  TRUE,
                                  FALSE)
filter(node.df, ancestral_blast) %>% 
  mutate(clust_num = X30pid25lngthI1.4) %>% 
  select(!contains('lngth'))

```

```{r}
vir14_blast <- blast_30pid25lngth.df %>% 
  filter(qseqid %in% 
           c('PVX_108770') | sseqid %in% 
           c('PVX_108770'))
hist(vir14_blast$pident, breaks = 100)
hist(vir14_blast$scovhsp, breaks = 100)
hist(vir14_blast$qcovhsp, breaks = 100)

vir14_blast_filter <- filter(vir14_blast,
                          qcovhsp > 50,
                          scovhsp > 50,
                          nident > 80) 
# vir14_blast_filter <- filter(vir14_blast, 
#                                  grepl(qseqid, pattern = 'Zhu') | grepl(sseqid, pattern = 'Zhu')) %>% 
#   arrange(pident)
vir14_blasters <- unique(c(vir14_blast_filter$qseqid, vir14_blast_filter$sseqid))

node.df$vir14_blast <- ifelse(node.df$label %in% vir14_blasters,
                                  TRUE,
                                  FALSE)
filter(node.df, vir14_blast) %>% 
  mutate(clust_num = X30pid25lngthI1.4) %>% 
  select(!contains('lngth'))

```