### 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_VvxBUSCO.Rmd
Add many required vivax files and change paths in vivax scripts

---
title: "tl22-01_VvxBUSCO"
output: 
  html_document:
    theme: cosmo
    code_download: true
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      results = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
options(scipen = 999) 
# Means that numbers are displayed normally not as 1e6 (1000000)
```

```{r loading_packages, include = FALSE, message = FALSE}
# Load the required packages
library('readr')
library('dplyr')
library('stringr')
library('purrr')
library('kableExtra')
library('plotly')
library('rmarkdown')
library('seqinr')
library(plotly)
```

```{r read_busco}
samplesheet <- read.delim("samplesheet.txt")

#Get the number of pirs stats
assembly_stats.df <- read.table('tl22-01_assembly_info_extra.txt')

colours_val <- unique(assembly_stats.df$expt_colour)
names(colours_val) <- unique(assembly_stats.df$experiment_num)
colours_val <- colours_val[match(sort(unique(assembly_stats.df$experiment_num)),
                                 names(colours_val))]

assembly_stats_busco <- read.table('tl22-01_assembly_info_extra_busco.txt')

#Include '.+(?=0.95)' pattern for searching as Gunalan has multiple cdhitest threshold assemblies.
samplesheet_busco.df <- mutate(rowwise(assembly_stats_busco), 
                               num_pir_predict = assembly_stats.df$num_uniq_mRNA_ORFs_over200aaLength[
                                 str_detect(assembly_stats.df$assembly, pattern = paste0(assembly_name, 
                                                                                         '.+(?=0.95)'))
                               ],
                               expt_num = unique(
                                 assembly_stats.df$experiment_num[
                                   str_detect(assembly_stats.df$assembly, pattern = paste0(assembly_name, 
                                                                                           '.+(?=0.95)'))
                                 ]),
                               num_transcripts = unique(
                                 assembly_stats.df$num_transcripts[
                                   str_detect(assembly_stats.df$assembly, pattern = paste0(assembly_name, 
                                                                                           '.+(?=0.95)'))
                                 ]),
                               num_mRNA_pir_over200aa_over1TPM = unique(
                                 assembly_stats.df$num_mRNA_pir_over200aa_over1TPM[
                                   str_detect(assembly_stats.df$assembly, pattern = paste0(assembly_name, 
                                                                                           '.+(?=0.95)'))
                                 ])
)
```

```{r plotly}
hline <- function(y = 0, color = "blue") {
  list(
    type = "line", 
    x0 = 0, 
    x1 = 1, 
    xref = "paper",
    y0 = y, 
    y1 = y, 
    line = list(color = 'black')
  )
}
vline <- function(x = 0, color = "red") {
  list(
    type = "line", 
    y0 = 0, 
    y1 = 1, 
    yref = "paper",
    x0 = x, 
    x1 = x, 
    line = list(color = color)
  )
}

plot_ly(samplesheet_busco.df,
        x = ~perc_busco_api_complete,
        y = ~perc_busco_api_nonMissing,
        text = ~assembly_name,
        color = ~Experiment,
        colors = rainbow(10),
        mode = 'markers') %>% 
  layout(shapes = list(hline(50)))

plot_ly(samplesheet_busco.df,
        x = ~perc_busco_plasmo_complete,
        y = ~perc_busco_plasmo_nonMissing,
        text = ~assembly_name,
        color = ~Experiment,
        colors = rainbow(10),
        mode = 'markers') %>% 
  layout(shapes = list(hline(50)))

plot_ly(samplesheet_busco.df,
        y = ~perc_busco_api_nonMissing,
        x = ~perc_busco_plasmo_nonMissing,
        text = ~assembly_name,
        color = ~expt_num,
        colors = colours_val,
        size = ~num_transcripts,
        # sizes = c(10,100),
        mode = 'markers') %>% 
  layout(shapes = list(hline(25), vline(25)))

plot_ly(samplesheet_busco.df,
        y = ~num_pir_predict,
        x = ~perc_busco_plasmo_nonMissing,
        text = ~assembly_name,
        color = ~expt_num,
        colors = colours_val,
        mode = 'markers')

```

```{r ggplot_busco_graph}

( plot_buscos <- ggplot(samplesheet_busco.df,
                      aes(x = perc_busco_plasmo_nonMissing,
                          y = perc_busco_api_nonMissing,
                          col = factor(expt_num, unique(sort(expt_num))),
                          size = num_pir_predict)) +
  geom_point(alpha = 0.75) +
  theme_classic() +
    theme(axis.text.x = element_text(size = 10),
          axis.text.y = element_text(size = 10),
          axis.title.x = element_text(size = 12),
          axis.title.y = element_text(size = 12)) +
  ylim(0,100) +
  scale_color_manual(values = colours_val,
                     name = 'Experiment') +
    guides(color = guide_legend(override.aes = list(size=5),
                             ncol = 2)) +
  scale_size_continuous(name = expression(paste('Num predicted ',italic(pir),'s'))) +
  xlab('Percentage of Plasmodium BUSCOs') +
  ylab('Percentage of Apicomplexan BUSCOs') )

ggsave(plot = plot_buscos,
       'plots/assemblies_buscoPlasmo-buscoApi-pirNum.pdf',
       device = 'pdf',
       width = 7,
       height = 7,
       units = "in")

ggplot(samplesheet_busco.df,
       aes(x = perc_busco_plasmo_nonMissing,
           y = num_mRNA_pir_over200aa_over1TPM,
           col = factor(expt_num, unique(sort(expt_num))),
           size = num_pir_predict)) +
  geom_point(alpha = 0.75) +
  theme_classic() +
  scale_color_manual(values = colours_val,
                     name = 'Experiment') +
  scale_size_continuous(name = 'Number of predicted pirs') +
  xlab('Percentage of Plasmodium BUSCOs') +
  ylab('Number of pir transcripts')

ggplot(samplesheet_busco.df,
       aes(x = perc_busco_plasmo_nonMissing,
           y = num_transcripts,
           col = factor(expt_num, unique(sort(expt_num))),
           size = num_pir_predict)) +
  geom_point(alpha = 0.75) +
  theme_classic() +
  scale_color_manual(values = colours_val,
                     name = 'Experiment') +
  scale_size_continuous(name = 'Number of predicted pirs') +
  xlab('Percentage of Plasmodium BUSCOs') +
  ylab('Number of transcripts')

```