### 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_VvxTPMover1.Rmd

---
title: "tl22-01_VvxTPMover1"
output: 
  html_document:
    theme: cosmo
    code_download: true
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---

# Intro

This script analyses the results of a _de novo_ assembly of Plasmodium vivax RNAseq data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      results = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
options(scipen = 999) 
# Means that numbers are displayed normally not as 1e6 (1000000)
```

```{r loading_packages, include = FALSE, message = FALSE}
# Load the required packages
library('readr')
library('dplyr')
library('stringr')
library('purrr')
library('kableExtra')
library('plotly')
library('rmarkdown')
library('seqinr')

safe_colorblind_palette <- c("#88CCEE", "#CC6677", "#DDCC77", "#117733", "#332288", "#AA4499", 
                             "#44AA99", "#999933", "#882255", "#661100", "#6699CC", "#888888")
global_save_var <- FALSE
```

```{r function_ordering}
ordering <- function(to_order, order_vector, fromLast = TRUE){
  order = unlist(sapply(order_vector, 
                        function(x){unique(str_subset(to_order, 
                                                      pattern = paste(x)))}))
  order_missing = unique(to_order[!to_order %in% order])
  unique(c(order, order_missing), fromLast = fromLast)
}
```

```{r function_uniquify}
#Make unique identifiers from a vector of names by appending '_X' where X is the number of the 'replicate'.
uniquify <- function(vector_names, first_one = TRUE){
  new_vector <- c()
  df_num <- data.frame(names = unique(vector_names), rep = 1)
  if(first_one){
    duplicated_names <- unique(vector_names[duplicated(vector_names)])
  }else{
    duplicated_names <- c()
  }
  for(i in 1:length(vector_names)){
    if(vector_names[i] %in% duplicated_names|!first_one){
      name <- paste(vector_names[i],
                    df_num$rep[match(vector_names[i], df_num$names)],
                    sep = '_')
      #Increase rep counter
      df_num$rep[match(vector_names[i], df_num$names)] <- df_num$rep[match(vector_names[i], df_num$names)] + 1
      new_vector[i] <- name
    }else{
      new_vector[i] <- vector_names[i]
    }
  }
  new_vector
}
```

```{r read_200aa_pirs}
G18G19_hmmerVir200aa.fa <- read.fasta('G18_B20_C19_Z16_M19_R20_K1719_G19-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_200aa.fasta')

#Need to read in the >200aa ORF length  pirs
pir_list.vec <- str_extract(
  names(G18G19_hmmerVir200aa.fa),
  pattern = 'NODE.+(?=_[[:digit:]])'
)
```

```{r load_salmon}
#Get the full file paths of the salmon quant.sf files.
quant_files <- list.files(path = "salmon/all_files",
                          recursive = TRUE,
                          pattern = 'quant.sf',
                          full.names = TRUE)

#Load the samplesheet.txt file of metadata, add new column for salmon files
samplesheet <- read.delim("samplesheet.txt") %>% 
  mutate(salmon_quant = '')

#Make a table of the files and the sample metadata from samplesheet.txt
# for loop over all the quant_files that exist
for(quant_name in quant_files){
  #Looking for the sample prefix in the file path between 'all_files' and '{prefix}_quant'
  prefix <- str_extract(quant_name, 
                        pattern = '(?<=\\/all_files/).+(?=/.+_quant)')
  if(nrow(samplesheet[grep(samplesheet$prefix, 
                   pattern = prefix),]) > 1){
                     stop(paste0(prefix,
                                 " matches more than one row of the samplesheet"))
                   }
  samplesheet[grep(samplesheet$prefix, 
                   pattern = prefix),'salmon_quant'] <- quant_name
}

sum(grepl(samplesheet[,'salmon_quant'], pattern = 'camp')) == 247
```

```{r list_expt_quant}
expt_quant <- lapply(unique(samplesheet$Experiment),
                     #For each experiment name
                     function(expt){
                       print(paste0('Starting ', expt))
                       #Get the experimental metadata by subsetting
                       expt_data <- samplesheet[samplesheet$Experiment == expt,]
                       if(!sum(grepl(expt_data$assembly, pattern = '.')) == 0){
                         #If sum of grepping for any character is not zero, do the analysis 
                         # i.e there are assembly names and probably assemblies/quant files
                         # For each individual assembly 
                         # (we can bind rows for samples that use the same assembly)
                         assembly_quants <- lapply(unique(expt_data$assembly), 
                                                   function(assembly){
                                                     #Get the prefix(es) that were quantified against this assembly
                                                     prefix <- expt_data[expt_data$assembly == assembly, 
                                                                         'prefix']
                                                     #Get the quant file name(s) and read it in
                                                     quant_file <- expt_data[expt_data$prefix %in% prefix,
                                                                             'salmon_quant']

                                                     #If the file name isn't zero length (i.e. there is no name)
                                                     if(!length(quant_file) == 0){
                                                       #If length is one then we can make the quant file for just one sample 
                                                       if(length(prefix) == 1){
                                                         quant <- read.delim(quant_file)
                                                         #Subset the Name and TPM columns and keep it as a data frame (drop = F)
                                                         quant.df <- quant[,c('Name','TPM'), drop = FALSE]

                                                         # Give a warning if the TPM values are all zero.
                                                         if(sum(as.double(quant$TPM)) == 0){
                                                           #Do not return a quant.df data.frame, making it nothing:
                                                           quant.df <- ''
                                                           print(paste0(prefix, 
                                                                        " associated file has recorded no TPM values"))
                                                           #empty return to end the lapply for this file
                                                           return()
                                                         }

                                                       } else {
                                                         #If longer than 1 so there's more than one quant file for this assembly
                                                         quant.rlist <- lapply(quant_file, read.delim)
                                                         #Use purrr reduce to bind the df in the list together with inner_join,
                                                         # combining by the 'Name' column which should be the same throughout.
                                                         quant.df <- purrr::reduce(quant.rlist, 
                                                                                   inner_join, 
                                                                                   by = 'Name')
                                                         quant.df <- quant.df[grepl(colnames(quant.df), 
                                                                                    pattern = 'Name|TPM')]

                                                         # Give a warning if the TPM values are all zero.
                                                         if(any(sapply(quant.df[grepl(colnames(quant.df), 
                                                                                      pattern = 'TPM')], 
                                                                       function(x) sum(as.double(x))) == 0)){
                                                           print(paste0("At least some of ", 
                                                                        expt, 
                                                                        " results have recorded no TPM values"))
                                                           return()
                                                         }

                                                       }

                                                       #Put the sample name (prefix) in the TPM column
                                                       colnames(quant.df)[grepl(colnames(quant.df), 
                                                                                pattern = 'TPM')] <- paste0(prefix,
                                                                                                            '_TPM')

                                                       # Mark the pir NODES as such (TRUE in the 'pir' column)
                                                       quant.df$pir <- quant.df$Name %in% pir_list.vec
                                                       print(paste0('Number of pirs = ', sum(quant.df$pir)))
                                                       return(quant.df)
                                                     }
                                                   })

                         names(assembly_quants) <- str_extract(unique(expt_data$assembly),
                                                               pattern = '(?<=\\/cdhitest/).+(?=/.+fasta)')
                         print(paste0('Finished ', expt))
                         return(assembly_quants)  
                       }
                     }
)

names(expt_quant) <- unique(samplesheet$Experiment)
```

```{r node-df_tpmvalues}

# Add the TPM values to the node.df dataset alongside a threshold expressed/not expressed

# Get a list of all the de novo transcripts and their max TPM. 
#  Shouldn't be any with the same name unless by exceptional chance?
pir_quant <- lapply(names(expt_quant), 
                     function(expt.name) {
                       expt.list <- expt_quant[[expt.name]]
                       # print(names(expt.list))
                       lapply(names(expt.list), 
                              function(quant.name) {
                                quant.df <- expt.list[[quant.name]]
                                if(!is.null(quant.df)){
                                  #Melt the dataframe to calculate the max TPM value for each de novo transcript
                                  # across all samples.
                                  #This may be faster without melting and using apply.
                                  filter(quant.df, pir) %>% 
                                    reshape2::melt(.,
                                                   id.vars = c('pir','Name'),
                                                   measure.vars = str_subset(colnames(.), 
                                                                             pattern = 'TPM'),
                                                   value.name = 'tpm') %>% 
                                    group_by(Name, pir) %>% 
                                    summarise(max_tpm = max(tpm)) %>% 
                                    mutate(label_no_end = paste(quant.name, Name, sep = '_'),
                                           expt_assembly = quant.name, 
                                           expt = expt.name)
                                }
                              }) %>% 
                         bind_rows
                     }) %>% 
  bind_rows

# sum(duplicated(pir_quant$Name))

write.csv(pir_quant,
          'pir_maxTPM.csv')

```

```{r pirRemoved}
pir_quant %>% 
  group_by(expt_assembly) %>% 
  summarise(num_over1TPM = sum(max_tpm >= 1),
            num_total = n_distinct(Name),
            perc_over1TPM = round(100*sum(max_tpm >= 1)/n_distinct(Name), digits = 2),
            med_max_over1tpm = median(max_tpm[max_tpm >= 1])) %>% 
  arrange(desc(perc_over1TPM))
```

```{r over1tpm_listFiles}
#Note that node_tpm.df includes none-quantified genes, like P01/Sal1/Zhu16 and the Rangel genes.
# Will use over or equal to 1 TPM in at least one sample (max_tpm > 1) as the threshold for active transcription.
expressed_pirTranscripts <- filter(pir_quant, max_tpm>=1)

paste0('Number of transcribed pirs = ',
       nrow(expressed_pirTranscripts),
       ' and number of non-transcribed pirs = ',
       nrow(pir_quant) - nrow(expressed_pirTranscripts))

names_no_end <- str_remove(names(G18G19_hmmerVir200aa.fa), 
                           pattern = '_[[:digit:]]{1,2}$')

#These aren't all unique, I guess this doesn't impact anything later.

#Subset the fasta file by the pirs which are actively transcribed.
# Note that this may be greater than expected due to the fact that there may be multiple pir ORFs for one expressed pir transcript.
# Also it is smaller than the number of expressed_pirTranscripts because not all of the expressed_pirTranscripts have a >200aa ORF.
# Note that the pirs are defined as such based on the *_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir.list files (so before 200aa filtering).

G18G19_hmmerVir200aa_tpmOver1.fa <- G18G19_hmmerVir200aa.fa[
  names(G18G19_hmmerVir200aa.fa)[names_no_end %in% expressed_pirTranscripts$label_no_end]
  ]

#Duplicates among these ORFs/transcripts.
# duplicates <- sapply(expressed_pirTranscripts$label_no_end,
#        function(x) names(G18G19_hmmerVir200aa.fa)[names_no_end %in% x] )
# duplicates[lapply(duplicates, length) > 1]

print(paste0('Number of ORFs (>200aa ORF tpm over 1) = ',
      length(G18G19_hmmerVir200aa_tpmOver1.fa)))
print(paste0('Number of unique transcripts (>200aa ORF tpm over 1) = ',
      length(unique(str_extract(names(G18G19_hmmerVir200aa_tpmOver1.fa), 
                   pattern = '.+(?=_[[:digit:]]$)')))))

table(str_extract(names(G18G19_hmmerVir200aa_tpmOver1.fa), pattern ='.+(?=_NODE)'))

#Save the fasta file of the over 1 TPM pirs.
# write.fasta(G18G19_hmmerVir200aa_tpmOver1.fa,
#             names(G18G19_hmmerVir200aa_tpmOver1.fa),
#             file.out = 'G18_B20_C19_Z16_M19_R20_K1719_G19-cat_scspades_0.95cdhitest_hmmerPfam-PlasmodiumVir_orf200aa_tpmOver1.fasta')

# sum(expressed_pirTranscripts$Name == 'NODE_2407_length_2160_cov_4.64561')

```

The percentage of overall transcripts and pir transcripts transcribed over 1 TPM for each sample.

```{r num_over_1}
# Count the number/percentage of de novo transcripts with an expression level over 1 TPM
over1.df <- lapply(names(expt_quant), 
       function(expt){
         expt_df.list <- lapply(names(expt_quant[[expt]]), 
                function(assembly){
                  if(!is.null(expt_quant[[expt]][[assembly]])){
                    df <- expt_quant[[expt]][[assembly]]
                    #reshape2::melt the data frame to put each sample (the measure.vars) on a separate row.
                    reshape2::melt(df, 
                         id.vars = c('Name','pir'),
                         measure.vars = str_subset(colnames(df), 
                                                   pattern = 'TPM'),
                         variable.name = 'sample',
                         value.name = 'tpm') %>% 
                      mutate(num_transcripts = n_distinct(Name),
                             num_pirs = n_distinct(Name[pir])) %>% 
                      group_by(sample, pir, 
                               num_pirs, num_transcripts
                      ) %>% 
                      summarise(over1_perc = unique(round(100*sum(tpm >= 1)/num_transcripts, 
                                                          digits = 2)),
                                over1_num = sum(tpm >= 1),
                                #For the pir percantage if it is over 100% (i.e. when this is
                                # calculating for all the transcripts total) then make it NA
                                over1_perc_pirs = ifelse(unique(sum(tpm >= 1)/num_pirs <= 1),
                                                         round(unique(100*sum(tpm >= 1)/num_pirs),
                                                               digits = 2),
                                                         NA),
                                total = n_distinct(Name)) %>% 
                      mutate(assembly = paste(assembly))
                  }
                })
         #Add the experiment as a row for binding all these together
         expt_df <- bind_rows(expt_df.list) %>% 
           mutate(expt = expt)
       }) %>% 
  #bind_rows together across all the experiments and assemblies.
  bind_rows

```

```{r assembly_info}

assembly_info.df <- read.delim('tl22-01_assembly_info.txt') %>% 
  filter(!grepl(assembly, pattern = '0\\.9c|0\\.99'),
         !grepl(assembly, pattern = 'SRX5158513_SRR8347514|SRX5158512_SRR8347515'))  
#Filter out any none-0.95 assembly made. And potential Kim19_A duplicate?

#Order experiments by years
expt_num.vec <-ordering(unique(assembly_info.df$experiment), 
                                   order_vector = c(16:22))
#Get experiment numbers and add '0' to single digit numbers.
names(expt_num.vec) <-  paste0('E', 
                              sapply(seq_along(expt_num.vec), function(x){
                                ifelse(nchar(x)==1, 
                                       paste0('0', x),
                                       x)
                              }))

#Manually add Gunalan19 and _Boonkaew20 (in addition to Guna19 and Boonkaew20)
expt_num.vec <- c(expt_num.vec, E07 = 'Gunalan19', E08 = '_Boonkaew20')

#Get the relevant info from samplesheet otherwise loads of rows will be duplicated when joining.
samplesheet_small.df <- mutate(samplesheet, assembly_file = assembly) %>% 
  select(assembly_file, Cat.Sample.stage.s.) %>% unique

assembly_info_samplesheet.df <- left_join(assembly_info.df, samplesheet_small.df, by = 'assembly_file')

assembly_info_samplesheet.df <- mutate(assembly_info_samplesheet.df,
                                        experiment_num = names(expt_num.vec)[match(experiment,
                                                                                  expt_num.vec)])

#Get consistent shapes and colours for the graphs

shapes_val <- 1:length(unique(assembly_info_samplesheet.df$Cat.Sample.stage.s.))
names(shapes_val) <- unique(assembly_info_samplesheet.df$Cat.Sample.stage.s.)

colours_val <- safe_colorblind_palette[1:length(unique(names(expt_num.vec)))]
names(colours_val) <- unique(names(expt_num.vec))

expt_num.df <- data.frame(expt = expt_num.vec, 
                          expt_code = names(expt_num.vec), 
                          col = c(colours_val,colours_val['E07'],colours_val['E08']))

#Get the stages of the life cycle used for the assembly (altering the Gunalan sample name in over1.df since they do not correspond).
over1.df$stage <- assembly_info_samplesheet.df$Cat.Sample.stage.s.[
  match(str_replace(over1.df$assembly, pattern = 'Gunalan19', replacement = 'Guna19'), 
        str_extract(assembly_info_samplesheet.df$assembly, 
                    pattern = '.+(?=_scspades_0.95cdhitest)'))
]

#Get the experiment numbers (altering the Gunalan sample name in over1.df since they do not correspond).
over1.df$experiment_num <- assembly_info_samplesheet.df$experiment_num[
  match(str_replace(over1.df$assembly, pattern = 'Gunalan19', replacement = 'Guna19'), 
        str_extract(assembly_info_samplesheet.df$assembly, 
                    pattern = '.+(?=_scspades_0.95cdhitest)'))
]

#Save the colour scheme
assembly_info_samplesheet.df$expt_colour <- colours_val[match(assembly_info_samplesheet.df$experiment_num,
                                                  names(colours_val))]

#Save number of pir transcripts over 1 TPM (altering the Gunalan sample name in over1.df since they do not correspond).
pir_over1.df <- filter(over1.df, pir)

assembly_info_samplesheet.df$num_mRNA_pir_over200aa_over1TPM <- pir_over1.df$over1_num[
  match(str_extract(assembly_info_samplesheet.df$assembly, 
                    pattern = '.+(?=_scspades_0.95cdhitest)'),
        str_replace(pir_over1.df$assembly, pattern = 'Gunalan19', replacement = 'Guna19'))
]

# Make NAs zero
assembly_info_samplesheet.df$num_mRNA_pir_over200aa_over1TPM <- ifelse(
  is.na(assembly_info_samplesheet.df$num_mRNA_pir_over200aa_over1TPM),
  0,
  assembly_info_samplesheet.df$num_mRNA_pir_over200aa_over1TPM 
)

write.table(assembly_info_samplesheet.df,
            'tl22-01_assembly_info_extra.txt')

```

```{r over1_plots}

(plot <- ggplot(data = filter(over1.df, !pir),
                aes(x = total, 
                    y = over1_perc,
                    col = experiment_num,
                    shape = stage)) +
    geom_point(size = 1, stroke = 2, alpha = .75) +
  scale_shape_manual(values=shapes_val,
                     name = 'Life cycle stage',
                     guide = guide_legend(ncol = 2)) +
  scale_color_manual(values = colours_val[!names(colours_val) %in% 'E10'],
                     name = 'Experiment',
                     guide = guide_legend(ncol = 2)) +
    xlab('Number of assembled transcripts') +
    ylab('Percentage of transcripts transcribed over 1 TPM') +
    theme_classic())

if(global_save_var){
  ggsave(plot = plot,
         'plots/assemblies_numTranscriptsOver1tpm_totalTranscripts.pdf',
         device = 'pdf',
         width = 7,
         height = 5,
         units = "in")
}

(plot <- ggplot(data = filter(over1.df, pir),
       aes(x = total, 
           y = over1_perc_pirs,
           col = experiment_num,
           shape = stage)) +
    geom_point(size = 1, stroke = 2, alpha = .75) +
    #Subset the shapes/colours_val since not all the experiments/stages have any predicted pirs and are hence not included.
    scale_shape_manual(values=shapes_val[
      names(shapes_val) %in% unique(filter(over1.df, 
                                           pir)$stage)
    ],
    name = 'Life cycle stage',
    guide = guide_legend(ncol = 2)) +
    scale_color_manual(values = colours_val[
      names(colours_val) %in% unique(filter(over1.df, 
                                            pir)$experiment_num)
    ],
    name = 'Experiment',
    guide = guide_legend(ncol = 2)) +
   xlab(expression(paste('Number of assembled ',italic(pir),' transcripts'))) +
   ylab(expression(paste('Percentage of ',italic(pir),'s transcribed over 1 TPM'))) +
   theme_classic())

if(global_save_var){
  ggsave(plot = plot,
         'plots/assemblies_numPirOver1tpm_totalTranscripts.pdf',
         device = 'pdf',
         width = 7,
         height = 5,
         units = "in")
}

```