### 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_analysingAssemblyStats.Rmd

---
title: "tl22-01_analysingAssemblyStats"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      results = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
# getwd()
```

```{r load_packages}
library(ggplot2)
library(stringr)
safe_colorblind_palette <- c("#88CCEE", "#CC6677", "#DDCC77", "#117733", "#332288", "#AA4499", 
                             "#44AA99", "#999933", "#882255", "#661100", "#6699CC", "#888888")
```

```{r read_data}
df <- read.delim('tl22-01_assembly_info.txt')
```

#Want to show the stages of development in this.

```{r expt_name}
 accession_table <- read.csv('samples_to_cat.csv')

#Extract the first two numbers of the name e.g. ERX2560125_ERR2541707
df$Cat.Sample.stage.s. <- factor(
  sapply(
    df$assembly, 
    function(assembly_name){
      accession_name <- str_extract(assembly_name, 
                                    pattern = '(SR|ER).{8,9}_(SR|ER).{8,9}(?=_)')
      #If this works (not NA and only two elements) then find the expt name in the table
      if(!is.na(accession_name) & length(unlist(strsplit(accession_name, '_'))) == 2){
        cat_sample_stage <- accession_table$Cat.Sample.stage.s.[grepl(accession_table$fastq_1, pattern = accession_name)]
      } else {
        #Get the first two elements of the name (separated by '_') - split then paste back together.
        expt_name <- paste(str_split_fixed(assembly_name, 
                                      pattern = '_', n = 3)[c(1,2)], collapse = '_')
        if(any(grepl(accession_table$Concatenated.file, 
                     pattern = expt_name))){
          #Otherwise this is probably a re-named concatenated file of samples, so
          # we need the 'Concatenated file' column.
          cat_sample_stage_all <- accession_table$Cat.Sample.stage.s.[grepl(accession_table$Concatenated.file, 
                                                                            pattern = expt_name)]
        } else {
          #Instead of being sensible I had a different naming scheme for the un-concatenated Kim and Gunalan files.
          cat_sample_stage_all <- accession_table$Cat.Sample.stage.s.[grepl(accession_table$file_prefix, 
                                                                            pattern = expt_name)]
        }
        cat_sample_stage <- unique(cat_sample_stage_all)
        if(rlang::is_empty(cat_sample_stage)){
          stop(paste0('No life cycle stage name for assembly ', assembly_name))
        } else if(!length(cat_sample_stage) == 1){
          cat_sample_stage <- 'Mixed.stages'
        }
      }
      return(cat_sample_stage)
    })
)
```

```{r numNovPir_v_totalTranscripts}
(plot <- ggplot(data = df, aes(x = num_transcripts, 
                               y = num_uniq_mRNA_pirHMMerORFs_nomatch95known_over200aa,
                               col = experiment,
                               shape = tolower(Cat.Sample.stage.s.))) +
   geom_point() +
   scale_color_manual(values = safe_colorblind_palette) +
   scale_shape_manual(values=1:nlevels(df$Cat.Sample.stage.s.),
                      name = 'Life cycle stage') +
   xlab('Number of assembled transcripts') +
   ylab('Number of novel pir ORFs') +
   theme_classic() +
   theme(legend.text = element_text(size = 7),
         legend.title = element_text(size = 7),
         legend.box = 'horizontal'))
ggsave(plot = plot,
       'plots/assemblies_numNovPir_v_totalTranscripts.pdf',
       device = 'pdf',
       width = 7,
       height = 5,
       units = "in")
```

```{r numPirORF_v_totalTranscripts}
ggplot(data = df, aes(x = num_transcripts, 
                      y = num_ORFs_predict,
                      col = experiment,
                      shape = tolower(Cat.Sample.stage.s.))) +
  geom_point() +  
  xlab('Number of assembled transcripts') +
  ylab('Number of predicted pir ORFs') +
  scale_color_manual(values = safe_colorblind_palette) +
  scale_shape_manual(values=1:nlevels(df$Cat.Sample.stage.s.),
                     name = 'Life cycle stage') +
  theme_classic() +
  theme(legend.text = element_text(size = 7),
        legend.title = element_text(size = 7),
        legend.box = 'horizontal')
```

```{r numNovel_v_numPirORF}
ggplot(data = df, aes(x = num_ORFs_predict, 
                      y = num_uniq_mRNA_pirHMMerORFs_nomatch95known_over200aa,
                      col = experiment,
                      shape = tolower(Cat.Sample.stage.s.))) +
  geom_point() +
  xlab('Number of predicted pir ORFs') +
  ylab('Number of novel pir ORFs') +
  scale_color_manual(values = safe_colorblind_palette) +
  scale_shape_manual(values=1:nlevels(df$Cat.Sample.stage.s.),
                     name = 'Life cycle stage') +
  theme_classic() +
  theme(legend.text = element_text(size = 7),
        legend.title = element_text(size = 7),
        legend.box = 'horizontal')
```